### 2018-10-19 paper.tex
first commit

\documentclass[10pt]{sigplanconf}

\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{amsmath}
\usepackage[font={small}]{caption}
\newcommand{\mycomment}[1]{}
\newcommand{\dfn}[1]{\textit{#1}} % For initial definitions of terms
\newcommand{\myfig}[3]
 {
 \begin{figure}
 \centerline{\includegraphics[width=#2]{graph/#1.pdf}}
 \caption{\sl \small #3}
 \label{#1:fig}
 \vspace*{-0.15in}
 \end{figure}
 }
\newcommand{\myfigcross}[3]
{
\begin{figure*}[ht]
\centerline{\includegraphics[width=#2]{graph/#1.pdf}}
\caption{\sl \small #3}
\label{#1:fig}
\vspace*{-0.15in}
\end{figure*}
}

% For marking text we're probably dropping.
\newcommand{\ignore}[1]{}

% For marking text that we might like to include in a final draft, if
% accepted to RADIO, but that will not fit within the 5-page limit.
\newcommand{\saveforlater}[1]{}

% For marking important work items.
\newcommand{\todo}[1]{\textbf{TODO: #1}}

\lstset{ %
    captionpos=b,
    frame=single
}

\begin{document}
\title{Bringing the Power of eBPF to Open vSwitch}
\authorinfo{
  \(
  \begin{matrix}
    % Alphabetical order.
    \textrm{William Tu} & \textrm{Joe Stringer} & \textrm{Yifeng Sun} & \textrm{Yi-Hung Wei}\\
    \textrm{u9012063@gmail.com} & \textrm{stringerjoe@vmware.com} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
  \end{matrix}
  \)
}
{VMware Inc.}{}

\maketitle

\begin{abstract}
In this paper we discuss two projects, the OVS-eBPF project, with
the goal of re-writing existing openvswitch kernel module's feature
into eBPF code at TC attachment hook.
And the OVS-AFXDP project, where we

\end{abstract}

\section{Introduction}\label{introduction}
Among the various ways of using eBPF, OVS has been exploring the power of eBPF in three:
(1) attaching eBPF to TC, (2) offloading a subset of processing to XDP, and (3) by-passing
the kernel using AF\_XDP. Unfortunately, as of today, none of the three approaches satisfies
the requirements of OVS. In this paper, we share the challenges we faced, experience learned,
and seek for feedbacks from the community for future direction.

Attaching eBPF to TC started first with the most aggressive goal: we planned to re-implement
the entire features of OVS kernel datapath under net/openvswitch/* into eBPF code.
We worked around a couple of limitations, for example, the lack of TLV support led us to
redefine a binary kernel-user API using a fixed-length array; and without a dedicated way to
execute a packet, we created a dedicated device for user to kernel packet transmission,
with a different BPF program attached to handle packet execute logic.
Currently, we are working on connection tracking.
Although a simple eBPF map can achieve basic operations of conntrack table lookup and commit,
how to handle NAT, (de)fragmentation, and ALG are still under discussion.

Moving one layer below TC is called XDP (eXpress Data Path), a much faster layer for packet
processing, but with almost no extra packet metadata and limited BPF helpers support.
Depending on the complexity of flows, OVS can offload a subset of its flow processing to XDP
when feasible. However, the fact that XDP has fewer helper function support implies that either
1) only very limited number of flows are eligible for offload, or 2) more flow processing logic
needed to be done in native eBPF.

AF\_XDP is another form of XDP but providing a socket interface for control plane and a shared
memory API for accessing packets from userspace application. OVS today has another full-fledged
datapath implementation in userspace, called dpif-netdev, used by DPDK community.
By treating the AF\_XDP as a fast packet-I/O channel, the OVS dpif-netdev can satisfy most of the
features. We are working on building the prototype and evaluating its performance.

Existing OVS consists of three datapath implementations:
dpif-netlink
dpif-netdev
dpif-windows
In this paper, OVS-eBPF project adds a new datapath type
dpif-bpf
and OVS-AFXDP project reuses the dpif-netdev interface, with
linux netdev receiving/sending packets from AFXDP sockets.

\section{Background}

\subsection{OVS Forwarding Model}

\begin{figure}
{\scriptsize
\begin{verbatim}

                +-------------------+
 slow path      |    ovs-vswitchd   |
                +-----^-------------+
                      |       |flow installation
----------------------|-------|-----------------
 fast path     upcall |       | 
 (datapath)  +----------------v---------+ 
  packets--> | parse, lookup, actions   |  --> output
             +--------------------------+
\end{verbatim}
}
\vspace{-1.0em}
\caption{An example of mirrored packet with outer header containing
the GRE and ERSPAN header, followed by the innter Ethernet frame.}
\label{erspanhdr}
\vspace{-1.0em}
\end{figure}

OVS is widely used in virtualized data center environments as a software
switching layer inside various operating systems, including FreeBSD,
Windows Hyper-V, Solaris and Linux. As shown in
Figure~\ref{ovsintro:fig}, the architecture of OVS consists of two major
components: a slow path and a fast path. OVS begins processing packets
in its datapath, the fast path, shortly after the packet is received by the NIC in the host OS.
The OVS datapath first performs packet parsing to extract relevant protocol
headers from the packet and store it locally in a manner that is efficient for
performing lookups (flow key), then it uses this information to look into the
match/action cache (flow table) and determines what needs to be done for this
packet. If there is no match in the flow table, the datapath passes the packet
from the kernel up to the slow path, \verb+ovs-vswitchd+, which maintains the
full determination of what needs to be executed to modify and forward
the packet correctly.  This process is called packet {\em upcall} and
usually happens at the first packet of a flow seen by the OVS datapath.
If the packet matches in this flow table, then the OVS datapath executes its
corresponding actions from the flow table lookup result and updates its flow
statistics.

In this model, the \verb+ovs-vswitchd+ determines how the packet should be handled,
and passes the information to the datapath inside the kernel using a
Linux generic netlink interface.  Over the years the OVS datapath features evolved.
The initial OVS datapath used a microflow cache for its flow table,
essentially caching exact-match entries for each
transport layer connection's forwarding decision.  And in later versions,
two layers of caching were used: a microflow cache and a megaflow cache,
which caches forwarding decisions for traffic aggregates beyond individual
connections.  In recent versions of OVS, datapath implementations include
features such as connection tracking, stateful network address translation, and
support for layer 3 tunneling protocols.

\mycomment{
As each new feature in OVS datapath requires deep knowledge and programming
paradigm of the target operating system, e.g., how Linux packets are received,
how Windows kernel module does firewalling, and each operating system
community has its own software upstream and maintenance processes, the software
development cycle of a new datapath feature becomes complex and requires
detailed platform-specific knowledge. Each platform introduces its own
development and maintenance burden, and it is common for new features to take
six months or more before implementations are available across all datapaths.
As a result, providing a programmable OVS datapath which allows extending the
features flexibly with lower maintenance cost becomes a critical requirement.
}

\subsection{eBPF Basics}\label{sec:ebpf}
Berkeley Packet Filter, BPF, is an instruction set architecture
proposed by Steven McCanne and Van Jacobson in 1993~\cite{cbpf}.  
BPF was designed as a generic packet filtering solution and is widely
used by every network operator today, through the well-known tcpdump/wireshark
applications. A BPF interpreter is attached early in the packet receive call
chain, and it executes a BPF program as a list of instructions.
A BPF program typically parses a packet and decides
whether to pass the packet to a userspace socket.  With its simple architecture
and early filtering decision logic, it can execute this logic efficiently.

%P2 from cBPF to eBPF
For the past few years, the Linux kernel community has improved the traditional
BPF (now renamed to classic BPF, cBPF) interpreter inside the kernel with
additional instructions, known as extended BPF (eBPF). eBPF was introduced with
the purpose of broadening the programmability of the Linux kernel. Within the
kernel, eBPF instructions run in a virtual machine environment. The virtual
machine provides a few registers, stack space, program counter, and a way to
interact with the rest of the kernel through a mechanism called helper
functions.  Similar to cBPF, eBPF operates in an event-driven model on a
particular hook point; each hook point has its own execution {\em context} and
execution at the hook point only starts when a particular type of event fires.
A BPF program is written against a specific context. For example, a BPF program
attached to a raw socket interface has a context which includes the packet, and
the program is only triggered to run when there is an incoming packet to the
raw socket.

% eBPF map, helper, tail call and syscall
%helper
The eBPF virtual machine provides a completely isolated environment for its
bytecode running inside; in other words, it cannot arbitrarily call other
kernel functions or access into memory outside its own environment. To interact
with the outside world, the eBPF architecture white-lists a set of helper
functions that a BPF program can call, depending on the {\em context} of the
BPF program.  For example, a BPF program attached to raw socket in a packet
context could invoke VLAN push or pop related helper functions, while a BPF
program with a kernel tracing context could not. %The complete list of helper
%functions can be found at \verb+linux/include/uapi/linux/bpf.h+.

%map
To store and share state, eBPF provides a mechanism to interact with a variety
of key/value stores, called \textit{maps}. eBPF maps reside in the kernel, and can be
shared and accessed from eBPF programs and userspace applications. eBPF
programs can access maps through helper functions, while userspace applications
can access maps through BPF system calls. There are a variety of map types for
different use cases, such as hash tables or arrays. These are created by a
userspace program and may be shared between multiple eBPF programs running in
any hook point.

%tail call
Finally, eBPF tail call~\cite{tailcall} is a mechanism allowing one eBPF program to trigger execution of 
another eBPF program, providing users the flexibility of composing a chain of
eBPF programs with each one focusing on particular features.  Unlike a
traditional function call, this tail call mechanism calls another program
without returning back to the caller's program. The tail call reuses the
caller's stack frame, which allows the eBPF implementation to minimize call
overhead and simplifies verification of eBPF programs.
% end of bpf intro

\subsection{XDP: eXpress Data Path}
\myfig{tcebpf}{2.6in}{The workflow of TC and XDP eBPF development process and its
packet flow.  The eBPF program compiled by LLVM+clang is loaded into the kernel
using iproute. The kernel runs the program through a verification stage, and
subsequently attaches the program to the TC/XDP ingress hook point.  Once
successfully loaded, an incoming packet received by XDP/TC ingress will execute the
eBPF program.}

There are several hook points where eBPF programs may be attached in recent
Linux kernels. XDP is still an eBPF program, but its attachment point is at the
lowest level of the network stack.  Due to its unique attachment point, the XDP
has its own {\em context}; the input parameters to the XDP eBPF program and return
values have different meanings than the TC eBPF program. XDP shows performance
closed to line rate of the device, because the XDP program is triggered immediately
at the network device driver's packet receiving code path.
Due to its lowest hook point at the networking stack, XDP input parameter has only
pointer to the beginning and end of the packet data plus a few metadatas.
XDP also supports accessing to eBPF maps and tail calls, but with much
less number of helper functions than the eBPF program at TC hook.

Figure~\ref{tcebpf:fig}
shows the typical workflow for installing an eBPF program to the TC/XDP hook point,
and how packets trigger eBPF execution.  Clang and LLVM takes a program
written in C and compiles it down to the eBPF instruction set, then emits an
ELF file that contains eBPF instructions.  An eBPF loader, such as iproute,
takes the ELF file, parses the programs and maps information from it and
issues BPF syscalls to load the program.  If it passes the BPF verifier,
then the program is attached to the hook point (in this case, TC/XDP), and
subsequent packets through the TC/XDP ingress hook will trigger execution of the
eBPF programs.

\subsection{AF\_XDP Socket}
% afxdp
AF\_XDP is a new Linux address family that aims for high packet I/O
performance. Traditionally, a userspace program receives packets from
kernel through the socket API.  By creating a socket with address family
such as AF\_PACKET, the userspace program can receive/send the raw packets
at the device driver layer.  Although the AF\_PACKET family has been using
in many places such as tcpdump, its performance does not catch up the the
recent high speed network devices, such as 40G/100G NICs.
Performance evaluation~\cite{danialafpacket,johnafxdp} of AF\_PACKET
shows less than 2 million packets per second using single core.

AF\_XDP was proposed and upstreamed to Linux kernel since 4.18~\cite{afxdp} 
The core idea behind the AF\_XDP is to leverage the XDP eBPF program's
early access to the raw packet, and create a high speed channel from the XDP to
a userspace socket interface. In other word, AF\_XDP socket family connects the
XDP packet receving/sending path to the user space, bypassing the rest of the
Linux networking stacks.
An AF\_XDP socket, called XSK, is create using the normal socket() system
call. Unlike AF\_PACKET using the send() and receive() syscalls,
XSK introduces two rings in userspace: the RX ring and the TX ring.
The userspace program using XSK needs to properly configure and maintain
the RX and TX ring structure in order to receive and send packets.

For example, to receive packets, a set of RX descriptors pointing to empty
packet buffer needs to be filled into the RX ring.  When packets arrives,
the userspace program fetches the packet data from the desciptors, and
refill the empty buffer to the RX ring structure, in order to receive new
incoming packets.  For sending packets, a set of descriptors pointing the
packet buffer with contents is filled to the TX ring (FILL queue).
Then, the userspace
program issues send() system call.  It is up to the userspace program to
make sure whether the packets have been sent or not, by checking the TX
ring structure.

All the packet data buffers used in TX/RX ring are allocated from a specific
memory area called UMEM which consists of a number of fixed size chunks.
A descriptor in RX/TX ring points to the element in UMEM by its address.
The address is simply an offset within the entire UMEM memory region.
The UMEM also has two rings: the FILL ring and the COMPLETION ring.
The Fill ring is used by the application to send down addr for the kernel
to fill in with RX packet data. References to these frames will then
appear in the RX ring once each packet has been received. The 
completion ring, on the other hand, contains frame addr that the 
kernel has transmitted completely and can now be used again by user
space, for either TX or RX.

Thus, the frame addrs appearing in the 
completion ring are addrs that were previously transmitted using the 
TX ring. In summary, the RX and FILL rings are used for the RX path
and the TX and COMPLETION rings are used for the TX path.

\section{OVS eBPF Datapath}
\mycomment{
\begin{figure}
{\scriptsize
\begin{verbatim}

               _
              |   +-------------------+
              |   |    ovs-vswitchd   |
              |   +-------------------+
    userspace |   |      ofproto      |<-->OpenFlow controllers
              |   +--------+-+--------+  
              |   | netdev | |ofproto-|
              |   +--------+ |  dpif  |
                  | netdev | +--------+
   *eBPF hook --> |provider| |  dpif  |
                  +---||---+ +--------+
              |       ||     |  dpif  | <--- *eBPF provider
              |       ||     |provider|
              |_      ||     +---||---+
                      ||         ||
               _  +---||-----+---||---+
              |   |          |datapath| <--- *eBPF datapath
       kernel |   |          +--------+
              |   |                   |
              |_  +--------||---------+
                           ||
                        physical
                           NIC
\end{verbatim}
}
\vspace{-1.0em}
\caption{An example of mirrored packet with outer header containing
the GRE and ERSPAN header, followed by the innter Ethernet frame.}
\label{erspanhdr}
\vspace{-1.0em}
\end{figure}
\subsection{Tunnel Support}
\subsection{Megaflow Support}
\subsection{Connection Tracking Support}
}

\section{Userspace OVS with AF\_XDP}
\begin{figure}
{\scriptsize
\begin{verbatim}
              |   +-------------------+
              |   |    ovs-vswitchd   |<-->ovsdb-server
              |   +-------------------+
              |   |      ofproto      |<-->OpenFlow controllers
              |   +--------+-+--------+ 
              |   | netdev | |ofproto-|
    userspace |   +--------+ |  dpif  |
              |   | AF_XDP | +--------+
              |   | netdev | |  dpif  |
              |   +---||---+ +--------+
              |       ||     |  dpif- |
              |       ||     | netdev |
              |_      ||     +--------+  
                      ||         
               _  +---||-----+--------+
              |   | XDP program +     |
 Linux kernel |   |   eBPF map        |
              |_  +--------||---------+
                           ||
                        physical
                           NIC
\end{verbatim}
}
\vspace{-1.0em}
\caption{OVS Architecture AF\_XDP}
\label{ovsafxdp}
\vspace{-1.0em}
\end{figure}

\subsection{Datapath and Netdev}
OVS includes a userspace datapath interface (dpif) implemenation, called dpif-netdev.
Unlike the dpif-netlink which talks to the Linux Open vSwitch kernel
module through netlink API, the dpif-netdev works completely in userspace.
As a result, dpif-netdev expects to receive and send packets from its userspace
interface.  One major use case of dpif-netdev is OVS-DPDK~\cite{ovsdpdk},
where the packet reception and transmission are all done in DPDK's userspace
library. The dpif-netdev is designed to be agnostic to how the network
device accessing the packets, by an abstraction layer called netdev.
So Packets might come from DPDK packet I/O library,
a Linux AF\_PACKET socket API, or in our case, the AF\_XDP socket interface,
as long as each mechanism implements its own netdev interface.
Once dpif-netdev receives a packet, it follows the same mechanism doing
parse, lookup the flow table, and apply actions to the packet.

Figure~\ref{ovsafxdp} shows the architecture of Userspace OVS with AF\_XDP.
We implement a new netdev type called netdev-afxdp, which receives and transmits
packets using the XSK.
In order for XSK to work, we insert a XDP program and a eBPF map with   
interact with XDP program to forward packets to the AF\_XDP socket.
Once the AF\_XDP netdev receives a packet, it passes the packet to the
dpif-netdev for packet processing.

\mycomment{
/* PMD: Poll modes drivers.  PMD accesses devices via polling to eliminate
 * the performance overhead of interrupt processing.  Therefore netdev can
 * not implement rx-wait for these devices.  dpif-netdev needs to poll
 * these device to check for recv buffer.  pmd-thread does polling for
 * devices assigned to itself.
 *
 * DPDK used PMD for accessing NIC.
 *}

\subsection{AF\_XDP netdev configuration}
ovs-vswitchd controls and manages the XDP program and maps.
When users attach a AF\_XDP netdev to the OVS bridge, the netdev-afxdp
does the following steps:
We implemnet most of the logic in lib/netdev-linux.c
{
\begin{verbatim}
    ovs-vsctl add-br br0
    ovs-vsctl add-port br0 eth0 -- \
        set int eth0 type="afxdp"
\end{verbatim}
}
\begin{enumerate}
\item Attach a XDP program to the netdev queue:
The XDP program OVS attaches is fixed and consists
of only a few line of code, which receives the packet and
calls the \texttt{bpf\_redirect\_map()} helper function with
the XSK eBPF maps and key equals zero.

\item Create a AF\_XDP socket:
This requires creating the XSK socket by calling socket() syscall,
setting up its RX and TX ring buffer, allocating a UMEM region,
and setting up FILL/COMPLETION ring of the UMEM.

\item Load and configure the XSK eBPF map:
The XSK eBPF maps consists of key value pairs, where key is an u32 index and value is
the netdev's queue ID.  In our case, we simply program one entry to the map with
key equals zero and the file descriptor, fd, of the XSK as its value.
As a result, the XDP program calling \texttt{bpf\_redirect\_map} with key=0 will
forwards the packet to the XSK.
\end{enumerate}

When a netdev-afxdp is detached or closed by user,
ovs-vswitchd closes the XSK socket, free its memory, and unload the eBPF program and map.

\subsection{Packet Reception and Transmission}

When receiving the packets from XSK, netdev-afxdp implements the following
operations:
\begin{enumeration}
\item Get N free elements from UMEM
\item Fille

\end{enumeration}
When sending out the packets

\subsection{Performance Evaluation}
PMD
HugePage
UMEM element list
Batching

\bibliography{paper.bib}
\bibliographystyle{plain}

\end{document}

### 2018-10-21 paper.tex
add afxdp txrx

@@
 
+% concept
+Application still HW agnostic with ZC
+Each application gets its own packet buffer and TX/RX descriptor rings
+Packet buffers can be shared if desired
+  TX/RX descriptor rings always private to process
+% AF\_XDP has 3 mode
+% zero copy, native mode,
+XDP\_SKB:
+– Works on any netdevice using sockets and generic XDP path
+XDP\_DRV:
+– Works on any device with XDP support (all three NDOs)
+XDP\_DRV + ZC:
+– Need buffer allocator support in driver + a new NDO for TX
 All the packet data buffers used in TX/RX ring are allocated from a specific
@@
 
-\mycomment{
-/* PMD: Poll modes drivers.  PMD accesses devices via polling to eliminate
- * the performance overhead of interrupt processing.  Therefore netdev can
- * not implement rx-wait for these devices.  dpif-netdev needs to poll
- * these device to check for recv buffer.  pmd-thread does polling for
- * devices assigned to itself.
- *
- * DPDK used PMD for accessing NIC.
- *}
-
 \subsection{AF\_XDP netdev configuration}
@@
 }
+
 \begin{enumerate}
@@
 forwards the packet to the XSK.
+
+\item Initialize the FILL queue
+Before any packet arrival, the rest three queues: RX, TX and COMPLETION
+remains the same
 \end{enumerate}
@@
 
+memory allocation:
+
+% 1 UMEM per XSK per queue, no sharing
+For each netdev, ovs-vswitchd attach XSK to each 
+For each XSK
+memorypool
+We implement a UMEM freelist to keep track of the used/unused element
+
+To handle multiple 
 \subsection{Packet Reception and Transmission}
+\begin{figure}
+{\scriptsize
+\begin{verbatim}
+UMEM: total 8 elements, each has 2K chunk size
+                 +-------------------------------+
+                 |   |   |   |   |   |   |   |   |
+                 +-------------------------------+
+            addr:  1   2   3   4   5   6   7  8 
+(1) Initial Stat
+           -------------------          UMEM freelist
+FILL Qu  ...| 1 | 2 | 3 | 4 | ...       {5, 6, 7, 8}
+           -------------------
+           -------------------
+RX ring  ...|   |   |   |   | ...
+           -------------------
+(2) 4 Packets Arrive
+           -------------------          UMEM freelist
+FILL Qu  ...|   |   |   |   | ...       {5, 6, 7, 8}
+          --------------------
+           -------------------
+RX ring  ...| 1 | 2 | 3 | 4 | ...
+           -------------------
+(3) Get 4 free UMEM elemet and refill
+           -------------------          UMEM freelist
+FILL Qu  ...| 5 | 6 | 7 | 8 | ...       {} 
+           -------------------
+           -------------------
+RX ring  ...| 1 | 2 | 3 | 4 | ...
+           -------------------
+(4) Pass packets {1, 2, 3, 4} to dpif-netdev
+    for parse, lookup, and actions.
 
-When receiving the packets from XSK, netdev-afxdp implements the following
-operations:
-\begin{enumeration}
-\item Get N free elements from UMEM
-\item Fille
+(5) Recycle to UMEM freelist, goto step (2)
+            +-------------------        UMEM freelist
+FILL Qu  ...| 5 | 6 | 7 | 8 | ...       {1, 2, 3, 4}
+            +------------------
+           -------------------
+RX ring  ...|   |   |   |   | ...
+           -------------------
+\end{verbatim}
+}
+\vspace{-1.0em}
+\caption{An example of how OVS programs the FILL queue and RX ring when
+processing incoming AF\_XDP packets.}
+\label{afxdprx}
+\vspace{-1.0em}
+\end{figure}
+Figure~\ref{afxdprx} shows how ovs-vswitchd sets up the FILL queue and
+the RX ring for receving packets from XSK.  For simplicity to demonstrate
+how AF\_XDP works, assume that there are only eight UMEM buffers, with each buffer's
+size equals 2KB.  Initially at step 1, ovs-vswitchd fills four empty UMEM elem into
+the FILL queue and waits for incoming packets.  When there are incoming packets,
+at step 2, the FILL queue's four buffer elems have been consumed and moved
+to the RX ring.  In order to keep receiving packets, ovs-vswitchd gets another
+four empty UMEM elems from thte UMEM freelist, and fills into FILL queue (step 3).
+Then ovs-vswitchd creates the metadata needed for the four packet
+buffer, i.e., struct dp\_packet and struct dp\_packet\_batch, and passes to 
+the dpif-netdev layer for parse, lookup and action executions.
+Finally, when ovs-vswitchd finishes processing the UMEM buffer, a recycle
+mechanism is triggered to place these buffer back to UMEM freelist (step 5).
+Step 5 makes sure that there are always available elems in FILL queue, so
+that the underlying XDP program in kernel can keep placing packets into
+the FILL queue while the userspace ovs-vswitchd is processing the previous
+received packets.
+When step 5 finishes, ovs-vswitchd goes back to step 2, waiting and processing
+new packets.
 
-\end{enumeration}
-When sending out the packets
+\begin{figure}
+{\scriptsize
+\begin{verbatim}
+ dpif-netdev has 4 packets at UMEM:{1, 2, 3, 4}
+ from eth1 and plan to send to eth2.
+(1) Get 4 free element from eth2 UMEM, copy packet data to it
+(2) Create TX descriptors and place into TX ring
+           -------------------        eth2 UMEM freelist
+TX ring  ...| 4 | 5 | 6 | 7 | ...       {1, 2, 3, 8}
+           -------------------
+           -------------------
+COMPL q  ...|   |   |   |   | ...
+           -------------------
+(3) Issue send syscall to eth2's XSK
+(4) Packet transmission is completed
+           -------------------        eth2 UMEM freelist
+TX ring  ...|   |   |   |   | ...       {1, 2, 3, 8}
+           -------------------
+           -------------------
+COMPL q  ...| 4 | 5 | 6 | 7 | ...
+           -------------------
+(5) Recycle to eth2 UMEM
+           -------------------        eth2 UMEM freelist
+TX ring  ...|   |   |   |   | ...       {1, 2, 3, 4,
+           -------------------           5, 6, 7, 8}
+           -------------------
+COMPL q  ...|   |   |   |   | ...
+           -------------------
+(6) Recycle {1, 2, 3, 4} to eth1's UMEM 
+\end{verbatim}
+}
+\vspace{-1.0em}
+\caption{An example of how OVS programs the COMPLETION queue and TX ring when
+processing incoming AF\_XDP packets from one netdev and sending to another netdev.}
+\label{afxdptx}
+\vspace{-1.0em}
+\end{figure}
+
+Figure~\ref{afxdptx} shows the process for sending packets to the XSK.
+Assuming that ovs-vswitch has a a bridge and two ports: eth1 and eth2.
+And there is a flow entry showing
+\texttt{in\_port=eth1, ..., actions=..., output:eth2}, meaning to forwarding
+packets received from eth1 to eth2.
+Assuming that both eth1 and eth2 are configured with AF\_XDP support.
+
+Initially, assume ovs-vswitchd receives four packets {1, 2, 3, 4} from eth1's XSK
+To send to eth2 using its XSK, ovs-vswitchd first gets four packet buffers,
+{4, 5, 6, 7}, from eth2's UMEM freelist and copy the packet data to their
+packet buffer in UMEM.  Then TX descriptors for the four packets are
+created and placed into eth2's TX ring (step 2).
+At step 3, send syscall is issued to signal the kernel to start transmit.
+As the send in XSK is asynchronos and send syscall only return return zero when
+no error, ovs-vswitch has to poll the COMPLETION queue to make sure these four
+packets have been sent out (step 4).
+Once the four packets' descriptors shown up at the COMPLETION ring, at step 5,
+ovs-vswitchd recycles their UMEM elememnts back to the eth2's UMEM freelist.
+In addition, the original four packet buffer from eth1, {1, 2, 3, 4}, is
+recycled back to eth1's UMEM freelist.
 
 \subsection{Performance Evaluation}
+When sending and receivng packets using AF\_XDP with OVS, two CPUs show
+100\% utilitization:
+\begin{itemize}
+\item ksoftirqd:
+\item ovs-vswitchd:
+\end{itemize}
 PMD
@@
 Batching
+2-CPU operation
+non-blocking send
+non-block receive
+
+\mycomment{
+/* PMD: Poll modes drivers.  PMD accesses devices via polling to eliminate
+ * the performance overhead of interrupt processing.  Therefore netdev can
+ * not implement rx-wait for these devices.  dpif-netdev needs to poll
+ * these device to check for recv buffer.  pmd-thread does polling for
+ * devices assigned to itself.
+ *
+ * DPDK used PMD for accessing NIC.
+ *}
 

### 2018-10-21 paper.tex
afxdp performance

@@
 \subsection{Performance Evaluation}
-When sending and receivng packets using AF\_XDP with OVS, two CPUs show
+% testbed
+
+When forwarding packets using AF\_XDP with OVS, two CPUs show up
 100\% utilitization:
 \begin{itemize}
-\item ksoftirqd:
-\item ovs-vswitchd:
+\item ovs-vswitchd: This is the process keeps doing the send and receive steps
+in Figure~\ref{afxdprx,afxdptx}.
+\item ksoftirqd: This is the kernel software interrupt thread handling the
+incoming packets, triggering XDP program to pass to the XSK, and also processing
+transmission.
 \end{itemize}
+
+For all our experiments, we use a microbenchmark program, called xdpsock,
+as the baseline to compare with our implementation.
+xdpsock is an AF\_XDP sample program doing no packet processing, and can be configured
+to do simply dropping the packet, rxdrop, or forwarding the packet to the same port as
+it receives, l2fwd.  In our testbed, we measured $19Mpps$ for rxdrop and $17Mpps$ for
+l2fwd.
+
+Compared to xdpsock, ovs-vswitchd adds on top of it a UMEM memory pool,
+and packets parsing, lookup and action execution.
+Our initial prototype shows pretty terrible performance, only $0.5Mpps$ rxdrop.
+We investigated and our first fix is to enable OVS's PMD-mode,
+Poll-Mode-Driver, to the afxdp netdev.
+In OVS, a non-pmd mode netdev does packet receiption by putting all the receiving
+netdev's file descript together and ovs-vswitch polls them when there is an fd ready
+to perform I/O.  As a result, the fd of the XSK is {\em shared} with other fds, and
+we observed that the poll system call has pretty high overhead.
+
+Applying OVS's PMD netdev avoids these problems and improving the rxdrop performance
+from $0.5Mpss$ to $3Mpps$.  Ideally, receing packets from XSK does not need to context
+switch to into kernel as mentioned in steps~\ref{afxdprx}. So an ideal implementation
+should show minimal context switches with most of the CPU stays in userspace. 
+When enabling PMD netdev for afxdp, ovs-vswitch can be
+configured to create a {\em dedicated} thread for the packet I/O.  
+In our case, we create a thread for each XSK receive queue which keeps polling
+the RX ring for new packets.  Each round of receive polling processes a batch of
+packet, up to maximum of 32 packets. In addition, we enable high memory page 1GB to
+reduce the page fault overhead.  With the above fixes/optimizations, the rxdrop
+performance increases to around $10Mpps$
+
+For every stage of the optimization, we use Linux perf extensively, e.g., perf stat
+and perf record/report.  With the above new design, we observe the new bottleneck
+is at the UMEM memory pool we introduce to maintain the UMEM freelist, and the
+packet metadata allocation, struct dp\_packet, for every receiving packet.
+We introduce two major API, umem\_elem\_get(), which gets N free element from the
+UMEM freelist, and umem\_elem\_put(), which places back the free UMEM buffer to
+the freelist.
+producer 
+consumer
+We implemented three data structures of UMEM memory pool as below
+to compare their performance
+\begin{itemize}
+\item list\_head: Similar to Linux kernel's list\_head, each element contains it
+UMEM addr, and a next pointer points to next free element. Elements are sparse in
+memory. Consumer always takes the first element, head, and producer also place back
+from the head of the list.
+\item ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
+allocated in continuous memory region, and pointers (head and tail) point for
+Consumer takes elements from the tail and when placing back, producer, puts
+them into where head pointer points to.
+\item ptr\_stack: Similar to ptr\_ring, but consumer and producer always use
+the top few element, thus similar to push pop. Compared
+
+\end{itemize}
+% lock
+% push/pop
+% preallocate array
+Since the UMEM
+
+We then observe that the UMEM memory pool implementation ha
+
 PMD
@@
 non-block receive
+pre-allocation dp\_packet
+rxhash
 

### 2018-10-22 paper.tex
afxdp experiment

@@
 linux netdev receiving/sending packets from AFXDP sockets.
+
+Table~\ref{compare} shows the difference between two projects:
+\begin{table}
+\centering
+\scriptsize
+\begin{tabular}{|c | c | c|}
+\hline
+ {\bf Projects} & {\bf OVS-eBPF } & {\bf OVS-AFXDP} \\ \hline\hline
+%  Performance & $<$2Mpps & $>14$Mpps \\ \hline
+  Performance & Low & High \\ \hline
+  Driver Support & No & Yes \\ \hline
+  Minimal Kernel & 4.18 & Unknown \\ \hline
+  BPF Code Size & Large & Minimal \\ \hline
+  Extensibility & Low & High \\ \hline
+  \end{tabular}
+\caption{\footnotesize
+Comparison of the OVS-eBPF project and OVS-AFXDP project.
+}
+\label{compare}
+\end{table}
 
@@
 
+Define:
+baseline rxdrop vs OVS-AFXDP rxdrop
+baseline l2fwd vs OVS-AFXDO l2fwd
 \subsection{Performance Evaluation}
-% testbed
+
+\subsubsection{Testbed}
 
@@
 Poll-Mode-Driver, to the afxdp netdev.
+
+\subsubsection{PMD netdev}
 In OVS, a non-pmd mode netdev does packet receiption by putting all the receiving
@@
 
+\subsubsection{Memory Management and Locks}
 For every stage of the optimization, we use Linux perf extensively, e.g., perf stat
@@
 them into where head pointer points to.
-\item ptr\_stack: Similar to ptr\_ring, but consumer and producer always use
-the top few element, thus similar to push pop. Compared
+\item ptr\_stack: Similar to ptr\_ring, an array of elements are allocated
+in continuous memory region, but both consumer and producer takes elements
+from beginning of the array.
+\end{itemize}
 
+Although we allocate one UMEM per netdev, there might be multiple queues per
+netdev sharing the same UMEM. As a result, the above three data structures
+all require a mutex lock before accessing UMEM freelist. Linux perf reports
+pthread\_mutex\_lock as one of the top CPU utilization function, and overall
+shows around $2Mpps$ overhead for rxdrop experiments.
+We change our design by 1) allocating per-queue UMEM region and 2)
+allocating one pmd thread per queue. As a result, no lock is needed because
+the each queue has only one thread and its own set of UMEM elements.
+
+%\subsubsection{Metadat Allocation}
+With the above change, Linux perf shows that the packets metadata allocation
+takes a lot of CPU cycles, the dp\_packet\_init, dp\_packet\_set\_data.
+So instead of allocating packet metadata at packet reception time, we
+implement two data structures and compare their performance:
+\begin{itemize}
+\item Inlining in UMEM buffer: Since we already allocate 2K chunk for each
+UMEM buffer, we reserve the first 256-byte in each UMEM element as struct dp\_packet
+and initialize the dp\_packet as much as we can at allocation time.
+So when packets arrive, each packet already has its packet metadata at a const
+offset ahead.
+\item Array on heap: This design simply allocate a continuous memory region
+storing an array of packet metadata, and initialize them as much as we can.
 \end{itemize}
-% lock
-% push/pop
-% preallocate array
-Since the UMEM
 
-We then observe that the UMEM memory pool implementation ha
+%complete separation of the allocation of metadata structures from the allocation of packet data buffers.
+% see mbuf dpdk https://doc.dpdk.org/guides/prog_guide/mbuf_lib.html
+%Embed metadata within a single memory buffer the structure followed by a fixed size area for the packet data.
+%Use separate memory buffers for the metadata structure and for the packet data.
 
-PMD
-HugePage
-UMEM element list
-Batching
-2-CPU operation
-non-blocking send
-non-block receive
-pre-allocation dp\_packet
-rxhash
+With the above design change, we found that using the ptr\_stack with
+array on heap get the best performance. Largely due to both data structures
+have better spatial locality and more batching friendly.
+For example, setting up 32 packet metadatas in an array definitely incurs
+less cache misses than setting up in 32 elements in UMEM.
+And getting 32 free UMEM elements from ptr\_stack has similar benefits.
+With the above decision, the OVS rxdrop shows $17Mpps$, only $2Mpps$ less
+than the baseline xdpsock rxdrop.
 
-\mycomment{
-/* PMD: Poll modes drivers.  PMD accesses devices via polling to eliminate
- * the performance overhead of interrupt processing.  Therefore netdev can
- * not implement rx-wait for these devices.  dpif-netdev needs to poll
- * these device to check for recv buffer.  pmd-thread does polling for
- * devices assigned to itself.
- *
- * DPDK used PMD for accessing NIC.
- *}
+\subsubsection{Batching Send Syscall}
+%574.790 M/sec  
+We continued measured the performance of OVS-AFXDP l2fwd and.
+observed only $4Mpps$, compared to $17Mpps$ baseline xdpsock.
+Again, by analysis using perf, we found that the OVS pmd thread under rxdrop
+has much fewer context switches compared to the l2fwd, indicating that
+the PMD process spends much more time in kernel space than in userspace.
+By using strace, we saw that the l2fwd experiement calls sendto system call
+at very high frequeuce, because from Figure~\ref{afxdptx}, we design to check
+the success of send (step 4) immediately after issusing send (step 3).
+We change the design by calling send syscall (step 3) only when TX ring is
+closed to full, e.g., when 3/4 ring elements have been used.
+In specific case, instead of issuing send syscall for a batch of 32 packets
+and making sure they are done, we only issue send when there are 512 outstanding
+packets.  With this change, the OVS-AFXDP l2fwd experiment shows around $14Mpps$.
+
+%rxhash
 

### 2018-10-22 paper.tex
revise

@@
 \begin{abstract}
-In this paper we discuss two projects, the OVS-eBPF project, with
-the goal of re-writing existing openvswitch kernel module's feature
-into eBPF code at TC attachment hook.
-And the OVS-AFXDP project, where we
+eBPF is an emerging technology in Linux kernel with the goal of making Linux kernel
+extensible by providing an eBPF virtual machine with safety guarantee.
+Open vSwitch, OVS, is a software switch running majorly in Linux
+operating systems and interacts with Linux kernel module,
+the openvswitch.ko.
+In this paper we describe our efforts to make use of eBPF technology in OVS.
+We present the design, implementation, and evaluation of two projects:
+the OVS-eBPF project and the OVS-AFXDP project.
+OVS-eBPF has the goal of re-writing existing flow processing features
+in openvswitch kernel module into eBPF code.
+On the other hand, the OVS-AFXDP project aims to by-passing the kernel
+using AF\_XDP socket and moving most of the flow processing features
+in userspace.
 
@@
 (1) attaching eBPF to TC, (2) offloading a subset of processing to XDP, and (3) by-passing
-the kernel using AF\_XDP. Unfortunately, as of today, none of the three approaches satisfies
-the requirements of OVS. In this paper, we share the challenges we faced, experience learned,
-and seek for feedbacks from the community for future direction.
-
+the kernel using AF\_XDP.
 Attaching eBPF to TC started first with the most aggressive goal: we planned to re-implement
@@
 kernel through the socket API.  By creating a socket with address family
-such as AF\_PACKET, the userspace program can receive/send the raw packets
+such as AF\_PACKET, userspace programs can receive/send the raw packets
 at the device driver layer.  Although the AF\_PACKET family has been using
-in many places such as tcpdump, its performance does not catch up the the
+in many places such as tcpdump, its performance does not catch up with the
 recent high speed network devices, such as 40G/100G NICs.
@@
 The core idea behind the AF\_XDP is to leverage the XDP eBPF program's
-early access to the raw packet, and create a high speed channel from the XDP to
+early access to the raw packet, and create a high speed channel from the XDP directly to
 a userspace socket interface. In other word, AF\_XDP socket family connects the
-XDP packet receving/sending path to the user space, bypassing the rest of the
+XDP packet receving/sending path to the userspace, bypassing the rest of the
 Linux networking stacks.
 An AF\_XDP socket, called XSK, is create using the normal socket() system
-call. Unlike AF\_PACKET using the send() and receive() syscalls,
+call. Unlike AF\_PACKET which uses the send() and receive() syscalls with packet buffer as parameter,
 XSK introduces two rings in userspace: the RX ring and the TX ring.
@@
 the RX and TX ring structure in order to receive and send packets.
+In addition, all the packet data buffers used in TX/RX ring are allocated from a specific
+memory area called UMEM which consists of a number of fixed size chunks.
+The UMEM also has two rings: the FILL ring and the COMPLETION ring.
+A descriptor in RX/TX ring or in FILL/COMPLETION ring points to the element
+in UMEM by its address.  The address is not system's virtual or physical address
+but simply an offset within the entire UMEM memory region.
 
-For example, to receive packets, a set of RX descriptors pointing to empty
-packet buffer needs to be filled into the RX ring.  When packets arrives,
-the userspace program fetches the packet data from the desciptors, and
-refill the empty buffer to the RX ring structure, in order to receive new
+For example, to receive packets, a set of descriptors pointing to empty
+packet buffer needs to be filled into the FILL ring.  When packets arrives,
+the userspace program checks the RX ring, fetches the packet data from the
+desciptors, and refills the empty buffer to the FILL ring structure, in order to receive new
 incoming packets.  For sending packets, a set of descriptors pointing the
-packet buffer with contents is filled to the TX ring (FILL queue).
-Then, the userspace
-program issues send() system call.  It is up to the userspace program to
-make sure whether the packets have been sent or not, by checking the TX
-ring structure.
-
-% concept
-Application still HW agnostic with ZC
-Each application gets its own packet buffer and TX/RX descriptor rings
-Packet buffers can be shared if desired
-  TX/RX descriptor rings always private to process
-% AF\_XDP has 3 mode
-% zero copy, native mode,
-XDP\_SKB:
-– Works on any netdevice using sockets and generic XDP path
-XDP\_DRV:
-– Works on any device with XDP support (all three NDOs)
-XDP\_DRV + ZC:
-– Need buffer allocator support in driver + a new NDO for TX
-All the packet data buffers used in TX/RX ring are allocated from a specific
-memory area called UMEM which consists of a number of fixed size chunks.
-A descriptor in RX/TX ring points to the element in UMEM by its address.
-The address is simply an offset within the entire UMEM memory region.
-The UMEM also has two rings: the FILL ring and the COMPLETION ring.
-The Fill ring is used by the application to send down addr for the kernel
-to fill in with RX packet data. References to these frames will then
-appear in the RX ring once each packet has been received. The 
-completion ring, on the other hand, contains frame addr that the 
-kernel has transmitted completely and can now be used again by user
-space, for either TX or RX.
-
-Thus, the frame addrs appearing in the 
-completion ring are addrs that were previously transmitted using the 
-TX ring. In summary, the RX and FILL rings are used for the RX path
-and the TX and COMPLETION rings are used for the TX path.
-
+packet buffer with contents filled to the TX ring then issus send() system call.
+It is up to the userspace program to make sure whether the packets have been
+sent or not, by checking the COMPLETION ring structure.
+In summary, users of XSK needs to properly interact with the following four
+rings:
+\begin{itemize}
+\item FILL ring: for users to fill UMEM addresses to kernel for receiving packets.
+\item RX ring: for users to access recevied packets. 
+\item TX ring: for users to place packets needed to be sent.
+\item COMPLETION ring: for users to make sure packets are sent.
+\end{itemize}
+Unlike AF\_PACKET which is bound to entire netdev, the XSK is more fine-grained.
+XSK is bound to a specific queue id on a device, so only the traffic/flow 
+sent to the queue id shows up in the XSK.
 \section{OVS eBPF Datapath}
@@
 
-\subsection{Datapath and Netdev}
+\subsection{Datapath and Netdev Interface}
 OVS includes a userspace datapath interface (dpif) implemenation, called dpif-netdev.
@@
 
-Figure~\ref{ovsafxdp} shows the architecture of Userspace OVS with AF\_XDP.
-We implement a new netdev type called netdev-afxdp, which receives and transmits
-packets using the XSK.
-In order for XSK to work, we insert a XDP program and a eBPF map with   
-interact with XDP program to forward packets to the AF\_XDP socket.
+Figure~\ref{ovsafxdp} shows the architecture of userspace OVS with AF\_XDP.
+We implement a new netdev type for AF\_XDP, which receives and transmits
+packets using the XSK.  We insert a XDP program and a eBPF map with interact
+with XDP program to forward packets to the AF\_XDP socket.
 Once the AF\_XDP netdev receives a packet, it passes the packet to the
@@
 \subsection{AF\_XDP netdev configuration}
-ovs-vswitchd controls and manages the XDP program and maps.
-When users attach a AF\_XDP netdev to the OVS bridge, the netdev-afxdp
-does the following steps:
-We implemnet most of the logic in lib/netdev-linux.c
+When users attach a AF\_XDP netdev to the OVS bridge, for example, by
+issuing the following command:
 {
@@
 }
-
+ovs-vswitch does the following steps to configure the
+AF\_XDP netdev:
 \begin{enumerate}
-\item Attach a XDP program to the netdev queue:
+\item Attach a XDP program to the netdev's queue:
 The XDP program OVS attaches is fixed and consists
@@
 \item Create a AF\_XDP socket:
-This requires creating the XSK socket by calling socket() syscall,
-setting up its RX and TX ring buffer, allocating a UMEM region,
-and setting up FILL/COMPLETION ring of the UMEM.
+Call socket() syscall to create a XSK,
+set up its RX and TX ring buffer, allocate a UMEM region,
+and set up FILL/COMPLETION ring of the UMEM.
 
@@
 The XSK eBPF maps consists of key value pairs, where key is an u32 index and value is
-the netdev's queue ID.  In our case, we simply program one entry to the map with
+the file descript of the XSK.  ovs-vswitchd simply programs one entry to the map with
 key equals zero and the file descriptor, fd, of the XSK as its value.
@@
 
-\item Initialize the FILL queue
-Before any packet arrival, the rest three queues: RX, TX and COMPLETION
-remains the same
+\item Fill the UMEM FILL ring: Get a couple UMEM elements and place into FILL ring.
 \end{enumerate}
-
-When a netdev-afxdp is detached or closed by user,
-ovs-vswitchd closes the XSK socket, free its memory, and unload the eBPF program and map.
-
-memory allocation:
+On the other hand, when a AF\_XDP netdev is detached or closed by user,
+ovs-vswitchd closes the XSK socket, free the UMEM memory region, and
+unload the eBPF program and map.
 
 % 1 UMEM per XSK per queue, no sharing
-For each netdev, ovs-vswitchd attach XSK to each 
-For each XSK
-memorypool
-We implement a UMEM freelist to keep track of the used/unused element
-
-To handle multiple 
-\subsection{Packet Reception and Transmission}
+%For each netdev, ovs-vswitchd attach XSK to each 
+%For each XSK
+\subsection{Packet Reception}
+In order to properly set-up the RX/TX/FILL/COMPLETION rings, we implement a
+UMEM memory management API. Basically, it's a list maintaining the unused elements
+in UMEM, called UMEM freelist, and APIs to get/put elements into the list.
 \begin{figure}
@@
 \vspace{-1.0em}
-\caption{An example of how OVS programs the FILL queue and RX ring when
+\caption{An example of how OVS programs the FILL ring and RX ring when
 processing incoming AF\_XDP packets.}
@@
 \end{figure}
-Figure~\ref{afxdprx} shows how ovs-vswitchd sets up the FILL queue and
+
+Figure~\ref{afxdprx} shows how ovs-vswitchd sets up the FILL ring and
 the RX ring for receving packets from XSK.  For simplicity to demonstrate
@@
 size equals 2KB.  Initially at step 1, ovs-vswitchd fills four empty UMEM elem into
-the FILL queue and waits for incoming packets.  When there are incoming packets,
-at step 2, the FILL queue's four buffer elems have been consumed and moved
+the FILL ring and waits for incoming packets.  When there are incoming packets,
+at step 2, the FILL ring's four buffer elems have been consumed and moved
 to the RX ring.  In order to keep receiving packets, ovs-vswitchd gets another
-four empty UMEM elems from thte UMEM freelist, and fills into FILL queue (step 3).
+four empty UMEM elems from thte UMEM freelist, and fills into FILL ring (step 3).
 Then ovs-vswitchd creates the metadata needed for the four packet
@@
 mechanism is triggered to place these buffer back to UMEM freelist (step 5).
-Step 5 makes sure that there are always available elems in FILL queue, so
+Step 5 makes sure that there are always available elems in FILL ring, so
 that the underlying XDP program in kernel can keep placing packets into
-the FILL queue while the userspace ovs-vswitchd is processing the previous
+the FILL ring while the userspace ovs-vswitchd is processing the previous
 received packets.
@@
 
+\subsection{Packet Transmission}
 \begin{figure}
@@
  from eth1 and plan to send to eth2.
-(1) Get 4 free element from eth2 UMEM, copy packet data to it
+(1) Get 4 free elements from eth2's UMEM, copy packet data to it
 (2) Create TX descriptors and place into TX ring
@@
 \vspace{-1.0em}
-\caption{An example of how OVS programs the COMPLETION queue and TX ring when
+\caption{An example of how OVS programs the COMPLETION ring and TX ring when
 processing incoming AF\_XDP packets from one netdev and sending to another netdev.}
@@
 packets received from eth1 to eth2.
-Assuming that both eth1 and eth2 are configured with AF\_XDP support.
+And both eth1 and eth2 are configured with AF\_XDP support.
 
@@
 As the send in XSK is asynchronos and send syscall only return return zero when
-no error, ovs-vswitch has to poll the COMPLETION queue to make sure these four
+no error, ovs-vswitch polls the COMPLETION ring to make sure these four
 packets have been sent out (step 4).
@@
 In addition, the original four packet buffer from eth1, {1, 2, 3, 4}, is
-recycled back to eth1's UMEM freelist.
+also srecycled back to eth1's UMEM freelist.
 
-Define:
-baseline rxdrop vs OVS-AFXDP rxdrop
-baseline l2fwd vs OVS-AFXDO l2fwd
 \subsection{Performance Evaluation}
+All of our performance results use a hardware test bed that consists of
+two Intel Xeon E5 2440 v2 1.9GHz servers, each with 1 CPU socket and
+8 physical cores with hyperthreading enabled.
+One server, the target server, has an Intel 40GbE XL710 single port NIC,
+the other server, the source server, has Netronome NFP-4000 40GbE device.
+The target server runs the Linux kernel 4.19-rc4, with i40e driver supporting
+the AF\_XDP mode.
+We installed OVS-AFXDP on the target server, and the other server,
+the source server, generates 64-byte single UDP flow packets
+at the rate of 19~Mpps using the DPDK library.
 
-\subsubsection{Testbed}
+For all our experiments, we fuse a microbenchmark program, called xdpsock,
+as the baseline to compare with OVS-AFXDP implementation.
+xdpsock is an AF\_XDP sample program doing no packet processing, and can be configured
+to do simply dropping the packet, rxdrop, or forwarding the packet to the same port as
+it receives, l2fwd.
+In our testbed, we measured $19Mpps$ for xdpsock-rxdrop and $17Mpps$ for xdpsock-l2fwd.
+For OVS-AFXDP, we conduct similar two experiments, but with OVS's OpenFlow rule below:
+\begin{itemize}
+\item OVS-AFXDP rxdrop: Install a single OpenFlow rule to drop every packets,
+e.g.,  \texttt{in\_port=eth1, actions=drop}.
+\item OVS-AFXDP l2fwd: Install a single OpenFlow rule to forward packet to the same port
+as it receives, e.g., \texttt{in\_port=eth1, actions=set\_fields:eth2->in\_port, output:eth1}.
+\end{itemize}
+
+Additionally, AF\_XDP provides three operation modes:
+1) XDP\_SKB: works on devices using generic XDP~\cite{genericxdp},
+2) XDP\_DRV: works on devices with XDP support.
+3) XDP\_DRV ZC (Zero Copy): works on devices with XDP\_XDP zerecopy support.
+For all our experiments, we use the XDP\_DRV ZC mode.
 
@@
 \end{itemize}
-
-For all our experiments, we use a microbenchmark program, called xdpsock,
-as the baseline to compare with our implementation.
-xdpsock is an AF\_XDP sample program doing no packet processing, and can be configured
-to do simply dropping the packet, rxdrop, or forwarding the packet to the same port as
-it receives, l2fwd.  In our testbed, we measured $19Mpps$ for rxdrop and $17Mpps$ for
-l2fwd.
 

### 2018-10-22 paper.tex
add OVS-eBPF section

@@
 \section{Introduction}\label{introduction}
+
+eBPF, extended Berkeley Packet Filter, enables userspace applications to
+customize and extend the Linux kernel's functionality.  It provides
+flexible platform abstractions for network functions, and is being ported
+to a variety of platforms. 
+The Linux kernel BPF~\cite{cbpf,bpfp},
+with its extension called eBPF~\cite{lwnebpf,daniel1,daniel2},
+provides the desired infrastructure for an OVS datapath implementation.
+Specifically, a packet flowing inside a Linux
+box can be intercepted and trigger the execution of arbitrary eBPF programs.
+
 Among the various ways of using eBPF, OVS has been exploring the power of eBPF in three:
@@
 sent to the queue id shows up in the XSK.
+
 \section{OVS eBPF Datapath}
-\mycomment{
-\begin{figure}
-{\scriptsize
-\begin{verbatim}
+\myfigcross{ovsebpf}{5.5in}{The overall architecture of OVS eBPF datapath
+consists of multiple eBPF programs which are tail-called dynamically, maps
+which are shared between eBPF programs and userspace applications, and
+ovs-vswitchd as the management plane for all components.}
 
-               _
-              |   +-------------------+
-              |   |    ovs-vswitchd   |
-              |   +-------------------+
-    userspace |   |      ofproto      |<-->OpenFlow controllers
-              |   +--------+-+--------+  
-              |   | netdev | |ofproto-|
-              |   +--------+ |  dpif  |
-                  | netdev | +--------+
-   *eBPF hook --> |provider| |  dpif  |
-                  +---||---+ +--------+
-              |       ||     |  dpif  | <--- *eBPF provider
-              |       ||     |provider|
-              |_      ||     +---||---+
-                      ||         ||
-               _  +---||-----+---||---+
-              |   |          |datapath| <--- *eBPF datapath
-       kernel |   |          +--------+
-              |   |                   |
-              |_  +--------||---------+
-                           ||
-                        physical
-                           NIC
-\end{verbatim}
+\subsection{eBPF Configuration and Maps}
+The OVS eBPF datapath consists of one ELF-formatted object file which provides
+the full functionality of an OVS datapath. This object defines
+a set of maps and a set of eBPF programs which implement a subset of
+the datapath functionality.  To bootstrap, we load the eBPF program into
+the kernel using \verb+iproute+. One of the programs is marked
+within the ELF file to indicate that it should be
+attached to the hook point; the other programs are only executed via tail
+calls rooted in the eBPF hook point.  The ELF file also defines multiple persistent eBPF maps, which are
+pinned to the BPF filesystem ~\cite{bpfpmap}
+for sharing between different eBPF programs and \verb+ovs-vswitchd+.
+The OVS datapath requires the following eBPF maps:
+\begin{itemize}
+\item Flow key. This is the internal representation of the protocol headers
+    and metadata for the current packet being processed, defined in the P4~\cite{p4} language.
+\item Flow Table. This is a hash table whose key is the 32-bit hash value
+of the flow key, from both packet and its metadata,
+and value equals an array of actions to execute upon the flow.
+\item Stats Table. This is similar to the flow table, but rather than holding
+an array of actions to execute for the packet, it contains packet and byte
+statistics for the flow.
+\item Perf Ring Buffer~\cite{ebpfperf,ebpfperf2}. The perf event map
+allows an eBPF program to put user-defined data into a ring buffer which
+may be read from a userspace program. \verb+ovs-vswitchd+ memory maps
+this ring buffer to read packets and metadata from the eBPF program for
+flow miss upcall processing.
+\item Program Array. This map allows eBPF programs to tail call other eBPF
+programs. When the BPF loader inserts eBPF programs into the kernel, it assigns
+unique indexes for each program and stores these into the map.  At run time,
+an eBPF program will tail call another program by referring to an index within
+this map.
+\end{itemize}
+
+% locking and synchronization
+The OVS eBPF program is triggered by the TC ingress hook associated with a
+particular network device. Multiple instances of the same eBPF program
+may be triggered simultaneously on different cores which are receiving traffic
+from different network devices.  The eBPF maps, unless specified, have a single
+instance across all cores.  Access to map entries are protected by the kernel
+RCU~\cite{rcu} mechanism which makes it safe to read concurrently.  However,
+there are no built-in mechanisms to protect writers to the maps.
+For the flow table map, OVS avoids the race by ensuring that
+only \verb+ovs-vswitchd+ inserts or removes elements from the map from a single
+thread.  For flow statistics, atomic operations are used to avoid race
+conditions.  Other maps such as the flow key perf ring buffer maps use
+per-cpu instances to manage synchronization.
+
+\subsection{Header Parsing and Lookup}
+\label{parsing}
+When a packet arrives on the TC ingress hook, the OVS eBPF datapath begins
+executing a series of programs, beginning with the parser/lookup program
+as shown in Figure~\ref{ovsebpf:fig}.  The eBPF parser program
+consists of two components; standard protocol parsing and Linux-specific
+metadata parsing.  The protocol parsing is executed directly on the packet
+bytes based on standardized protocols, while the platform-specific metadata
+parsing must occur on the {\em context} provided by the eBPF environment.
+
+The resulting code will assemble the protocol headers and metadata,
+collectively known as the flow key. This flow key is then used to look for an
+entry in the flow table map, to get an array of actions to execute. If there is
+no entry in the flow table map, then the packet and the flow key will be
+written to the perf event map for further processing by \verb+ovs-vswitchd+.
+
+\subsection{Action Execution}
+\label{execution}
+When a lookup is successful the eBPF gets a list of actions to be executed,
+such as outputting the packet to a certain port, or pushing a VLAN tag. The
+list of actions is configured in \verb+ovs-vswitchd+ and may be a variable
+length depending on the desired network processing behaviour. For example, an
+L2 switch doing unknown broadcast sends packet to all its current ports.  The
+OVS datapath's actions is derived from the OpenFlow action specification and
+the OVSDB schema for \verb+ovs-vswitchd+.
+
+One might expect to intuitively write an eBPF program to iterate across the
+list of actions to execute, with each iteration of the loop dealing with a
+particular action in the list.  Unfortunately, this type of iteration implies
+dynamic loops, which are restricted within the eBPF forwarding model. Moreover,
+the variable number of actions also implies that there is no way to guarantee
+the bounded program size, which is limited to 4,096 eBPF bytecode instructions.
+
+To solve these challenges, we first break each type of action logic into an
+independent eBPF program and tail call from one eBPF action program to
+another, as shown in Figure~\ref{ovsebpf:fig}.  This alleviates
+the problem from having 4k instructions for the {\em entire} action list to 4k
+instructions {\em per action}.  Our proof-of-concept implementation shows that
+this limitation is sufficient for all existing actions.  Additionally, the
+design allows each action to be implemented and verified independently.
+Second, to solve the dynamic looping problem, we convert the variable length
+list of actions into a fixed length 32-element array.  As a result,
+flow table lookup always returns an array of 32 actions to be executed,
+and the LLVM compiler unrolls the loop to pass the BPF verifier.
+If a matching flow has less than 32 actions to execute, the rest of the
+actions is no-op, and we short-cut the action execution to the deparser.
+If a matching flow has more than 32 actions, then the eBPF datapath delegates
+the execution to userspace datapath, i.e., the slow path. For many use cases,
+this is sufficient; there is room to further optimize this path in the future
+if common cases require more than 32 actions.
+
+Each action also requires certain action-specific metadata to execute.
+For example, an output action would require an {\em ifindex} to
+indicate the output interface, while a push\_vlan action needs the {\em VLAN ID}
+to be inserted into the VLAN tag.  To accommodate this, the array element not
+only contains the action to execute, but also the metadata required by the
+particular action. The size of each element must be big enough to hold the
+metadata for any action, as it is a fixed sized array. Future work may
+relax this requirement.
+
+Each eBPF program executes in an isolated environment. As a consequence
+of breaking the action list into individual eBPF action programs, some state needs to be
+transferred between the current eBPF program and the next.
+The state to transfer from one eBPF action program to another includes
+(1) The flow key, so that the currently executing eBPF action program can
+lookup the flow table, and (2) The action index in the actions list, so that
+the eBPF action program knows it is executing the $N$th element in the list,
+and at the end of its processing, to tail-call the $n+1$th action program.
+Currently eBPF could use either per-CPU maps as scratch buffers or the
+context's packet control block (\verb+cb+) to pass data between programs. In our
+design, we use one 32-bit value in the \verb+cb+ to keep the action
+index, and per-cpu maps to save the flow key.
+%As a result, there is additional overhead of map lookup per action.
+
+% an example to illustrate
+Figure~\ref{ovsebpf:fig} also demonstrates an example eBPF execution of a
+packet forwarded to port 2 as well as port mirroring to a VLAN port with VLAN ID
+100 at port 3.  Once the packet is parsed,
+the flow table lookup returns an action list of output:2, push\_vlan:100,
+output:3.  At the end of the parser+lookup program, it tail calls the eBPF
+output program as the first step to kick start action execution.
+The execution of this output program {\em overwrites} the caller's stack,
+so it has to look up the flow table map to retrieve the flow key,
+and execute the output action.  Once done, the output action increments the
+action index from 0 to 1, saves it in \verb+cb+, and tail-calls the next action
+program, which is push\_vlan.  The push\_vlan eBPF program again 
+looks up the flow table, and fetches the action metadata at index 1
+and executes the push VLAN action using a BPF helper function.
+The third output action follows the same procedure and finally sends the
+packet out to port 3.
+\saveforlater{Appendix A shows the eBPF helper functions used in
+each eBPF action program implementation.}
+
+\subsection{Flow Miss Upcall and Installation}
+One of the important tasks of the OVS datapath is to pass any packet that
+misses its flow table to the slow path to get instructions for
+further processing.  In the existing OVS {\em kernel} datapath implementation,
+the missed packet is sent to \verb+ovs-vswitchd+, which processes the packet,
+inserts a flow into the flow table, and re-injects the packet into the kernel
+datapath using the Linux generic netlink interface.  For the eBPF datapath, this
+design implies two requirements: (1) a way for the eBPF program to communicate
+with userspace, and (2) a mechanism for the userspace to re-insert packet into
+the eBPF program.
+
+% upcall: how ebpf work 
+To address the first requirement, we use the support for Linux perf ring
+buffers and the \verb+skb_event_output()+ eBPF helper function~\cite{ebpfperf}
+which allows the eBPF program to pass data to the userspace through the Linux perf
+ring buffer~\cite{ebpfperf}.  During miss upcall processing, the eBPF program
+will insert the full packet and the current flow key into the perf ring buffer.
+To receive the data from userspace, \verb+ovs-vswitchd+ runs a thread which
+maps the ring buffer to its virtual memory using mmap system call, and polls
+the buffer using the poll system call. If there is incoming data from the ring buffer, the
+thread is woken up, it reads from the buffer and processes the packet and
+metadata. The result of this processing will be inserted into the flow table
+map.
+% down call: how to re-insert packet
+To address the second requirement, we construct a dedicated Linux TAP device
+which also has the OVS eBPF datapath program attached to it.
+\verb+ovs-vswitchd+ sends the missed packet using an AF\_PACKET socket,
+triggering the underlying eBPF program for further processing.
+This program is very similar to the
+previously-used parser+lookup program, with minor changes. Specifically, this
+packet was originally received on one device, however when \verb+ovs-vswitchd+
+sends the packet on the TAP device, the eBPF program is triggered for the TAP
+device instead. So, the platform metadata for incoming port misidentifies the
+source as the TAP device. To ensure that the packet lookup occurs correctly,
+\verb+ovs-vswitchd+ prepends the port number to the beginning of the packet
+data, then when the eBPF program for the dedicated TAP device processes the
+packet, it will read this port number into the metadata, then strip this port
+from the packet. The resulting packet is identical to the originally-received
+packet, and now the metadata will match the metadata originally generated by
+the parser+lookup the first time the packet was received. The rest of the
+lookup, actions execution, and deparser is then executed as per the description
+in the previous sections.
+
+\subsection{Evaluation}\label{evaluation}
+To quantify the performance of the OVS eBPF datapath, we measure the packet
+forwarding rate in millions of packets per second (Mpps), using 64-byte
+packets under different forwarding scenarios. The hardware testbed
+consists of two Intel Xeon E5 2650 servers, each with an Intel 10GbE
+X540-AT2 dual port NIC, with the two ports of the Intel NIC on one server
+connected to the two ports on the identical NIC on the other server. The
+OVS eBPF datapath is installed on one server, acting as a bridge to
+forward packets from one port on the NIC to the other, and vice-versa.
+The other server acts as packet generating host, which runs DPDK
+packet-gen sending at the maximum packet rate of
+$14.88$Mpps. This server sends to one port to the target server, and
+receives the forwarded packets on the other port. All experiments use
+only one CPU core running Linux kernel 4.9-rc6.
+\begin{table}
+\centering
+\small
+\begin{tabular}{|c | c | c| c|}
+\hline
+ {\bf Linux Bridge} & {\bf OVS Kernel } & {\bf Linux TC} & {\bf OVS eBPF}\\ \hline
+ 1.6 Mpps & 1.4 Mpps & 1.9 Mpps & 1.12 Mpps\\ \hline
+\end{tabular}
+\caption{\footnotesize
+Baseline port-to-port forwarding rate using existing Linux subsystems and
+OVS kernel and eBPF.
 }
-\vspace{-1.0em}
-\caption{An example of mirrored packet with outer header containing
-the GRE and ERSPAN header, followed by the innter Ethernet frame.}
-\label{erspanhdr}
-\vspace{-1.0em}
-\end{figure}
-\subsection{Tunnel Support}
-\subsection{Megaflow Support}
-\subsection{Connection Tracking Support}
-}
+\label{table:baseperf}
+\end{table}
+
+\textbf{Baseline Forwarding Performance.}
+We start by conducting a simple port-to-port packet forwarding experiment,
+i.e., receiving packets from one port and outputting to the other,
+using the Linux native bridge, Linux TC, and OVS kernel datapath, as shown in
+Table~\ref{table:baseperf}.
+The native Linux bridge is a simple L2 mac-learning switch with no
+programmability, showing the forwarding rate of $1.6$ Mpps.
+The OVS kernel module, which does additional flow key extraction,
+shows a slower forwarding rate of $1.4$ Mpps.
+For forwarding packets using TC, we have an eBPF program loaded
+into TC, with the program only having one BPF helper function call, the
+bpf\_skb\_redirect, that redirects the incoming packet from one interface
+to another.  Since TC accesses the incoming packets closest to the driver
+layer, it shows the highest performance of $1.9$ Mpps.
+Finally, we measure our proposed OVS eBPF forwarding speed,
+with incoming packets traversing through an eBPF parser+lookup program,
+an output action program, and the deparser program.
+Since the OVS eBPF is implemented based on tail calling these
+additional eBPF programs at TC, we observe the overhead of $0.78$ Mpps,
+a reduction from $1.9$ Mpps to $1.12$ Mpps.
+
+\textbf{Forwarding with action execution.}
+To further investigate the overhead,
+we program one additional action executed before the packet is
+forwarded to the other port.  Since the Linux bridge has no
+programmability, we only compare the OVS kernel datapath with the OVS eBPF
+datapath.  Table~\ref{table:ebpfperf} shows forwarding
+packets while executing the additional action type: hash, push\_vlan,
+set\_dst\_mac, and set\_gre\_tunnel, with the OVS kernel and eBPF datapaths.
+The forwarding performance of these results is bounded by the CPU cost.
+In the case of the hash and vlan actions, the NIC allows the processing
+to be offloaded from the CPU, resulting in minimal performance overhead for
+executing these actions.  As a result, the forwarding rate exhibits little
+to no overhead above the baseline.
+Moreover, for the hash and push\_vlan actions, due to the NIC offloading of
+actions processing, the datapath does not need to execute any modification to
+the packet contents, so the deparser program execution may be optimized out.
+
+To ensure invoking the deparser, the third experiment, set\_dst\_mac, involves
+altering packet data, specifically to modify the destination MAC address. In
+this case, the deparser writes stored metadata from actions processing back to
+the packet. The performance in this experiment drops to $0.84$ Mpps. The drop
+of $0.28$ Mpps from baseline is the result of the deparser writing all known
+packet header metadata back to the packet, and due to extraneous memory copies
+introduced by the P4-to-eBPF compiler. With a more intelligent deparser
+implementation, and by using recent improvements to the eBPF
+API~\cite{dpa1,dpa2}, the performance gap compared with the native kernel
+implementation is expected to shrink.
+
+%The Set DST\_MAC result of $0.84$ Mpps makes the packet modification using the P4-generated deparser.
+Finally, the set\_get\_tunnel action experiment yielded $0.48$ Mpps, which
+represents the additional cost of tunnelling traffic; such traffic must
+traverse the Linux network stack twice, once for overlay traffic and
+once for underlay traffic. The eBPF result is also more expensive than
+the existing Linux implementation, but with a narrower performance gap
+than the earlier experiments. The majority of processing in this case occurs
+outside of OVS, so the overhead of the eBPF datapath has less effect.
+
+\begin{table}
+\centering
+\small
+\begin{tabular}{|c | p{1.4cm} | p{1.4cm}| c|}
+\hline
+{\bf Action} & {\bf eBPF DP} & {\bf Kernel DP} & {\bf Overhead}\\ \hline 
+hash & 1.12 & 1.34 & 16\%\\ \hline
+push\_vlan & 1.11 & 1.32 & 15\% \\ \hline
+set\_dst\_mac & 0.84 &  1.28 & 34\% \\ \hline
+%set\_dst\_mac (no parser) & 1.14 & -- & --\\ \hline
+set\_gre\_tunnel & 0.48 & 0.57 & 8\%\\ \hline 
+\end{tabular}
+\caption{\footnotesize Comparison of single core forwarding rate in Mpps
+with eBPF and Kernel Datapath, with the additional
+action executed before forwarding the packet to another port.}
+\label{table:ebpfperf}
+\end{table}
 

### 2018-10-22 paper.tex
revise introducation

@@
 OVS-eBPF has the goal of re-writing existing flow processing features
-in openvswitch kernel module into eBPF code.
+in openvswitch kernel module into eBPF code, by attaching eBPF program
+to Linux TC.
 On the other hand, the OVS-AFXDP project aims to by-passing the kernel
-using AF\_XDP socket and moving most of the flow processing features
-in userspace.
+using AF\_XDP socket and moves most of the flow processing features
+into userspace.
 
@@
 flexible platform abstractions for network functions, and is being ported
-to a variety of platforms. 
-The Linux kernel BPF~\cite{cbpf,bpfp},
-with its extension called eBPF~\cite{lwnebpf,daniel1,daniel2},
-provides the desired infrastructure for an OVS datapath implementation.
-Specifically, a packet flowing inside a Linux
-box can be intercepted and trigger the execution of arbitrary eBPF programs.
-
-Among the various ways of using eBPF, OVS has been exploring the power of eBPF in three:
-(1) attaching eBPF to TC, (2) offloading a subset of processing to XDP, and (3) by-passing
-the kernel using AF\_XDP.
-Attaching eBPF to TC started first with the most aggressive goal: we planned to re-implement
-the entire features of OVS kernel datapath under net/openvswitch/* into eBPF code.
-We worked around a couple of limitations, for example, the lack of TLV support led us to
-redefine a binary kernel-user API using a fixed-length array; and without a dedicated way to
-execute a packet, we created a dedicated device for user to kernel packet transmission,
+to a variety of platforms.  Among the various different ways of using eBPF,
+OVS has been exploring the power of eBPF in three:
+1) in-kernel flow processing by attaching eBPF programs to TC,
+2) offloading a subset of processing to XDP (eXpress Data Path),
+and 3) moving the flow processing to userspace by using AF\_XDP.
+
+Attaching eBPF to TC, OVS-eBPF project, started first with the most aggressive goal:
+we planned to re-implement the entire features of OVS kernel datapath under net/openvswitch/* into eBPF code.
+We worked around a couple of limitations, for example, the lack of TLV (Type-Length-Value) support
+led us to redefine a binary kernel-user API using a fixed-length array;
+and without a dedicated way to execute a packet, we created a dedicated
+device for user to kernel packet transmission,
 with a different BPF program attached to handle packet execute logic.
-Currently, we are working on connection tracking.
-Although a simple eBPF map can achieve basic operations of conntrack table lookup and commit,
-how to handle NAT, (de)fragmentation, and ALG are still under discussion.
+Currently, OVS-eBPF can satisfy most of the basic feature requirement
+for flow processing, including tunneling protocol support.
+But more complicated processing such as connection tracking, NAT,
+(de)fragmentation, and ALG are still under discussion.
 
@@
 1) only very limited number of flows are eligible for offload, or 2) more flow processing logic
-needed to be done in native eBPF.
-
-AF\_XDP is another form of XDP but providing a socket interface for control plane and a shared
-memory API for accessing packets from userspace application. OVS today has another full-fledged
-datapath implementation in userspace, called dpif-netdev, used by DPDK community.
-By treating the AF\_XDP as a fast packet-I/O channel, the OVS dpif-netdev can satisfy most of the
-features. We are working on building the prototype and evaluating its performance.
-
-Existing OVS consists of three datapath implementations:
-dpif-netlink
-dpif-netdev
-dpif-windows
-In this paper, OVS-eBPF project adds a new datapath type
-dpif-bpf
-and OVS-AFXDP project reuses the dpif-netdev interface, with
-linux netdev receiving/sending packets from AFXDP sockets.
-
-Table~\ref{compare} shows the difference between two projects:
+needed to be done in native eBPF code.
+
+XDP provides another way for interacting with userspace programs, called AF\_XDP. 
+AF\_XDP is another socket interface for control plane and a shared
+memory API for accessing packets from userspace application.
+OVS today has another full-fledged
+datapath implementation in userspace, called dpif-netdev, mostly used by DPDK community.
+OVS-AFXDP project treats the AF\_XDP as a fast packet-I/O channel, and move
+most of the flow processing into the userspace OVS.
+
+A rough comparison of the two project is shown in Table~\ref{compare}
+This paper describes the design, implementation, and evaluation of the
+OVS-eBPF and OVS-AFXDP projects.
 \begin{table}
@@
  {\bf Projects} & {\bf OVS-eBPF } & {\bf OVS-AFXDP} \\ \hline\hline
-%  Performance & $<$2Mpps & $>14$Mpps \\ \hline
-  Performance & Low & High \\ \hline
+  Packet Rate & Low & High \\ \hline
   Driver Support & No & Yes \\ \hline
-  Minimal Kernel & 4.18 & Unknown \\ \hline
+  Minimal Kernel Req. & 4.18 & Unknown \\ \hline
   BPF Code Size & Large & Minimal \\ \hline
@@
 \section{Background}
-
-\subsection{OVS Forwarding Model}
-
-\begin{figure}
-{\scriptsize
-\begin{verbatim}
-
-                +-------------------+
- slow path      |    ovs-vswitchd   |
-                +-----^-------------+
-                      |       |flow installation
-----------------------|-------|-----------------
- fast path     upcall |       | 
- (datapath)  +----------------v---------+ 
-  packets--> | parse, lookup, actions   |  --> output
-             +--------------------------+
-\end{verbatim}
-}
-\vspace{-1.0em}
-\caption{An example of mirrored packet with outer header containing
-the GRE and ERSPAN header, followed by the innter Ethernet frame.}
-\label{erspanhdr}
-\vspace{-1.0em}
-\end{figure}
 
@@
 
-\mycomment{
-As each new feature in OVS datapath requires deep knowledge and programming
-paradigm of the target operating system, e.g., how Linux packets are received,
-how Windows kernel module does firewalling, and each operating system
-community has its own software upstream and maintenance processes, the software
-development cycle of a new datapath feature becomes complex and requires
-detailed platform-specific knowledge. Each platform introduces its own
-development and maintenance burden, and it is common for new features to take
-six months or more before implementations are available across all datapaths.
-As a result, providing a programmable OVS datapath which allows extending the
-features flexibly with lower maintenance cost becomes a critical requirement.
+\subsection{OVS Forwarding Model}
+
+\begin{figure}
+{\scriptsize
+\begin{verbatim}
+
+                +-------------------+
+ slow path      |    ovs-vswitchd   |
+                +-----^-------------+
+                      |       |flow installation
+----------------------|-------|-----------------
+ fast path     upcall |       | 
+ (datapath)  +----------------v---------+ 
+  packets--> | parse, lookup, actions   |  --> output
+             +--------------------------+
+\end{verbatim}
 }
+\vspace{-1.0em}
+\caption{An example of mirrored packet with outer header containing
+the GRE and ERSPAN header, followed by the innter Ethernet frame.}
+\label{erspanhdr}
+\vspace{-1.0em}
+\end{figure}
 
@@
 the freelist.
-producer 
-consumer
+producer, consumer
 We implemented three data structures of UMEM memory pool as below
@@
 
-%\subsubsection{Metadat Allocation}
+\subsubsection{Metadat Allocation}
 With the above change, Linux perf shows that the packets metadata allocation
@@
 \begin{itemize}
-\item Inlining in UMEM buffer: Since we already allocate 2K chunk for each
+\item Embedding in UMEM buffer: Since we already allocate 2K chunk for each
 UMEM buffer, we reserve the first 256-byte in each UMEM element as struct dp\_packet
 and initialize the dp\_packet as much as we can at allocation time.
-So when packets arrive, each packet already has its packet metadata at a const
-offset ahead.
+This is similar to the DPDK mbuf design~\cite{dpdkmbuf}, where a single memory buffer
+is used for both packet data and metadata.
 \item Array on heap: This design simply allocate a continuous memory region
@@
 % see mbuf dpdk https://doc.dpdk.org/guides/prog_guide/mbuf_lib.html
-%Embed metadata within a single memory buffer the structure followed by a fixed size area for the packet data.
+%Embed metadata within a single memory buffer the structure followed by a fixed size area for
+% the packet data.
 %Use separate memory buffers for the metadata structure and for the packet data.

### 2018-10-22 paper.tex
more revise

@@
 Windows Hyper-V, Solaris and Linux. As shown in
-Figure~\ref{ovsintro:fig}, the architecture of OVS consists of two major
+Figure~\ref{ovsintro}, the architecture of OVS consists of two major
 components: a slow path and a fast path. OVS begins processing packets
@@
  fast path     upcall |       | 
- (datapath)  +----------------v---------+ 
-  packets--> | parse, lookup, actions   |  --> output
-             +--------------------------+
+ (datapath)  +----------------v---------------+ 
+  packets--> | parse, lookup match, actions   |  --> output
+             |        & action table          |
+             +--------------------------------+
 \end{verbatim}
@@
 \vspace{-1.0em}
-\caption{An example of mirrored packet with outer header containing
-the GRE and ERSPAN header, followed by the innter Ethernet frame.}
-\label{erspanhdr}
+\caption{
+The forwarding plane of OVS consists of two components;
+ovs-vswitch handles the complexity of the OpenFlow protocol, and the
+datapath acts as a caching layer to optimize the performance.  A flow
+missed by the match/action table in the datapath triggers an {\em
+upcall}, which forwards the information to ovs-vswitchd.  ovs-vswitchd
+installs an appropriate flow entry into the datapath's match/action table.}
+\label{ovsintro}
 \vspace{-1.0em}
@@
 \subsection{AF\_XDP netdev configuration}
-When users attach a AF\_XDP netdev to the OVS bridge, for example, by
-issuing the following command:
-{
+When users attach a AF\_XDP netdev to an OVS bridge, for example, by
+issuing the following commands:
+{\small
 \begin{verbatim}
@@
 }
-ovs-vswitch does the following steps to configure the
+ovs-vswitch does the following steps to bring up the
 AF\_XDP netdev:
@@
 
-\item Fill the UMEM FILL ring: Get a couple UMEM elements and place into FILL ring.
+\item Populate the UMEM FILL ring: Get a couple UMEM elements and place into FILL ring.
 \end{enumerate}
@@
 
-% 1 UMEM per XSK per queue, no sharing
-%For each netdev, ovs-vswitchd attach XSK to each 
-%For each XSK
-\subsection{Packet Reception}
 In order to properly set-up the RX/TX/FILL/COMPLETION rings, we implement a
 UMEM memory management API. Basically, it's a list maintaining the unused elements
-in UMEM, called UMEM freelist, and APIs to get/put elements into the list.
+in UMEM, called UMEM freelist, and function calls to get/put elements into the list.
+\subsection{Packet Reception}
 \begin{figure}
@@
 One server, the target server, has an Intel 40GbE XL710 single port NIC,
-the other server, the source server, has Netronome NFP-4000 40GbE device.
+and the other server, the source server, has Netronome NFP-4000 40GbE device.
 The target server runs the Linux kernel 4.19-rc4, with i40e driver supporting
@@
 at the rate of 19~Mpps using the DPDK library.
+On the target server, we also disable the spectre and meltdown fixes.
 
@@
 xdpsock is an AF\_XDP sample program doing no packet processing, and can be configured
-to do simply dropping the packet, rxdrop, or forwarding the packet to the same port as
+to do simply dropping all packets, rxdrop, or forwarding the packet to the same port as
 it receives, l2fwd.
 In our testbed, we measured $19Mpps$ for xdpsock-rxdrop and $17Mpps$ for xdpsock-l2fwd.
-For OVS-AFXDP, we conduct similar two experiments, but with OVS's OpenFlow rule below:
+For OVS-AFXDP, we conduct similar two experiments, but with OVS's OpenFlow rule installed as
+below:
 \begin{itemize}
@@
 Additionally, AF\_XDP provides three operation modes:
-1) XDP\_SKB: works on devices using generic XDP~\cite{genericxdp},
-2) XDP\_DRV: works on devices with XDP support.
-3) XDP\_DRV ZC (Zero Copy): works on devices with XDP\_XDP zerecopy support.
+XDP\_SKB: works on devices using generic XDP~\cite{genericxdp},
+XDP\_DRV: works on devices with XDP support, and
+XDP\_DRV ZC (Zero Copy): works on devices with XDP\_XDP zerecopy support.
 For all our experiments, we use the XDP\_DRV ZC mode.
 
-When forwarding packets using AF\_XDP with OVS, two CPUs show up
+\subsubsection{PMD netdev}
+Our initial prototype shows pretty terrible performance, only $0.5Mpps$ for rxdrop.
+We observe that when dropping/forwarding packets under OVS-AFXDP, two CPUs show up
 100\% utilitization:
@@
 \item ovs-vswitchd: This is the process keeps doing the send and receive steps
-in Figure~\ref{afxdprx,afxdptx}.
-\item ksoftirqd: This is the kernel software interrupt thread handling the
-incoming packets, triggering XDP program to pass to the XSK, and also processing
-transmission.
+in Figure~\ref{afxdprx} and~\ref{afxdptx}.
+\item ksoftirqd/core\_id: This is the kernel software interrupt thread handling the
+incoming packets, triggering XDP program to pass packets to the XSK, and also
+processing transmission.
 \end{itemize}
-
-Compared to xdpsock, ovs-vswitchd adds on top of it a UMEM memory pool,
-and packets parsing, lookup and action execution.
-Our initial prototype shows pretty terrible performance, only $0.5Mpps$ rxdrop.
-We investigated and our first fix is to enable OVS's PMD-mode,
+\mycomment{Compared to xdpsock, ovs-vswitchd adds on top of it a UMEM memory pool,
+and packets parsing, lookup and action execution.}
+We investigate using Linux perf and our first fix is to enable OVS's PMD-mode,
 Poll-Mode-Driver, to the afxdp netdev.
 
-\subsubsection{PMD netdev}
 In OVS, a non-pmd mode netdev does packet receiption by putting all the receiving
-netdev's file descript together and ovs-vswitch polls them when there is an fd ready
-to perform I/O.  As a result, the fd of the XSK is {\em shared} with other fds, and
+netdev's file descriptors, fds, together and ovs-vswitchd polls them when there is an fd
+ready to perform I/O.  As a result, the fd of the XSK is {\em shared} with other fds, and
 we observed that the poll system call has pretty high overhead.
@@
 Applying OVS's PMD netdev avoids these problems and improving the rxdrop performance
-from $0.5Mpss$ to $3Mpps$.  Ideally, receing packets from XSK does not need to context
+from $0.5Mpss$ to $3Mpps$.  Ideally, receing packets from XSK should not need to context
 switch to into kernel as mentioned in steps~\ref{afxdprx}. So an ideal implementation
-should show minimal context switches with most of the CPU stays in userspace. 
+should show minimal context switches with most of the CPU time stays in userspace. 
 When enabling PMD netdev for afxdp, ovs-vswitch can be
-configured to create a {\em dedicated} thread for the packet I/O.  
-In our case, we create a thread for each XSK receive queue which keeps polling
+configured to create a {\em dedicated} thread for the packet processing.
+In OVS-AFXDP, we create a thread for each XSK's receive queue and the thread keeps polling
 the RX ring for new packets.  Each round of receive polling processes a batch of
-packet, up to maximum of 32 packets. In addition, we enable high memory page 1GB to
-reduce the page fault overhead.  With the above fixes/optimizations, the rxdrop
+packet, up to maximum of 32 packets. In addition, we enable 1GB huge memory page support
+to reduce the page fault overhead.  With the above fixes/optimizations, the rxdrop
 performance increases to around $10Mpps$
 
-\subsubsection{Memory Management and Locks}
+\subsubsection{Memory Management and Locks}\label{memorym}
 For every stage of the optimization, we use Linux perf extensively, e.g., perf stat
 and perf record/report.  With the above new design, we observe the new bottleneck
-is at the UMEM memory pool we introduce to maintain the UMEM freelist, and the
+is at the UMEM memory pool APIs we introduce to maintain the UMEM freelist, as well as the
 packet metadata allocation, struct dp\_packet, for every receiving packet.
@@
 UMEM freelist, and umem\_elem\_put(), which places back the free UMEM buffer to
-the freelist.
-producer, consumer
-We implemented three data structures of UMEM memory pool as below
-to compare their performance
+the freelist, similar to the concept of producer (put) and consumer (get).
+We implemented three different data structures of UMEM memory pool as below
+to compare their performance.
 \begin{itemize}
-\item list\_head: Similar to Linux kernel's list\_head, each element contains it
-UMEM addr, and a next pointer points to next free element. Elements are sparse in
-memory. Consumer always takes the first element, head, and producer also place back
+\item list\_head: Similar to Linux kernel's list\_head, each element contains its
+UMEM element address, and a next pointer points to next free element.
+Due to the nature of list\_head, elements are sparse in memory.
+Consumer always takes the first element, head, and producer also place back
 from the head of the list.
 \item ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
-allocated in continuous memory region, and pointers (head and tail) point for
-Consumer takes elements from the tail and when placing back, producer, puts
+allocated in continuous memory region.
+Consumer takes elements from the tail pointer and when placing back, producer, puts
 them into where head pointer points to.
@@
 netdev sharing the same UMEM. As a result, the above three data structures
-all require a mutex lock before accessing UMEM freelist. Linux perf reports
-pthread\_mutex\_lock as one of the top CPU utilization function, and overall
-shows around $2Mpps$ overhead for rxdrop experiments.
+require a mutex lock before accessing UMEM freelist. Linux perf reports
+pthread mutex lock related APIs as one of the top CPU utilization function.
 We change our design by 1) allocating per-queue UMEM region and 2)
@@
 
-\subsubsection{Metadat Allocation}
-With the above change, Linux perf shows that the packets metadata allocation
-takes a lot of CPU cycles, the dp\_packet\_init, dp\_packet\_set\_data.
+\subsubsection{Metadata Pre-allocation}
+Moving forward, Linux perf shows that the packets metadata allocation
+takes a lot of CPU cycles, i.e.,  dp\_packet\_init(), dp\_packet\_set\_data().
 So instead of allocating packet metadata at packet reception time, we
-implement two data structures and compare their performance:
+implemented two data structures and compare their performance:
 \begin{itemize}
-\item Embedding in UMEM buffer: Since we already allocate 2K chunk for each
-UMEM buffer, we reserve the first 256-byte in each UMEM element as struct dp\_packet
+\item Embedding in packet buffer: Since we already allocate 2K chunk for each
+UMEM packet buffer, we reserve the first 256-byte in each UMEM element as struct dp\_packet
 and initialize the dp\_packet as much as we can at allocation time.
@@
 is used for both packet data and metadata.
-\item Array on heap: This design simply allocate a continuous memory region
+\item Separate from packet buffer: This design simply allocate a continuous memory region
 storing an array of packet metadata, and initialize them as much as we can.
@@
 % the packet data.
-%Use separate memory buffers for the metadata structure and for the packet data.
-
-With the above design change, we found that using the ptr\_stack with
-array on heap get the best performance. Largely due to both data structures
-have better spatial locality and more batching friendly.
+
+With the above design change, we found that using the ptr\_stack in section~\ref{memorym} with
+separate from packet buffer design yields the best performance. Largely due to both data structures
+have better spatial locality and more batching friendly, when all data are allocated together.
 For example, setting up 32 packet metadatas in an array definitely incurs
@@
 And getting 32 free UMEM elements from ptr\_stack has similar benefits.
-With the above decision, the OVS rxdrop shows $17Mpps$, only $2Mpps$ less
-than the baseline xdpsock rxdrop.
+With the above decision, the OVS rxdrop shows around $19Mpps$, closed
+to the baseline xdpsock rxdrop performance.
 
@@
 %574.790 M/sec  
-We continued measured the performance of OVS-AFXDP l2fwd and.
-observed only $4Mpps$, compared to $17Mpps$ baseline xdpsock.
-Again, by analysis using perf, we found that the OVS pmd thread under rxdrop
+We continued measuring the performance of OVS-AFXDP l2fwd and
+observed only $4Mpps$, compared to $17Mpps$ baseline xdpsock-l2fwd.
+We found that the OVS pmd thread under rxdrop
 has much fewer context switches compared to the l2fwd, indicating that
-the PMD process spends much more time in kernel space than in userspace.
-By using strace, we saw that the l2fwd experiement calls sendto system call
+the pmd process spends much more time in kernel space than in userspace.
+By using strace, we saw that the OVS-AFXDP l2fwd experiement calls sendto system call
 at very high frequeuce, because from Figure~\ref{afxdptx}, we design to check
@@
 
-%rxhash
-
+\subsubsection{Summary}
+In summary, with the above design and optimization, OVS-AFXDP shows
+\begin{table}
+\centering
+\small
+\begin{tabular}{|c | c | c|}
+\hline
+ {\bf Projects} & {\bf xdpsock } & {\bf OVS-AFXDP} \\ \hline\hline
+  rxdrop & 19Mpps & 19Mpps \\ \hline
+  l2fwd  & 17Mpps & 14Mpps \\ \hline
+  \end{tabular}
+\caption{\footnotesize
+Performance comparison of the xdpsock and and OVS-AFXDP.
+}
+\label{compare}
+\end{table}
+Although there are still rooms for improvement, we are now working on
+making the patch upstream to the OVS code base for more people to use.
+
+\section{Conclusion}
 \bibliography{paper.bib}

### 2018-10-22 paper.tex
fix bib

@@
     \textrm{William Tu} & \textrm{Joe Stringer} & \textrm{Yifeng Sun} & \textrm{Yi-Hung Wei}\\
-    \textrm{u9012063@gmail.com} & \textrm{stringerjoe@vmware.com} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
+    \textrm{u9012063@gmail.com} & \textrm{joe@ovn.org} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
   \end{matrix}
@@
 how AF\_XDP works, assume that there are only eight UMEM buffers, with each buffer's
-size equals 2KB.  Initially at step 1, ovs-vswitchd fills four empty UMEM elem into
+size equals 2KB.  Initially at step 1, ovs-vswitchd fills four available UMEM elements into
 the FILL ring and waits for incoming packets.  When there are incoming packets,
@@
 to the RX ring.  In order to keep receiving packets, ovs-vswitchd gets another
-four empty UMEM elems from thte UMEM freelist, and fills into FILL ring (step 3).
+four available UMEM elements from the UMEM freelist, and fills into FILL ring (step 3).
 Then ovs-vswitchd creates the metadata needed for the four packet
-buffer, i.e., struct dp\_packet and struct dp\_packet\_batch, and passes to 
+buffer \{1, 2, 3, 4\}, i.e., struct dp\_packet and struct dp\_packet\_batch, and passes to 
 the dpif-netdev layer for parse, lookup and action executions.
@@
 mechanism is triggered to place these buffer back to UMEM freelist (step 5).
-Step 5 makes sure that there are always available elems in FILL ring, so
-that the underlying XDP program in kernel can keep placing packets into
-the FILL ring while the userspace ovs-vswitchd is processing the previous
-received packets.
-When step 5 finishes, ovs-vswitchd goes back to step 2, waiting and processing
+Step 5 makes sure that there are always available elements in FILL ring, so
+that the underlying XDP program in kernel can keep processing packets 
+while the userspace ovs-vswitchd is processing the previous
+received packets on RX ring.
+When step 5 finishes, ovs-vswitchd goes back to step 2, waiting for
 new packets.
@@
 \begin{verbatim}
- dpif-netdev has 4 packets at UMEM:{1, 2, 3, 4}
+ OVS dpif-netdev has 4 packets at UMEM:{1, 2, 3, 4} received
  from eth1 and plan to send to eth2.
@@
 
-Initially, assume ovs-vswitchd receives four packets {1, 2, 3, 4} from eth1's XSK
-To send to eth2 using its XSK, ovs-vswitchd first gets four packet buffers,
-{4, 5, 6, 7}, from eth2's UMEM freelist and copy the packet data to their
-packet buffer in UMEM.  Then TX descriptors for the four packets are
+Initially, assume ovs-vswitchd receives four packets \{1, 2, 3, 4\} from eth1's XSK
+To send to eth2 using eth2's XSK, ovs-vswitchd first gets four packet buffers,
+\{4, 5, 6, 7\} from eth2's UMEM freelist and copy the packet data from eth1's UMEM
+to eth2's \{4, 5, 6, 7\}.  Then TX descriptors for the four packets are
 created and placed into eth2's TX ring (step 2).
@@
 ovs-vswitchd recycles their UMEM elememnts back to the eth2's UMEM freelist.
-In addition, the original four packet buffer from eth1, {1, 2, 3, 4}, is
-also srecycled back to eth1's UMEM freelist.
+In addition, the original four packet buffer from eth1, \{1, 2, 3, 4\}, is
+also recycled back to eth1's UMEM freelist.
 
@@
 at the rate of 19~Mpps using the DPDK library.
-On the target server, we also disable the spectre and meltdown fixes.
+On the target server, we also disable the Intel spectre and meltdown fixes.
 
@@
 \subsubsection{Summary}
-In summary, with the above design and optimization, OVS-AFXDP shows
+Through the step-by-step analysis, the key to performance is to keep
+the process CPU time in userspace, and making sure the userspace process
+is dedicated to do packet processing. Moreover, standard optmization
+techniques such as batching is critical. Here we apply batching in a
+couple of places such as issuing send syscall, packet reception and
+transmission. In summary, OVS-AFXDP shows performance below:
 \begin{table}
@@
 \end{table}
+
 Although there are still rooms for improvement, we are now working on
@@
 \section{Conclusion}
+This paper describes two eBPF projects related to OVS: OVS-eBPF and OVS-AFXDP.
+
 \bibliography{paper.bib}

### 2018-10-23 paper.tex
Edit abstract

@@
 Open vSwitch, OVS, is a software switch running majorly in Linux
-operating systems and interacts with Linux kernel module,
-the openvswitch.ko.
-In this paper we describe our efforts to make use of eBPF technology in OVS.
-We present the design, implementation, and evaluation of two projects:
-the OVS-eBPF project and the OVS-AFXDP project.
-OVS-eBPF has the goal of re-writing existing flow processing features
-in openvswitch kernel module into eBPF code, by attaching eBPF program
+operating systems, its fast path packet process is implemented in Linux kernel
+module, openvswitch.ko.
+To provide greater flexibility and extensibility to OVS datapath, in this
+work, we present our design on making use of eBPF technology in OVS datapath
+development with two projects: the OVS-eBPF project and the OVS-AFXDP project.
+The goal of OVS-eBPF project is to re-write existing flow processing features
+in openvswitch kernel datapath into eBPF program, and attaching it
 to Linux TC.
@@
 into userspace.
-
+We demonstrate the feasibility of implementing OVS datapath with the
+aforementioned technologies and present the performance number in this paper.
 \end{abstract}

### 2018-10-23 paper.tex
section 4 fixes

--- 2018-10-22 paper.tex
@@
 We implement a new netdev type for AF\_XDP, which receives and transmits
-packets using the XSK.  We insert a XDP program and a eBPF map with interact
+packets using the XSK.  We insert a XDP program and a eBPF map which interacts
 with XDP program to forward packets to the AF\_XDP socket.
@@
 
-\item Populate the UMEM FILL ring: Get a couple UMEM elements and place into FILL ring.
+\item Populate the UMEM FILL ring: Get a couple of UMEM elements and place into FILL ring.
 \end{enumerate}
@@
 how AF\_XDP works, assume that there are only eight UMEM buffers, with each buffer's
-size equals 2KB.  Initially at step 1, ovs-vswitchd fills four available UMEM elements into
+size equals 2KB (?).  Initially at step 1, ovs-vswitchd fills four available UMEM elements into
 the FILL ring and waits for incoming packets.  When there are incoming packets,
@@
 Finally, when ovs-vswitchd finishes processing the UMEM buffer, a recycle
-mechanism is triggered to place these buffer back to UMEM freelist (step 5).
+mechanism is triggered to place this buffer back to UMEM freelist (step 5).
 Step 5 makes sure that there are always available elements in FILL ring, so
@@
 Figure~\ref{afxdptx} shows the process for sending packets to the XSK.
-Assuming that ovs-vswitch has a a bridge and two ports: eth1 and eth2.
+Assuming that ovs-vswitch has one bridge and two ports: eth1 and eth2.
 And there is a flow entry showing
-\texttt{in\_port=eth1, ..., actions=..., output:eth2}, meaning to forwarding
+\texttt{in\_port=eth1, ..., actions=..., output:eth2}, meaning to forward
 packets received from eth1 to eth2.
@@
 To send to eth2 using eth2's XSK, ovs-vswitchd first gets four packet buffers,
-\{4, 5, 6, 7\} from eth2's UMEM freelist and copy the packet data from eth1's UMEM
+\{4, 5, 6, 7\} from eth2's UMEM freelist and copies the packet data from eth1's UMEM
 to eth2's \{4, 5, 6, 7\}.  Then TX descriptors for the four packets are
@@
 At step 3, send syscall is issued to signal the kernel to start transmit.
-As the send in XSK is asynchronos and send syscall only return return zero when
+As the send in XSK is asynchronous and send syscall only returns zero when
 no error, ovs-vswitch polls the COMPLETION ring to make sure these four
 packets have been sent out (step 4).
-Once the four packets' descriptors shown up at the COMPLETION ring, at step 5,
-ovs-vswitchd recycles their UMEM elememnts back to the eth2's UMEM freelist.
-In addition, the original four packet buffer from eth1, \{1, 2, 3, 4\}, is
+Once the four packets' descriptors show up at the COMPLETION ring, at step 5,
+ovs-vswitchd recycles their UMEM elements back to eth2's UMEM freelist.
+In addition, the original four packet buffers from eth1, \{1, 2, 3, 4\}, are
 also recycled back to eth1's UMEM freelist.
@@
 One server, the target server, has an Intel 40GbE XL710 single port NIC,
-and the other server, the source server, has Netronome NFP-4000 40GbE device.
-The target server runs the Linux kernel 4.19-rc4, with i40e driver supporting
+and the other server, the source server, has a Netronome NFP-4000 40GbE device.
+The target server runs Linux kernel 4.19-rc4, with i40e driver supporting
 the AF\_XDP mode.
-We installed OVS-AFXDP on the target server, and the other server,
-the source server, generates 64-byte single UDP flow packets
+We installed OVS-AFXDP on the target server, and used the source server to
+generates 64-byte single UDP flow packets
 at the rate of 19~Mpps using the DPDK library.
@@
 XDP\_DRV: works on devices with XDP support, and
-XDP\_DRV ZC (Zero Copy): works on devices with XDP\_XDP zerecopy support.
+XDP\_DRV ZC (Zero Copy): works on devices with XDP\_XDP zerocopy support.
 For all our experiments, we use the XDP\_DRV ZC mode.
@@
 
-Applying OVS's PMD netdev avoids these problems and improving the rxdrop performance
-from $0.5Mpss$ to $3Mpps$.  Ideally, receing packets from XSK should not need to context
-switch to into kernel as mentioned in steps~\ref{afxdprx}. So an ideal implementation
+Applying OVS's PMD netdev avoids these problems and improves the rxdrop performance
+from $0.5Mpss$ to $3Mpps$.  Ideally, receiving packets from XSK should not need to context
+switch into kernel as mentioned in steps~\ref{afxdprx}. So an ideal implementation
 should show minimal context switches with most of the CPU time stays in userspace. 
@@
 the RX ring for new packets.  Each round of receive polling processes a batch of
-packet, up to maximum of 32 packets. In addition, we enable 1GB huge memory page support
+packets, up to a maximum of 32 packets. In addition, we enable 1GB huge memory page support
 to reduce the page fault overhead.  With the above fixes/optimizations, the rxdrop
@@
 and perf record/report.  With the above new design, we observe the new bottleneck
-is at the UMEM memory pool APIs we introduce to maintain the UMEM freelist, as well as the
-packet metadata allocation, struct dp\_packet, for every receiving packet.
-We introduce two major API, umem\_elem\_get(), which gets N free element from the
+is at the UMEM memory pool APIs we introduce to maintain the UMEM freelist, as well as at the
+packet metadata allocation, struct dp\_packet, for every received packet.
+We introduce two major APIs, umem\_elem\_get(), which gets N free element from the
 UMEM freelist, and umem\_elem\_put(), which places back the free UMEM buffer to
@@
 Due to the nature of list\_head, elements are sparse in memory.
-Consumer always takes the first element, head, and producer also place back
-from the head of the list.
+Consumer always takes the first element, head, and producer also places back
+to the head of the list.
 \item ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
@@
 \item ptr\_stack: Similar to ptr\_ring, an array of elements are allocated
-in continuous memory region, but both consumer and producer takes elements
+in continuous memory region, but both consumer and producer take elements
 from beginning of the array.
@@
 require a mutex lock before accessing UMEM freelist. Linux perf reports
-pthread mutex lock related APIs as one of the top CPU utilization function.
+pthread mutex lock related APIs as ones of the top CPU utilization functions.
 We change our design by 1) allocating per-queue UMEM region and 2)
 allocating one pmd thread per queue. As a result, no lock is needed because
-the each queue has only one thread and its own set of UMEM elements.
+each queue has only one thread and its own set of UMEM elements.
 
 \subsubsection{Metadata Pre-allocation}
-Moving forward, Linux perf shows that the packets metadata allocation
+Moving forward, Linux perf shows that the packet metadata allocation
 takes a lot of CPU cycles, i.e.,  dp\_packet\_init(), dp\_packet\_set\_data().
 So instead of allocating packet metadata at packet reception time, we
-implemented two data structures and compare their performance:
+implement two data structures and compare their performance:
 \begin{itemize}
@@
 is used for both packet data and metadata.
-\item Separate from packet buffer: This design simply allocate a continuous memory region
+\item Separate from packet buffer: This design simply allocates a continuous memory region
 storing an array of packet metadata, and initialize them as much as we can.
@@
 By using strace, we saw that the OVS-AFXDP l2fwd experiement calls sendto system call
-at very high frequeuce, because from Figure~\ref{afxdptx}, we design to check
-the success of send (step 4) immediately after issusing send (step 3).
+at very high frequency, because from Figure~\ref{afxdptx}, we design to check
+the success of send (step 4) immediately after issuing send (step 3).
 We change the design by calling send syscall (step 3) only when TX ring is
-closed to full, e.g., when 3/4 ring elements have been used.
+close to full, e.g., when 3/4 ring elements have been used.
 In specific case, instead of issuing send syscall for a batch of 32 packets
-and making sure they are done, we only issue send when there are 512 outstanding
+and keeping polling to check transmission is done, we only issue send when there are 512 outstanding
 packets.  With this change, the OVS-AFXDP l2fwd experiment shows around $14Mpps$.

### 2018-10-23 paper.tex
Edit ch 01

--- 2018-10-23 paper.tex
@@
 flexible platform abstractions for network functions, and is being ported
-to a variety of platforms.  Among the various different ways of using eBPF,
-OVS has been exploring the power of eBPF in three:
+to a variety of platforms. In Linux kernel, users can attach eBPF programs
+to TC and XDP hook point as shown in Fig~\ref{ebpf-hook}. Based on this
+design, we explore the possibilities of utilizing eBPF to implement OVS
+datapath in threefold:
 1) in-kernel flow processing by attaching eBPF programs to TC,
-2) offloading a subset of processing to XDP (eXpress Data Path),
+2) offloading a subset of flow processing to XDP (eXpress Data Path),
 and 3) moving the flow processing to userspace by using AF\_XDP.
 
-Attaching eBPF to TC, OVS-eBPF project, started first with the most aggressive goal:
-we planned to re-implement the entire features of OVS kernel datapath under net/openvswitch/* into eBPF code.
-We worked around a couple of limitations, for example, the lack of TLV (Type-Length-Value) support
-led us to redefine a binary kernel-user API using a fixed-length array;
-and without a dedicated way to execute a packet, we created a dedicated
-device for user to kernel packet transmission,
-with a different BPF program attached to handle packet execute logic.
-Currently, OVS-eBPF can satisfy most of the basic feature requirement
-for flow processing, including tunneling protocol support.
-But more complicated processing such as connection tracking, NAT,
-(de)fragmentation, and ALG are still under discussion.
-
-Moving one layer below TC is called XDP (eXpress Data Path), a much faster layer for packet
-processing, but with almost no extra packet metadata and limited BPF helpers support.
-Depending on the complexity of flows, OVS can offload a subset of its flow processing to XDP
-when feasible. However, the fact that XDP has fewer helper function support implies that either
-1) only very limited number of flows are eligible for offload, or 2) more flow processing logic
-needed to be done in native eBPF code.
+\begin{figure}
+{\scriptsize
+\begin{verbatim}
+
++-----------------------+
+|    Network Stack      |
++---------^-------------+
+          |
++---------|-------------+
+| Traffic Control (TC)  | --> eBPF TC hook
++---------^-------------+
+          |
++---------|-------------+
+|     Driver (XDP)      | --> eBPF XDP hook
++---------^-------------+
+          |
++---------|-------------+
+|     Network card      |
++---------^-------------+
+          |
+          | packets
+
+\end{verbatim}
+}
+\vspace{-1.0em}
+\caption{eBPF program hook points in the Linux kernel.}
+\label{ebpf-hook}
+\vspace{-1.0em}
+\end{figure}
+
+Firstly, in OVS-eBPF project, we attach flow processing eBPF programs to TC.
+We start with the most aggressive goal that we plan to re-implement the entire
+features of OVS kernel datapath under net/openvswitch/* into eBPF code.
+We work around a couple of eBPF limitations, for example, the lack of TLV
+(Type-Length-Value) support leads us to redefine a binary kernel-user API using
+a fixed-length array; without a dedicated way to execute a packet, we
+create a dedicated device that attached a eBPF program to handle packet
+execute logic for userspace to kernel packet transmission.
+Currently, OVS-eBPF can satisfy most of the basic feature for flow processing
+and tunneling protocol support, and we are investigating more complicated
+features such as connection tracking, NAT, (de)fragmentation, and ALG.
+
+We can attach eBPF programs one layer below TC to XDP (eXpress Data Path).
+It is a much faster layer for packet processing, but there is almost no extra
+packet metadata available, and XDP only provides limited kernel helpers.
+Depending on the type of flows, OVS can offload a subset of its flow processing
+to XDP. However, the fact that XDP has fewer helper function support implies
+that either 1) only very limited number of flows are eligible for offload,
+or 2) more flow processing logic needed to be implemented in native eBPF code.
+For example, it is more difficult for OVS datapath to provide tunnel support in
+XDP, since lightweight tunnel kernel helpers is not available.
 
 XDP provides another way for interacting with userspace programs, called AF\_XDP. 
-AF\_XDP is another socket interface for control plane and a shared
-memory API for accessing packets from userspace application.
-OVS today has another full-fledged
-datapath implementation in userspace, called dpif-netdev, mostly used by DPDK community.
-OVS-AFXDP project treats the AF\_XDP as a fast packet-I/O channel, and move
-most of the flow processing into the userspace OVS.
-
-A rough comparison of the two project is shown in Table~\ref{compare}
-This paper describes the design, implementation, and evaluation of the
-OVS-eBPF and OVS-AFXDP projects.
+AF\_XDP is a socket interface for control plane and a shared memory API for
+accessing packets from userspace application. With AF\_XDP, OVS-AFXDP project
+redirect packets to userspace, and processes the packets by OVS's full-fledged
+userspace datapath implementation, dpif-netdev. In this approach, OVS-AFXDP
+project treats the AF\_XDP as a fast packet-I/O channel, and move most of
+the flow processing into the userspace OVS.
+
+This paper focuses on the OVS-eBPF and OVS-AFXDP projects.
+A rough comparison of the two project is shown in Table~\ref{proj-compare}.
+The remainder of this paper is organized as follows:
+In Section~\ref{sec:background}, we first provide some background information
+on eBPF, XDP, and AF\_XDP.  We then present the design, implementation, and
+evaluation of the OVS-eBPF and OVS-AFXDP projects in Section~\ref{sec:ebpf}
+and Section~\ref{sec:af_xdp} respectively. Finally,
+Section~\ref{sec:conclusion} concludes the paper and discusses the future work.
+
 \begin{table}
@@
   BPF Code Size & Large & Minimal \\ \hline
-  Extensibility & Low & High \\ \hline
+  Extensibility & Good & Great \\ \hline
   \end{tabular}
@@
 }
-\label{compare}
+\label{proj-compare}
 \end{table}
@@
 \section{Background}
+\label{sec:background}
 
@@
 \section{OVS eBPF Datapath}
+\label{sec:ebpf}
 \myfigcross{ovsebpf}{5.5in}{The overall architecture of OVS eBPF datapath
@@
 \section{Userspace OVS with AF\_XDP}
+\label{sec:af_xdp}
 \begin{figure}
@@
 \section{Conclusion}
+\label{sec:conclusion}
 This paper describes two eBPF projects related to OVS: OVS-eBPF and OVS-AFXDP.

### 2018-10-24 paper.tex
fix sendmsg, and bib

@@
            -------------------
-(3) Issue send syscall to eth2's XSK
+(3) Issue sendmsg syscall to eth2's XSK
 (4) Packet transmission is completed
@@
 To send to eth2 using eth2's XSK, ovs-vswitchd first gets four packet buffers,
-\{4, 5, 6, 7\} from eth2's UMEM freelist and copy the packet data from eth1's UMEM
+\{4, 5, 6, 7\} from eth2's UMEM freelist and copies the packet data from eth1's UMEM
 to eth2's \{4, 5, 6, 7\}.  Then TX descriptors for the four packets are
 created and placed into eth2's TX ring (step 2).
-At step 3, send syscall is issued to signal the kernel to start transmit.
-As the send in XSK is asynchronos and send syscall only return return zero when
+At step 3, sendmsg syscall is issued to signal the kernel to start transmit.
+As the send in XSK is asynchronos and sendmsg syscall only return return zero when
 no error, ovs-vswitch polls the COMPLETION ring to make sure these four
 packets have been sent out (step 4).
-Once the four packets' descriptors shown up at the COMPLETION ring, at step 5,
-ovs-vswitchd recycles their UMEM elememnts back to the eth2's UMEM freelist.
-In addition, the original four packet buffer from eth1, \{1, 2, 3, 4\}, is
+Once the four packets' descriptors show up at the COMPLETION ring, at step 5,
+ovs-vswitchd recycles their UMEM elements back to eth2's UMEM freelist.
+In addition, the original four packet buffers from eth1, \{1, 2, 3, 4\}, are
 also recycled back to eth1's UMEM freelist.
@@
 One server, the target server, has an Intel 40GbE XL710 single port NIC,
-and the other server, the source server, has Netronome NFP-4000 40GbE device.
-The target server runs the Linux kernel 4.19-rc4, with i40e driver supporting
+and the other server, the source server, has a Netronome NFP-4000 40GbE device.
+The target server runs Linux kernel 4.19-rc4, with i40e driver supporting
 the AF\_XDP mode.
-We installed OVS-AFXDP on the target server, and the other server,
-the source server, generates 64-byte single UDP flow packets
+We installed OVS-AFXDP on the target server, and used the source server to
+generates 64-byte single UDP flow packets
 at the rate of 19~Mpps using the DPDK library.
-On the target server, we also disable the Intel spectre and meltdown fixes.
+On the target server, for demonstrating the best performance,
+we disabled the Intel Spectre and Meltdown fixes~\cite{jesper_xdp_perf_drop, retpoline}.
 
@@
 the success of send (step 4) immediately after issusing send (step 3).
-We change the design by calling send syscall (step 3) only when TX ring is
+We change the design by calling sendmsg syscall (step 3) only when TX ring is
 closed to full, e.g., when 3/4 ring elements have been used.
-In specific case, instead of issuing send syscall for a batch of 32 packets
+In specific case, instead of issuing sendmsg syscall for a batch of 32 packets
 and making sure they are done, we only issue send when there are 512 outstanding
@@
 techniques such as batching is critical. Here we apply batching in a
-couple of places such as issuing send syscall, packet reception and
+couple of places such as issuing sendmsg syscall, packet reception and
 transmission. In summary, OVS-AFXDP shows performance below:

### 2018-10-24 paper.tex
fix xdp related citation

@@
 There are several hook points where eBPF programs may be attached in recent
-Linux kernels. XDP is still an eBPF program, but its attachment point is at the
+Linux kernels. XDP~\cite{xdp,xdp2} is still an eBPF program, but its attachment point is at the
 lowest level of the network stack.  Due to its unique attachment point, the XDP
@@
 recent high speed network devices, such as 40G/100G NICs.
-Performance evaluation~\cite{danialafpacket,johnafxdp} of AF\_PACKET
+Performance evaluation~\cite{af_packet_v4,af_packet_performance} of AF\_PACKET
 shows less than 2 million packets per second using single core.
 
-AF\_XDP was proposed and upstreamed to Linux kernel since 4.18~\cite{afxdp} 
+AF\_XDP was proposed and upstreamed to Linux kernel since 4.18~\cite{af_xdp_lwn} 
 The core idea behind the AF\_XDP is to leverage the XDP eBPF program's

### 2018-10-24 paper.tex

--- 2018-10-24 paper.tex
@@
            -------------------
-(3) Issue sendmsg syscall to eth2's XSK
+(3) Issue send syscall to eth2's XSK
 (4) Packet transmission is completed
@@
 created and placed into eth2's TX ring (step 2).
-At step 3, sendmsg syscall is issued to signal the kernel to start transmit.
-As the send in XSK is asynchronos and sendmsg syscall only return return zero when
+At step 3, send syscall is issued to signal the kernel to start transmit.
+As the send in XSK is asynchronous and send syscall only returns zero when
 no error, ovs-vswitch polls the COMPLETION ring to make sure these four
@@
 at the rate of 19~Mpps using the DPDK library.
-On the target server, for demonstrating the best performance,
-we disabled the Intel Spectre and Meltdown fixes~\cite{jesper_xdp_perf_drop, retpoline}.
+On the target server, we also disable the Intel spectre and meltdown fixes.
 
@@
 By using strace, we saw that the OVS-AFXDP l2fwd experiement calls sendto system call
-at very high frequeuce, because from Figure~\ref{afxdptx}, we design to check
-the success of send (step 4) immediately after issusing send (step 3).
-We change the design by calling sendmsg syscall (step 3) only when TX ring is
-closed to full, e.g., when 3/4 ring elements have been used.
-In specific case, instead of issuing sendmsg syscall for a batch of 32 packets
-and making sure they are done, we only issue send when there are 512 outstanding
+at very high frequency, because from Figure~\ref{afxdptx}, we design to check
+the success of send (step 4) immediately after issuing send (step 3).
+We change the design by calling send syscall (step 3) only when TX ring is
+close to full, e.g., when 3/4 ring elements have been used.
+In specific case, instead of issuing send syscall for a batch of 32 packets
+and keeping polling to check transmission is done, we only issue send when there are 512 outstanding
 packets.  With this change, the OVS-AFXDP l2fwd experiment shows around $14Mpps$.
@@
 techniques such as batching is critical. Here we apply batching in a
-couple of places such as issuing sendmsg syscall, packet reception and
+couple of places such as issuing send syscall, packet reception and
 transmission. In summary, OVS-AFXDP shows performance below:

### 2018-10-24 paper.tex
fix afxdp section

--- 2018-10-24 paper.tex
@@
 a userspace socket interface. In other word, AF\_XDP socket family connects the
-XDP packet receving/sending path to the userspace, bypassing the rest of the
+XDP packet receving/sending path to the userspace, by-passing the rest of the
 Linux networking stacks.
@@
 call. Unlike AF\_PACKET which uses the send() and receive() syscalls with packet buffer as parameter,
-XSK introduces two rings in userspace: the RX ring and the TX ring.
+XSK introduces two rings in userspace: the Rx ring and the Tx ring.
 The userspace program using XSK needs to properly configure and maintain
-the RX and TX ring structure in order to receive and send packets.
-In addition, all the packet data buffers used in TX/RX ring are allocated from a specific
-memory area called UMEM which consists of a number of fixed size chunks.
-The UMEM also has two rings: the FILL ring and the COMPLETION ring.
-A descriptor in RX/TX ring or in FILL/COMPLETION ring points to the element
-in UMEM by its address.  The address is not system's virtual or physical address
-but simply an offset within the entire UMEM memory region.
+the Rx and Tx ring structures in order to receive and send packets.
+In addition, all the packet data buffers used in Tx/Rx rings are allocated from a specific
+memory area called umem which consists of a number of fixed size chunks.
+The umem also has two rings: the Fill ring and the Completion ring.
+A descriptor in Rx/Tx ring or in Fill/Completion ring points to the element
+in umem by its address. The address is not system's virtual or physical address
+but simply an offset within the entire umem memory region.
 
 For example, to receive packets, a set of descriptors pointing to empty
-packet buffer needs to be filled into the FILL ring.  When packets arrives,
-the userspace program checks the RX ring, fetches the packet data from the
-desciptors, and refills the empty buffer to the FILL ring structure, in order to receive new
+packet buffer needs to be filled into the Fill ring.  When packets arrives,
+the userspace program checks the Rx ring, fetches the packet data from the
+desciptors, and refills the empty buffer to the Fill ring structure, in order to receive new
 incoming packets.  For sending packets, a set of descriptors pointing the
-packet buffer with contents filled to the TX ring then issus send() system call.
+packet buffers with contents filled to the Tx ring, then issus sendmsg() system call.
 It is up to the userspace program to make sure whether the packets have been
-sent or not, by checking the COMPLETION ring structure.
-In summary, users of XSK needs to properly interact with the following four
+sent or not, by checking the Completion ring structure.
+In summary, users of XSK needs to properly program following four
 rings:
 \begin{itemize}
-\item FILL ring: for users to fill UMEM addresses to kernel for receiving packets.
-\item RX ring: for users to access recevied packets. 
-\item TX ring: for users to place packets needed to be sent.
-\item COMPLETION ring: for users to make sure packets are sent.
+\item Fill ring: for users to fill umem addresses to kernel for receiving packets.
+\item Rx ring: for users to access recevied packets. 
+\item Tx ring: for users to place packets needed to be sent.
+\item Completion ring: for users to make sure packets are sent.
 \end{itemize}
@@
 Call socket() syscall to create a XSK,
-set up its RX and TX ring buffer, allocate a UMEM region,
-and set up FILL/COMPLETION ring of the UMEM.
+set up its Rx and Tx ring buffer, allocate a umem region,
+and set up Fill/Completion ring of the umem.
 
@@
 
-\item Populate the UMEM FILL ring: Get a couple of UMEM elements and place into FILL ring.
+\item Populate the umem Fill ring: Get a couple of umem elements and place into Fill ring.
 \end{enumerate}
 On the other hand, when a AF\_XDP netdev is detached or closed by user,
-ovs-vswitchd closes the XSK socket, free the UMEM memory region, and
+ovs-vswitchd closes the XSK socket, free the umem memory region, and
 unload the eBPF program and map.
 
-In order to properly set-up the RX/TX/FILL/COMPLETION rings, we implement a
-UMEM memory management API. Basically, it's a list maintaining the unused elements
-in UMEM, called UMEM freelist, and function calls to get/put elements into the list.
+In order to properly program Rx/Tx/Fill/Completion rings, we implement a
+umem memory management layer, call umempool. Basically, it's a data structure
+maintaining the unused/available elements in umem, and GET and PUT access APIs.
 \subsection{Packet Reception}
@@
 \begin{verbatim}
-UMEM: total 8 elements, each has 2K chunk size
+umem: total 8 elements, each has 2K chunk size
                  +-------------------------------+
@@
 (1) Initial Stat
-           -------------------          UMEM freelist
-FILL Qu  ...| 1 | 2 | 3 | 4 | ...       {5, 6, 7, 8}
-           -------------------
-           -------------------
-RX ring  ...|   |   |   |   | ...
+           -------------------          umempool
+Fill r   ...| 1 | 2 | 3 | 4 | ...       {5, 6, 7, 8}
+           -------------------
+           -------------------
+Rx ring  ...|   |   |   |   | ...
            -------------------
 (2) 4 Packets Arrive
-           -------------------          UMEM freelist
-FILL Qu  ...|   |   |   |   | ...       {5, 6, 7, 8}
+           -------------------          umempool
+Fill r   ...|   |   |   |   | ...       {5, 6, 7, 8}
           --------------------
            -------------------
-RX ring  ...| 1 | 2 | 3 | 4 | ...
-           -------------------
-(3) Get 4 free UMEM elemet and refill
-           -------------------          UMEM freelist
-FILL Qu  ...| 5 | 6 | 7 | 8 | ...       {} 
-           -------------------
-           -------------------
-RX ring  ...| 1 | 2 | 3 | 4 | ...
+Rx ring  ...| 1 | 2 | 3 | 4 | ...
+           -------------------
+(3) Get 4 free umem elemet and refill
+           -------------------          umempool
+Fill r   ...| 5 | 6 | 7 | 8 | ...       {} 
+           -------------------
+           -------------------
+Rx ring  ...| 1 | 2 | 3 | 4 | ...
            -------------------
@@
 
-(5) Recycle to UMEM freelist, goto step (2)
-            +-------------------        UMEM freelist
-FILL Qu  ...| 5 | 6 | 7 | 8 | ...       {1, 2, 3, 4}
+(5) Recycle to umempool, goto step (2)
+            +-------------------        umempool
+Fill r  ...| 5 | 6 | 7 | 8 | ...       {1, 2, 3, 4}
             +------------------
            -------------------
-RX ring  ...|   |   |   |   | ...
+Rx ring  ...|   |   |   |   | ...
            -------------------
@@
 \vspace{-1.0em}
-\caption{An example of how OVS programs the FILL ring and RX ring when
+\caption{An example of how OVS programs the Fill ring and Rx ring when
 processing incoming AF\_XDP packets.}
@@
 
-Figure~\ref{afxdprx} shows how ovs-vswitchd sets up the FILL ring and
-the RX ring for receving packets from XSK.  For simplicity to demonstrate
-how AF\_XDP works, assume that there are only eight UMEM buffers, with each buffer's
-size equals 2KB (?).  Initially at step 1, ovs-vswitchd fills four available UMEM elements into
-the FILL ring and waits for incoming packets.  When there are incoming packets,
-at step 2, the FILL ring's four buffer elems have been consumed and moved
-to the RX ring.  In order to keep receiving packets, ovs-vswitchd gets another
-four available UMEM elements from the UMEM freelist, and fills into FILL ring (step 3).
+Figure~\ref{afxdprx} shows how ovs-vswitchd sets up the Fill ring and
+the Rx ring for receving packets from XSK.  For simplicity to demonstrate
+how AF\_XDP works, assume that there are only eight umem buffers, with each buffer's
+size equals 2KB (?).  Initially at step 1, ovs-vswitchd fills four available umem elements into
+the Fill ring and waits for incoming packets.  When there are incoming packets,
+at step 2, the Fill ring's four buffer elems have been consumed and moved
+to the Rx ring.  In order to keep receiving packets, ovs-vswitchd gets another
+four available umem elements from the umempool, and fills into Fill ring (step 3).
 Then ovs-vswitchd creates the metadata needed for the four packet
@@
 the dpif-netdev layer for parse, lookup and action executions.
-Finally, when ovs-vswitchd finishes processing the UMEM buffer, a recycle
-mechanism is triggered to place this buffer back to UMEM freelist (step 5).
-Step 5 makes sure that there are always available elements in FILL ring, so
+Finally, when ovs-vswitchd finishes processing the umem buffer, a recycle
+mechanism is triggered to place this buffer back to umempool (step 5).
+Step 5 makes sure that there are always available elements in Fill ring, so
 that the underlying XDP program in kernel can keep processing packets 
 while the userspace ovs-vswitchd is processing the previous
-received packets on RX ring.
+received packets on Rx ring.
 When step 5 finishes, ovs-vswitchd goes back to step 2, waiting for
@@
 \begin{verbatim}
- OVS dpif-netdev has 4 packets at UMEM:{1, 2, 3, 4} received
+ OVS dpif-netdev has 4 packets at umem:{1, 2, 3, 4} received
  from eth1 and plan to send to eth2.
-(1) Get 4 free elements from eth2's UMEM, copy packet data to it
-(2) Create TX descriptors and place into TX ring
-           -------------------        eth2 UMEM freelist
-TX ring  ...| 4 | 5 | 6 | 7 | ...       {1, 2, 3, 8}
-           -------------------
-           -------------------
-COMPL q  ...|   |   |   |   | ...
+(1) Get 4 free elements from eth2's umem, copy packet data to it
+(2) Create TX descriptors and place into Tx ring
+           -------------------        eth2 umempool
+Tx ring  ...| 4 | 5 | 6 | 7 | ...       {1, 2, 3, 8}
+           -------------------
+           -------------------
+Compl r  ...|   |   |   |   | ...
            -------------------
@@
 (4) Packet transmission is completed
-           -------------------        eth2 UMEM freelist
-TX ring  ...|   |   |   |   | ...       {1, 2, 3, 8}
-           -------------------
-           -------------------
-COMPL q  ...| 4 | 5 | 6 | 7 | ...
-           -------------------
-(5) Recycle to eth2 UMEM
-           -------------------        eth2 UMEM freelist
-TX ring  ...|   |   |   |   | ...       {1, 2, 3, 4,
+           -------------------        eth2 umempool
+Tx ring  ...|   |   |   |   | ...       {1, 2, 3, 8}
+           -------------------
+           -------------------
+Compl r  ...| 4 | 5 | 6 | 7 | ...
+           -------------------
+(5) Recycle to eth2 umem
+           -------------------        eth2 umempool
+Tx ring  ...|   |   |   |   | ...       {1, 2, 3, 4,
            -------------------           5, 6, 7, 8}
            -------------------
-COMPL q  ...|   |   |   |   | ...
-           -------------------
-(6) Recycle {1, 2, 3, 4} to eth1's UMEM 
+Compl r  ...|   |   |   |   | ...
+           -------------------
+(6) Recycle {1, 2, 3, 4} to eth1's umem 
 \end{verbatim}
@@
 \vspace{-1.0em}
-\caption{An example of how OVS programs the COMPLETION ring and TX ring when
+\caption{An example of how OVS programs the COMPLETION ring and Tx ring when
 processing incoming AF\_XDP packets from one netdev and sending to another netdev.}
@@
 To send to eth2 using eth2's XSK, ovs-vswitchd first gets four packet buffers,
-\{4, 5, 6, 7\} from eth2's UMEM freelist and copies the packet data from eth1's UMEM
+\{4, 5, 6, 7\} from eth2's umempool and copies the packet data from eth1's umem
 to eth2's \{4, 5, 6, 7\}.  Then TX descriptors for the four packets are
-created and placed into eth2's TX ring (step 2).
+created and placed into eth2's Tx ring (step 2).
 At step 3, sendmsg syscall is issued to signal the kernel to start transmit.
@@
 Once the four packets' descriptors show up at the COMPLETION ring, at step 5,
-ovs-vswitchd recycles their UMEM elements back to eth2's UMEM freelist.
+ovs-vswitchd recycles their umem elements back to eth2's umempool.
 In addition, the original four packet buffers from eth1, \{1, 2, 3, 4\}, are
-also recycled back to eth1's UMEM freelist.
+also recycled back to eth1's umempool.
 
 \subsection{Performance Evaluation}
-All of our performance results use a hardware test bed that consists of
+All of our performance results use a hardware test bed consisting of
 two Intel Xeon E5 2440 v2 1.9GHz servers, each with 1 CPU socket and
@@
 The target server runs Linux kernel 4.19-rc4, with i40e driver supporting
-the AF\_XDP mode.
+the AF\_XDP zero copy mode~\cite{af_xdp_patch_i40e, af_xdp_patch_zc}.
 We installed OVS-AFXDP on the target server, and used the source server to
@@
 
-For all our experiments, we fuse a microbenchmark program, called xdpsock,
+For all our experiments, we use a microbenchmark program, called xdpsock,
 as the baseline to compare with OVS-AFXDP implementation.
@@
 
-Additionally, AF\_XDP provides three operation modes:
-XDP\_SKB: works on devices using generic XDP~\cite{genericxdp},
-XDP\_DRV: works on devices with XDP support, and
-XDP\_DRV ZC (Zero Copy): works on devices with XDP\_XDP zerocopy support.
-For all our experiments, we use the XDP\_DRV ZC mode.
+Additionally, AF\_XDP provides three operation modes, from slowest to fastest:
+skb mode: works on devices using generic XDP~\cite{genericxdp},
+driver mode: works on devices with XDP support, and
+zero-copy driver mode: works on devices with XDP\_XDP zerocopy support.
+For all our experiments, we use the zero-copy driver mode for showing the best performance.
 
@@
 \end{itemize}
-\mycomment{Compared to xdpsock, ovs-vswitchd adds on top of it a UMEM memory pool,
+\mycomment{Compared to xdpsock, ovs-vswitchd adds on top of it a umempool,
 and packets parsing, lookup and action execution.}
@@
 packets, up to a maximum of 32 packets. In addition, we enable 1GB huge memory page support
-to reduce the page fault overhead.  With the above fixes/optimizations, the rxdrop
-performance increases to around $10Mpps$
+to reduce the page fault overhead.
 
 \subsubsection{Memory Management and Locks}\label{memorym}
-For every stage of the optimization, we use Linux perf extensively, e.g., perf stat
+For every stage of optimizations, we use Linux perf extensively, e.g., perf stat
 and perf record/report.  With the above new design, we observe the new bottleneck
-is at the UMEM memory pool APIs we introduce to maintain the UMEM freelist, as well as at the
+is at the umempool APIs we introduce, as well as at the
 packet metadata allocation, struct dp\_packet, for every received packet.
 We introduce two major APIs, umem\_elem\_get(), which gets N free element from the
-UMEM freelist, and umem\_elem\_put(), which places back the free UMEM buffer to
-the freelist, similar to the concept of producer (put) and consumer (get).
-We implemented three different data structures of UMEM memory pool as below
+umempool, and umem\_elem\_put(), which places back the free umem buffer to
+the umempool, similar to the concept of producer (put) and consumer (get).
+We implemented three different data structures of umempool as below
 to compare their performance.
 \begin{itemize}
-\item list\_head: Similar to Linux kernel's list\_head, each element contains its
-UMEM element address, and a next pointer points to next free element.
-Due to the nature of list\_head, elements are sparse in memory.
-Consumer always takes the first element, head, and producer also places back
-to the head of the list.
-\item ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
+\item FIFO-list\_head: We reversed first 8-byte of the umem buffer as a next pointer,
+pointing to the next available element in umempool, and implement FIFO access get/set
+APIs. The list\_head pointer always points to the first available umem element.
+\item FILO-ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
 allocated in continuous memory region.
@@
 them into where head pointer points to.
-\item ptr\_stack: Similar to ptr\_ring, an array of elements are allocated
+\item FIFO-ptr\_array: Similar to Linux kernel's skb\_array~\cite{skbarray},
+an array of elements are allocated
 in continuous memory region, but both consumer and producer take elements
@@
 
-Although we allocate one UMEM per netdev, there might be multiple queues per
-netdev sharing the same UMEM. As a result, the above three data structures
-require a mutex lock before accessing UMEM freelist. Linux perf reports
+Although we allocate one umem per netdev, there might be multiple queues per
+netdev sharing the same umem. As a result, the above three data structures
+require a mutex lock before accessing umem freelist. Linux perf reports
 pthread mutex lock related APIs as ones of the top CPU utilization functions.
-We change our design by 1) allocating per-queue UMEM region and 2)
+We change our design by 1) allocating per-queue umem region and 2)
 allocating one pmd thread per queue. As a result, no lock is needed because
-each queue has only one thread and its own set of UMEM elements.
-
-\subsubsection{Metadata Pre-allocation}
+each queue has only one thread and its own set of umem elements.
+
+\subsubsection{Metadata Allocation}
 Moving forward, Linux perf shows that the packet metadata allocation
 takes a lot of CPU cycles, i.e.,  dp\_packet\_init(), dp\_packet\_set\_data().
-So instead of allocating packet metadata at packet reception time, we
-implement two data structures and compare their performance:
+So instead of allocating packet metadata at packet reception time, we pre-allocate
+the packet metadata and implement two data structures to compare their performances:
 \begin{itemize}
 \item Embedding in packet buffer: Since we already allocate 2K chunk for each
-UMEM packet buffer, we reserve the first 256-byte in each UMEM element as struct dp\_packet
+umem packet buffer, we reserve the first 256-byte in each umem element as struct dp\_packet
 and initialize the dp\_packet as much as we can at allocation time.
@@
 
-With the above design change, we found that using the ptr\_stack in section~\ref{memorym} with
+With the above design change, we found that using the FIFO-ptr\_array in section~\ref{memorym} with
 separate from packet buffer design yields the best performance. Largely due to both data structures
@@
 For example, setting up 32 packet metadatas in an array definitely incurs
-less cache misses than setting up in 32 elements in UMEM.
-And getting 32 free UMEM elements from ptr\_stack has similar benefits.
+less cache misses than setting up in 32 elements in umem.
+And getting 32 free umem elements from ptr\_array has similar benefits.
 With the above decision, the OVS rxdrop shows around $19Mpps$, closed

### 2018-10-24 paper.tex
fix FIFO / FILO

@@
 \begin{itemize}
-\item FIFO-list\_head: We reversed first 8-byte of the umem buffer as a next pointer,
+\item FILO-list\_head: We reversed first 8-byte of the umem buffer as a next pointer,
 pointing to the next available element in umempool, and implement FIFO access get/set
 APIs. The list\_head pointer always points to the first available umem element.
-\item FILO-ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
+\item FIFO-ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
 allocated in continuous memory region.
@@
 them into where head pointer points to.
-\item FIFO-ptr\_array: Similar to Linux kernel's skb\_array~\cite{skbarray},
+\item FILO-ptr\_array: Similar to Linux kernel's skb\_array~\cite{skbarray},
 an array of elements are allocated
@@
 
-With the above design change, we found that using the FIFO-ptr\_array in section~\ref{memorym} with
+With the above design change, we found that using the FILO-ptr\_array in section~\ref{memorym} with
 separate from packet buffer design yields the best performance. Largely due to both data structures

### 2018-10-24 paper.tex
Edit ch-2.1 section title, ch-2.3

@@
 
+\subsection{OVS Forwarding Model}
 OVS is widely used in virtualized data center environments as a software
@@
 
-\subsection{OVS Forwarding Model}
-
 \begin{figure}
@@
 
-There are several hook points where eBPF programs may be attached in recent
-Linux kernels. XDP~\cite{xdp,xdp2} is still an eBPF program, but its attachment point is at the
-lowest level of the network stack.  Due to its unique attachment point, the XDP
-has its own {\em context}; the input parameters to the XDP eBPF program and return
-values have different meanings than the TC eBPF program. XDP shows performance
-closed to line rate of the device, because the XDP program is triggered immediately
-at the network device driver's packet receiving code path.
-Due to its lowest hook point at the networking stack, XDP input parameter has only
-pointer to the beginning and end of the packet data plus a few metadatas.
-XDP also supports accessing to eBPF maps and tail calls, but with much
-less number of helper functions than the eBPF program at TC hook.
+There are several hook points where eBPF programs can be attached in recent
+Linux kernels. XDP~\cite{xdp,xdp2} is another eBPF program hook point where
+its attachment point is at the lowest level of the network stack.
+XDP demonstrates high performance that closed to the line rate of the device,
+since the eBPF programs attached to XDP hook point are triggered immediately
+in the network device driver's packet receiving code path. For the same reason,
+eBPF program in XDP can only access the packet data and a few metadata.
+XDP supports accessing to eBPF maps and tail calls, but much less number of
+helper functions is available compared to the TC hook point.
 
@@
 takes the ELF file, parses the programs and maps information from it and
-issues BPF syscalls to load the program.  If it passes the BPF verifier,
-then the program is attached to the hook point (in this case, TC/XDP), and
+issues BPF syscalls to load the program.  If the program passes the BPF verifier,
+then it is attached to the hook point (in this case, TC/XDP), and
 subsequent packets through the TC/XDP ingress hook will trigger execution of the

### 2018-10-24 paper.tex
Edit ch 2.4

@@
 AF\_XDP is a new Linux address family that aims for high packet I/O
-performance. Traditionally, a userspace program receives packets from
-kernel through the socket API.  By creating a socket with address family
-such as AF\_PACKET, userspace programs can receive/send the raw packets
+performance. It enables another way for a userspace program to receive
+packets from kernel through the socket API. For example, currently, creating
+a socket with address family AF\_PACKET, userspace programs can receive/send
+the raw packets
 at the device driver layer.  Although the AF\_PACKET family has been using
@@
 
-AF\_XDP was proposed and upstreamed to Linux kernel since 4.18~\cite{af_xdp_lwn} 
+AF\_XDP was proposed and upstreamed to Linux kernel since 4.18~\cite{af_xdp_lwn}.
 The core idea behind the AF\_XDP is to leverage the XDP eBPF program's
-early access to the raw packet, and create a high speed channel from the XDP directly to
-a userspace socket interface. In other word, AF\_XDP socket family connects the
-XDP packet receving/sending path to the userspace, by-passing the rest of the
-Linux networking stacks.
-An AF\_XDP socket, called XSK, is create using the normal socket() system
-call. Unlike AF\_PACKET which uses the send() and receive() syscalls with packet buffer as parameter,
-XSK introduces two rings in userspace: the Rx ring and the Tx ring.
-The userspace program using XSK needs to properly configure and maintain
-the Rx and Tx ring structures in order to receive and send packets.
-In addition, all the packet data buffers used in Tx/Rx rings are allocated from a specific
-memory area called umem which consists of a number of fixed size chunks.
-The umem also has two rings: the Fill ring and the Completion ring.
-A descriptor in Rx/Tx ring or in Fill/Completion ring points to the element
-in umem by its address. The address is not system's virtual or physical address
-but simply an offset within the entire umem memory region.
-
-For example, to receive packets, a set of descriptors pointing to empty
-packet buffer needs to be filled into the Fill ring.  When packets arrives,
-the userspace program checks the Rx ring, fetches the packet data from the
-desciptors, and refills the empty buffer to the Fill ring structure, in order to receive new
-incoming packets.  For sending packets, a set of descriptors pointing the
-packet buffers with contents filled to the Tx ring, then issus sendmsg() system call.
-It is up to the userspace program to make sure whether the packets have been
-sent or not, by checking the Completion ring structure.
-In summary, users of XSK needs to properly program following four
-rings:
+early access to the raw packet, and provide a high speed channel from the XDP
+directly to a userspace socket interface. In other word, AF\_XDP socket family
+connects the XDP packet receiving/sending path to the userspace, by-passing the
+rest of the Linux networking stacks.
+An AF\_XDP socket, called XSK, is created by using the normal socket() system
+call. Unlike AF\_PACKET which uses the send() and receive() syscalls with packet
+buffer as parameter, XSK introduces two rings in userspace: the Rx ring and the
+Tx ring. These two rings are per-XSK data structure that the userspace program
+needs to properly configure and maintain in order to receive and send packets.
+Moreover, to provide zerocopy support for AF\_XDP, all the packet data buffers
+used in Tx/Rx rings are allocated from a specific memory region called \em{umem}
+which consists of a number of fixed size memory chunks.
+Two rings, the Fill ring and the Completion ring, are associated with a umem,
+and a umem can be shared between multiple XSKs.
+Each element in the Rx/Tx ring or in Fill/Completion ring is a descriptor that
+contains an address that points to a chunk in umem. The address is not system's
+virtual or physical address but simply an offset within the entire umem memory
+region.
+
+For example, to receive packets from XSK, firstly, a set of descriptors pointing
+to empty packet buffer need to be pushed into the Fill ring.  When a packet
+arrives, kernel pops descriptors from the Fill ring, fills in the data into the
+memory chunks pointed by the descriptors, and pushes the descriptors back to
+the Rx ring.
+The userspace program then checks the Rx ring, fetches the packet data from the
+descriptors, and refills the empty buffer back to the Fill ring structure, so
+that kernel can fill in new incoming packets later on.
+For sending packets, the userspace program pushes a set of descriptors that
+point to the packet buffers to the Tx ring, then issues sendmsg() system call.
+Kernel consumes packet buffers asscoaited with the Tx ring, and pushes
+transmitted descriptors to the Tx ring.
+The userspace program then checks the Completion ring to determine whether the
+packets have been sent.
+In summary, XSK users need to properly program following four rings:
 \begin{itemize}
 \item Fill ring: for users to fill umem addresses to kernel for receiving packets.
-\item Rx ring: for users to access recevied packets. 
+\item Rx ring: for users to access received packets.
 \item Tx ring: for users to place packets needed to be sent.
-\item Completion ring: for users to make sure packets are sent.
+\item Completion ring: for users to check if packets are sent.
 \end{itemize}
-Unlike AF\_PACKET which is bound to entire netdev, the XSK is more fine-grained.
-XSK is bound to a specific queue id on a device, so only the traffic/flow 
-sent to the queue id shows up in the XSK.
+Unlike AF\_PACKET which is bound to entire netdev, the binding of XSK is more
+fine-grained. XSK is bound to a specific queue on a device, so only the
+traffic/flow sent to the queue shows up in the XSK.
 

### 2018-10-24 paper.tex
Edit ch 4.1 and ch 4.2

@@
 \subsection{Datapath and Netdev Interface}
-OVS includes a userspace datapath interface (dpif) implemenation, called dpif-netdev.
-Unlike the dpif-netlink which talks to the Linux Open vSwitch kernel
-module through netlink API, the dpif-netdev works completely in userspace.
-As a result, dpif-netdev expects to receive and send packets from its userspace
-interface.  One major use case of dpif-netdev is OVS-DPDK~\cite{ovsdpdk},
-where the packet reception and transmission are all done in DPDK's userspace
+OVS provides a userspace datapath interface (dpif) implementation, called
+dpif-netdev. The dpif-netdev userspace datapath receives and transmits
+packets from its userspace interface.  One major use case of dpif-netdev
+is OVS-DPDK~\cite{ovsdpdk}, where the packet reception and transmission
+are all conducted in DPDK's userspace
 library. The dpif-netdev is designed to be agnostic to how the network
 device accessing the packets, by an abstraction layer called netdev.
-So Packets might come from DPDK packet I/O library,
+Therefore, packets can come from DPDK packet I/O library,
 a Linux AF\_PACKET socket API, or in our case, the AF\_XDP socket interface,
 as long as each mechanism implements its own netdev interface.
-Once dpif-netdev receives a packet, it follows the same mechanism doing
+Once dpif-netdev receives a packet, it follows the same mechanism performing
 parse, lookup the flow table, and apply actions to the packet.
@@
 }
-ovs-vswitch does the following steps to bring up the
+ovs-vswitchd does the following steps to bring up the
 AF\_XDP netdev:
@@
 \item Attach a XDP program to the netdev's queue:
-The XDP program OVS attaches is fixed and consists
-of only a few line of code, which receives the packet and
-calls the \texttt{bpf\_redirect\_map()} helper function with
-the XSK eBPF maps and key equals zero.
+OVS attaches a simple and fixed XDP program to each netdev's queue.
+The program only consists of a few lines of code, which receives the
+packets and redirects them to XSK by calling the \texttt{bpf\_redirect\_map()}
+helper function.
 
@@
 \item Load and configure the XSK eBPF map:
-The XSK eBPF maps consists of key value pairs, where key is an u32 index and value is
-the file descript of the XSK.  ovs-vswitchd simply programs one entry to the map with
-key equals zero and the file descriptor, fd, of the XSK as its value.
-As a result, the XDP program calling \texttt{bpf\_redirect\_map} with key=0 will
-forwards the packet to the XSK.
+The XSK eBPF maps consists of key value pairs, where key is an u32 index and
+value is the file descriptor of the XSK.  ovs-vswitchd programs an entry to
+the map with the key as queue id and the file descriptor, fd, of the XSK as
+its value.
+Therefore, the XDP program calling \texttt{bpf\_redirect\_map} will derive
+the corresponding XSK with the queue id.
 
@@
 \end{enumerate}
-On the other hand, when a AF\_XDP netdev is detached or closed by user,
+Finally, when a AF\_XDP netdev is detached or closed by user,
 ovs-vswitchd closes the XSK socket, free the umem memory region, and
@@
 
+\subsection{umem memory management}
 In order to properly program Rx/Tx/Fill/Completion rings, we implement a
@@
 maintaining the unused/available elements in umem, and GET and PUT access APIs.
-\subsection{Packet Reception}
+\subsubsection{Packet Reception}
 \begin{figure}
@@
 
-\subsection{Packet Transmission}
+\subsubsection{Packet Transmission}
 \begin{figure}

### 2018-10-25 paper.tex
Edit ch 4.3.1

@@
 Moreover, to provide zerocopy support for AF\_XDP, all the packet data buffers
-used in Tx/Rx rings are allocated from a specific memory region called \em{umem}
+used in Tx/Rx rings are allocated from a specific memory region called {\em umem}
 which consists of a number of fixed size memory chunks.
@@
 umem memory management layer, call umempool. Basically, it's a data structure
-maintaining the unused/available elements in umem, and GET and PUT access APIs.
+maintaining the unused/available elements in umem with GET and PUT access
+APIs. We will demonstrate the use case of umempool with packet reception and
+transmission in the following subsections.
+
 \subsubsection{Packet Reception}
@@
 \begin{verbatim}
-umem: total 8 elements, each has 2K chunk size
+umem: total 8 elements, each has 2KB chunk size
                  +-------------------------------+
@@
 Figure~\ref{afxdprx} shows how ovs-vswitchd sets up the Fill ring and
-the Rx ring for receving packets from XSK.  For simplicity to demonstrate
-how AF\_XDP works, assume that there are only eight umem buffers, with each buffer's
-size equals 2KB (?).  Initially at step 1, ovs-vswitchd fills four available umem elements into
+the Rx ring for receiving packets from XSK.  For simplicity, in this example,
+we assume that there are only eight umem buffers, and each buffer's
+size is 2KB.  Initially, at step 1, ovs-vswitchd pushes four available umem
+elements into
 the Fill ring and waits for incoming packets.  When there are incoming packets,
-at step 2, the Fill ring's four buffer elems have been consumed and moved
+at step 2, the Fill ring's four buffer elements have been consumed and moved
 to the Rx ring.  In order to keep receiving packets, ovs-vswitchd gets another
 four available umem elements from the umempool, and fills into Fill ring (step 3).
-Then ovs-vswitchd creates the metadata needed for the four packet
+Then, ovs-vswitchd creates the metadata needed for the four packet
 buffer \{1, 2, 3, 4\}, i.e., struct dp\_packet and struct dp\_packet\_batch, and passes to 

### 2018-10-25 paper.tex
Fix up affiliations

@@
     % Alphabetical order.
-    \textrm{William Tu} & \textrm{Joe Stringer} & \textrm{Yifeng Sun} & \textrm{Yi-Hung Wei}\\
-    \textrm{u9012063@gmail.com} & \textrm{joe@ovn.org} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
+    \textrm{William Tu$^1$} & \textrm{Joe Stringer$^2$} & \textrm{Yifeng Sun$^1$} & \textrm{Yi-Hung Wei$^1$}\\
+    \textrm{u9012063@gmail.com} & \textrm{joe@cilium.io} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
   \end{matrix}
@@
 }
-{VMware Inc.}{}
+{
+  \(
+  \begin{matrix}
+    \textrm{$^1$VMware Inc.} & \textrm{$^2$Cilium.io}\\
+  \end{matrix}
+  \)
+}{}
 

### 2018-10-25 paper.tex
abstract: Minor grammar fixes

@@
 extensible by providing an eBPF virtual machine with safety guarantee.
-Open vSwitch, OVS, is a software switch running majorly in Linux
+Open vSwitch (OVS), is a software switch running majorly in Linux
 operating systems, its fast path packet process is implemented in Linux kernel
@@
 to Linux TC.
-On the other hand, the OVS-AFXDP project aims to by-passing the kernel
-using AF\_XDP socket and moves most of the flow processing features
+On the other hand, the OVS-AFXDP project aims to by-pass the kernel
+using an AF\_XDP socket and moves most of the flow processing features
 into userspace.
 We demonstrate the feasibility of implementing OVS datapath with the
-aforementioned technologies and present the performance number in this paper.
+aforementioned technologies and present the performance numbers in this paper.
 \end{abstract}

### 2018-10-25 paper.tex
abstract: Minor grammar fixes

@@
     % Alphabetical order.
-    \textrm{William Tu$^1$} & \textrm{Joe Stringer$^2$} & \textrm{Yifeng Sun$^1$} & \textrm{Yi-Hung Wei$^1$}\\
-    \textrm{u9012063@gmail.com} & \textrm{joe@cilium.io} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
+    \textrm{William Tu$^1$} & \textrm{Joe Stringer} & \textrm{Yifeng Sun$^1$} & \textrm{Yi-Hung Wei$^1$}\\
+    \textrm{u9012063@gmail.com} & \textrm{joe@ovn.org} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
   \end{matrix}
@@
 }
-{
-  \(
-  \begin{matrix}
-    \textrm{$^1$VMware Inc.} & \textrm{$^2$Cilium.io}\\
-  \end{matrix}
-  \)
-}{}
+{$^1$VMware Inc.}{}
 

### 2018-10-25 paper.tex
intro: Minor grammar fixes

@@
 flexible platform abstractions for network functions, and is being ported
-to a variety of platforms. In Linux kernel, users can attach eBPF programs
-to TC and XDP hook point as shown in Fig~\ref{ebpf-hook}. Based on this
+to a variety of platforms. In the Linux kernel, users can attach eBPF programs
+to TC and XDP hook points as shown in Fig~\ref{ebpf-hook}. Based on this
 design, we explore the possibilities of utilizing eBPF to implement OVS
@@
 execute logic for userspace to kernel packet transmission.
-Currently, OVS-eBPF can satisfy most of the basic feature for flow processing
+Currently, OVS-eBPF can satisfy most of the basic features for flow processing
 and tunneling protocol support, and we are investigating more complicated
@@
 For example, it is more difficult for OVS datapath to provide tunnel support in
-XDP, since lightweight tunnel kernel helpers is not available.
+XDP, since lightweight tunnel kernel helpers are not available.
 
@@
 AF\_XDP is a socket interface for control plane and a shared memory API for
-accessing packets from userspace application. With AF\_XDP, OVS-AFXDP project
-redirect packets to userspace, and processes the packets by OVS's full-fledged
+accessing packets from userspace application. Using AF\_XDP, the OVS-AFXDP project
+redirect packets to userspace, and processes the packets using OVS's full-fledged
 userspace datapath implementation, dpif-netdev. In this approach, OVS-AFXDP
-project treats the AF\_XDP as a fast packet-I/O channel, and move most of
-the flow processing into the userspace OVS.
+project treats the AF\_XDP as a fast packet-I/O channel.
 

### 2018-10-25 paper.tex
intro: Minor grammar fixes

@@
     % Alphabetical order.
-    \textrm{William Tu$^1$} & \textrm{Joe Stringer} & \textrm{Yifeng Sun$^1$} & \textrm{Yi-Hung Wei$^1$}\\
-    \textrm{u9012063@gmail.com} & \textrm{joe@ovn.org} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
+    \textrm{William Tu$^1$} & \textrm{Joe Stringer$^2$} & \textrm{Yifeng Sun$^1$} & \textrm{Yi-Hung Wei$^1$}\\
+    \textrm{u9012063@gmail.com} & \textrm{joe@cilium.io} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
   \end{matrix}
@@
 }
-{$^1$VMware Inc.}{}
+{
+  \(
+  \begin{matrix}
+    \textrm{$^1$VMware Inc.} & \textrm{$^2$Cilium.io}\\
+  \end{matrix}
+  \)
+}{}
 

### 2018-10-25 paper.tex
intro: Improve comparison table

--- 2018-10-25 paper.tex
@@
   Packet Rate & Low & High \\ \hline
-  Driver Support & No & Yes \\ \hline
-  Minimal Kernel Req. & 4.18 & Unknown \\ \hline
+  Minimal Kernel Req. & 4.10 & 4.18 \\ \hline
+  Driver Support Req. & No & Yes \\ \hline
   BPF Code Size & Large & Minimal \\ \hline

### 2018-10-25 paper.tex
Edit ch 4.3.2

--- 2018-10-25 paper.tex
@@
     % Alphabetical order.
-    \textrm{William Tu$^1$} & \textrm{Joe Stringer$^2$} & \textrm{Yifeng Sun$^1$} & \textrm{Yi-Hung Wei$^1$}\\
-    \textrm{u9012063@gmail.com} & \textrm{joe@cilium.io} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
+    \textrm{William Tu} & \textrm{Joe Stringer} & \textrm{Yifeng Sun} & \textrm{Yi-Hung Wei}\\
+    \textrm{u9012063@gmail.com} & \textrm{joe@ovn.org} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
   \end{matrix}
@@
 }
-{
-  \(
-  \begin{matrix}
-    \textrm{$^1$VMware Inc.} & \textrm{$^2$Cilium.io}\\
-  \end{matrix}
-  \)
-}{}
+{VMware Inc.}{}
 
@@
 extensible by providing an eBPF virtual machine with safety guarantee.
-Open vSwitch (OVS), is a software switch running majorly in Linux
+Open vSwitch, OVS, is a software switch running majorly in Linux
 operating systems, its fast path packet process is implemented in Linux kernel
@@
 to Linux TC.
-On the other hand, the OVS-AFXDP project aims to by-pass the kernel
-using an AF\_XDP socket and moves most of the flow processing features
+On the other hand, the OVS-AFXDP project aims to by-passing the kernel
+using AF\_XDP socket and moves most of the flow processing features
 into userspace.
 We demonstrate the feasibility of implementing OVS datapath with the
-aforementioned technologies and present the performance numbers in this paper.
+aforementioned technologies and present the performance number in this paper.
 \end{abstract}
@@
 flexible platform abstractions for network functions, and is being ported
-to a variety of platforms. In the Linux kernel, users can attach eBPF programs
-to TC and XDP hook points as shown in Fig~\ref{ebpf-hook}. Based on this
+to a variety of platforms. In Linux kernel, users can attach eBPF programs
+to TC and XDP hook point as shown in Fig~\ref{ebpf-hook}. Based on this
 design, we explore the possibilities of utilizing eBPF to implement OVS
@@
 execute logic for userspace to kernel packet transmission.
-Currently, OVS-eBPF can satisfy most of the basic features for flow processing
+Currently, OVS-eBPF can satisfy most of the basic feature for flow processing
 and tunneling protocol support, and we are investigating more complicated
@@
 For example, it is more difficult for OVS datapath to provide tunnel support in
-XDP, since lightweight tunnel kernel helpers are not available.
+XDP, since lightweight tunnel kernel helpers is not available.
 
@@
 AF\_XDP is a socket interface for control plane and a shared memory API for
-accessing packets from userspace application. Using AF\_XDP, the OVS-AFXDP project
-redirect packets to userspace, and processes the packets using OVS's full-fledged
+accessing packets from userspace application. With AF\_XDP, OVS-AFXDP project
+redirect packets to userspace, and processes the packets by OVS's full-fledged
 userspace datapath implementation, dpif-netdev. In this approach, OVS-AFXDP
-project treats the AF\_XDP as a fast packet-I/O channel.
+project treats the AF\_XDP as a fast packet-I/O channel, and move most of
+the flow processing into the userspace OVS.
 
@@
 \begin{verbatim}
- OVS dpif-netdev has 4 packets at umem:{1, 2, 3, 4} received
- from eth1 and plan to send to eth2.
-(1) Get 4 free elements from eth2's umem, copy packet data to it
+ OVS sending 4 packets recevied from eth1 to eth2 by XSK.
+(1) Get 4 free elements from eth2's umempool, copy packet data to them
 (2) Create TX descriptors and place into Tx ring
@@
            -------------------
-(6) Recycle {1, 2, 3, 4} to eth1's umem 
+(6) Recycle the 4 packet buffers back to eth1's umempool
 \end{verbatim}
@@
 \caption{An example of how OVS programs the COMPLETION ring and Tx ring when
-processing incoming AF\_XDP packets from one netdev and sending to another netdev.}
+processing incoming AF\_XDP packets from one netdev and sending them to another
+netdev.}
 \label{afxdptx}
@@
 Figure~\ref{afxdptx} shows the process for sending packets to the XSK.
-Assuming that ovs-vswitch has one bridge and two ports: eth1 and eth2.
-And there is a flow entry showing
-\texttt{in\_port=eth1, ..., actions=..., output:eth2}, meaning to forward
+In this example, there is one bridge and two ports, eth1 and eth2, in OVS.
+Both ports are configured with AF\_XDP support, and there is a flow,
+\texttt{in\_port=eth1, actions=output:eth2} that forwards
 packets received from eth1 to eth2.
-And both eth1 and eth2 are configured with AF\_XDP support.
-
-Initially, assume ovs-vswitchd receives four packets \{1, 2, 3, 4\} from eth1's XSK
-To send to eth2 using eth2's XSK, ovs-vswitchd first gets four packet buffers,
+
+Initially, assume ovs-vswitchd receives four packets from eth1's XSK.
+To send packets to eth2 using eth2's XSK, ovs-vswitchd first gets four packet buffers,
 \{4, 5, 6, 7\} from eth2's umempool and copies the packet data from eth1's umem
-to eth2's \{4, 5, 6, 7\}.  Then TX descriptors for the four packets are
+to eth2's umem at \{4, 5, 6, 7\}.  Later on, TX descriptors for the four packets are
 created and placed into eth2's Tx ring (step 2).
-At step 3, sendmsg syscall is issued to signal the kernel to start transmit.
-As the send in XSK is asynchronos and sendmsg syscall only return return zero when
-no error, ovs-vswitch polls the COMPLETION ring to make sure these four
-packets have been sent out (step 4).
+At step 3, sendmsg syscall is issued to signal the kernel to start the transmission.
+As the send msg syscall in XSK is asynchronous, ovs-vswitchd needs to poll
+the COMPLETION ring to make sure that these four packets have been sent out
+(step 4).
 Once the four packets' descriptors show up at the COMPLETION ring, at step 5,
 ovs-vswitchd recycles their umem elements back to eth2's umempool.
-In addition, the original four packet buffers from eth1, \{1, 2, 3, 4\}, are
-also recycled back to eth1's umempool.
+The original four packet buffers from eth1 are recycled back to eth1's
+umempool as well.
 

### 2018-10-25 paper.tex
Edit 4.4.0 and 4.4.1

@@
 
-\subsection{Performance Evaluation}
-All of our performance results use a hardware test bed consisting of
+\subsection{Performance Evaluation and Optimization}
+All of our performance evaluation are conducted on a test bed consisting of
 two Intel Xeon E5 2440 v2 1.9GHz servers, each with 1 CPU socket and
 8 physical cores with hyperthreading enabled.
-One server, the target server, has an Intel 40GbE XL710 single port NIC,
-and the other server, the source server, has a Netronome NFP-4000 40GbE device.
-The target server runs Linux kernel 4.19-rc4, with i40e driver supporting
+The first server, the source server, is served as a traffic generator.
+It is equipped with a Netronome NFP-4000 40GbE device, and runs a customized
+traffic generator that generates 64-byte single UDP flow packets
+at the rate of $19Mpps$ using the DPDK library.
+The second server, the target server, is equipped with an Intel 40GbE XL710
+single port NIC, and it runs Linux kernel 4.19-rc4 with i40e driver supporting
 the AF\_XDP zero copy mode~\cite{af_xdp_patch_i40e, af_xdp_patch_zc}.
-We installed OVS-AFXDP on the target server, and used the source server to
-generates 64-byte single UDP flow packets
-at the rate of 19~Mpps using the DPDK library.
-On the target server, for demonstrating the best performance,
-we disabled the Intel Spectre and Meltdown fixes~\cite{jesper_xdp_perf_drop, retpoline}.
+We install OVS-AFXDP on the target server, and
+we disable the Intel Spectre and Meltdown fixes~\cite{jesper_xdp_perf_drop, retpoline}
+for demonstrating the best performance. In addition, we enable 1GB huge
+memory page support to reduce the page fault overhead.
 
@@
 as the baseline to compare with OVS-AFXDP implementation.
-xdpsock is an AF\_XDP sample program doing no packet processing, and can be configured
-to do simply dropping all packets, rxdrop, or forwarding the packet to the same port as
-it receives, l2fwd.
-In our testbed, we measured $19Mpps$ for xdpsock-rxdrop and $17Mpps$ for xdpsock-l2fwd.
-For OVS-AFXDP, we conduct similar two experiments, but with OVS's OpenFlow rule installed as
+xdpsock is an AF\_XDP sample program that can be configured into two modes.
+In the rxdrop mode, xdpsock simply drops all incoming packets without any
+packet processing. In the l2fwd mode, xdpsock forwards the packet to the
+same port from where it is received.
+In our testbed, we measure $19Mpps$ for xdpsock-rxdrop and $17Mpps$ for xdpsock-l2fwd.
+For OVS-AFXDP, we conduct two similar experiments, with OVS's OpenFlow rule installed as
 below:
@@
 
-Additionally, AF\_XDP provides three operation modes, from slowest to fastest:
-skb mode: works on devices using generic XDP~\cite{genericxdp},
-driver mode: works on devices with XDP support, and
-zero-copy driver mode: works on devices with XDP\_XDP zerocopy support.
-For all our experiments, we use the zero-copy driver mode for showing the best performance.
+AF\_XDP can provide the following three operation modes depending on the device
+and driver support. We list the three mode from the slowest to the fastest
+throughput.
+1) skb mode: works on devices using generic XDP~\cite{genericxdp},
+2) driver mode: works on devices with XDP support, and
+3) zero-copy driver mode: works on devices with XDP\_XDP zerocopy support.
+For all our experiments, we use the zero-copy driver mode since it provides the
+best performance.
 
 \subsubsection{PMD netdev}
-Our initial prototype shows pretty terrible performance, only $0.5Mpps$ for rxdrop.
-We observe that when dropping/forwarding packets under OVS-AFXDP, two CPUs show up
-100\% utilitization:
+In the beginning, our initial prototype of OVS-AFXDP only provides $0.5Mpps$
+throughput for rxdrop.
+We observe that when dropping/forwarding packets under OVS-AFXDP, two processes
+below can easily reach 100\% CPU utilization:
 \begin{itemize}
-\item ovs-vswitchd: This is the process keeps doing the send and receive steps
+\item ovs-vswitchd: This is the process that keeps doing the send and receive steps
 in Figure~\ref{afxdprx} and~\ref{afxdptx}.
@@
 and packets parsing, lookup and action execution.}
-We investigate using Linux perf and our first fix is to enable OVS's PMD-mode,
-Poll-Mode-Driver, to the afxdp netdev.
-
-In OVS, a non-pmd mode netdev does packet receiption by putting all the receiving
-netdev's file descriptors, fds, together and ovs-vswitchd polls them when there is an fd
-ready to perform I/O.  As a result, the fd of the XSK is {\em shared} with other fds, and
-we observed that the poll system call has pretty high overhead.
+With further investigation using Linux perf tools, our first improvement is to
+to enable OVS's Poll-Mode-Driver (PMD) mode to the afxdp netdev.
+
+In OVS's non-PMD mode, ovs-vswitchd does packet reception by putting all the
+receiving netdev's file descriptors (fd)s together and invokes the poll system
+call to determine if any one of the fd is ready to perform I/O.  As a result,
+the polling of XSK's fd  is {\em shared} with other fds, and
+we observed that the poll system call incurs high overhead.
 
 Applying OVS's PMD netdev avoids these problems and improves the rxdrop performance
-from $0.5Mpss$ to $3Mpps$.  Ideally, receiving packets from XSK should not need to context
-switch into kernel as mentioned in steps~\ref{afxdprx}. So an ideal implementation
-should show minimal context switches with most of the CPU time stays in userspace. 
-When enabling PMD netdev for afxdp, ovs-vswitch can be
-configured to create a {\em dedicated} thread for the packet processing.
-In OVS-AFXDP, we create a thread for each XSK's receive queue and the thread keeps polling
-the RX ring for new packets.  Each round of receive polling processes a batch of
-packets, up to a maximum of 32 packets. In addition, we enable 1GB huge memory page support
-to reduce the page fault overhead.
+from $0.5Mpss$ to $3Mpps$. This performance gain is because of reducing userspace
+and kernel space context switch overhead by omitting poll system call.
+That is instead of using poll system call, we implement packet reception by using
+PMD netdev for AFXDP, so that ovs-vswitchd allocate a dedicated thread for each
+XSK's receive queue and the thread keeps polling the RX ring for new packets.
+Currently, each round of receive polling processes a batch up to 32 packets.
 
@@
 We change our design by 1) allocating per-queue umem region and 2)
-allocating one pmd thread per queue. As a result, no lock is needed because
+allocating one PMD thread per queue. As a result, no lock is needed because
 each queue has only one thread and its own set of umem elements.
@@
 observed only $4Mpps$, compared to $17Mpps$ baseline xdpsock-l2fwd.
-We found that the OVS pmd thread under rxdrop
+We found that the OVS PMD thread under rxdrop
 has much fewer context switches compared to the l2fwd, indicating that
-the pmd process spends much more time in kernel space than in userspace.
+the PMD process spends much more time in kernel space than in userspace.
 By using strace, we saw that the OVS-AFXDP l2fwd experiement calls sendto system call

### 2018-10-25 paper.tex
Edit 4.4.2 - 4.4.4

--- 2018-10-25 paper.tex
@@
 For every stage of optimizations, we use Linux perf extensively, e.g., perf stat
-and perf record/report.  With the above new design, we observe the new bottleneck
-is at the umempool APIs we introduce, as well as at the
-packet metadata allocation, struct dp\_packet, for every received packet.
-We introduce two major APIs, umem\_elem\_get(), which gets N free element from the
-umempool, and umem\_elem\_put(), which places back the free umem buffer to
-the umempool, similar to the concept of producer (put) and consumer (get).
-We implemented three different data structures of umempool as below
-to compare their performance.
+and perf record/report.  With the above PMD netdev optimization, we observe the
+new bottleneck is the umempool APIs that we introduced.
+We introduce two major umempool APIs, umem\_elem\_get(), which gets N free
+element from the umempool, and umem\_elem\_put(), which places back the free
+umem buffer to the umempool. We implement three different data structures for
+umempool management as below.
 \begin{itemize}
-\item FILO-list\_head: We reversed first 8-byte of the umem buffer as a next pointer,
-pointing to the next available element in umempool, and implement FIFO access get/set
-APIs. The list\_head pointer always points to the first available umem element.
-\item FIFO-ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
-allocated in continuous memory region.
-Consumer takes elements from the tail pointer and when placing back, producer, puts
-them into where head pointer points to.
-\item FILO-ptr\_array: Similar to Linux kernel's skb\_array~\cite{skbarray},
-an array of elements are allocated
-in continuous memory region, but both consumer and producer take elements
-from beginning of the array.
+\item LIFO-list\_head: We reverse the first 8-byte of a umem chunk as a next
+pointer that pointing to the next available element in umempool, and implement
+LIFO (last in, first out) access get/put APIs. In this design, we maintain the
+list\_head pointer that always points to the first available umem element,
+and we get and put elements from list\_head.
+\item FIFO-ptr\_ring: This design maintains an extra FIFO (first in, first out)
+pointer ring that is similar to Linux kernel's ptr\_ring. The ring is an array
+of elements that allocated in a continuous memory region, where each element
+points to an available umem chunk. We keep track of head and tail pointer for
+the FIFO-ptr\_ring. The consumer gets elements from the tail pointer,
+and then the producer puts elements back to where head pointer points to.
+\item LIFO-ptr\_array: This design is similar to Linux kernel's
+skb\_array~\cite{skbarray} where an array of pointers are allocated
+in contiguous memory region. Each element in the array points to an available
+memory chunk in umem.  We keep track of a head pointer to the array,
+and both consumer and producer get/put elements from the head pointer.
 \end{itemize}
 
-Although we allocate one umem per netdev, there might be multiple queues per
-netdev sharing the same umem. As a result, the above three data structures
-require a mutex lock before accessing umem freelist. Linux perf reports
-pthread mutex lock related APIs as ones of the top CPU utilization functions.
-We change our design by 1) allocating per-queue umem region and 2)
+Initially, we allocate one umem per netdev. Since there might be multiple
+queues per netdev sharing the same umem, the above three data structures
+require a mutex lock to protect umempool accessing. In this design, Linux
+perf reports pthread mutex lock related APIs as one of the top CPU
+utilization function.
+We then change our design by 1) allocating per-queue umem region and 2)
 allocating one PMD thread per queue. As a result, no lock is needed because
@@
 So instead of allocating packet metadata at packet reception time, we pre-allocate
-the packet metadata and implement two data structures to compare their performances:
+the packet metadata and implement two data structures to compare their performance:
 \begin{itemize}
-\item Embedding in packet buffer: Since we already allocate 2K chunk for each
-umem packet buffer, we reserve the first 256-byte in each umem element as struct dp\_packet
-and initialize the dp\_packet as much as we can at allocation time.
+\item Embedding in packet buffer: As we already allocate 2KB chunk for each
+umem packet buffer, we reserve the first 256-byte in each umem chunk as struct dp\_packet
+and initialize the dp\_packet's packet independent fields  at allocation time.
 This is similar to the DPDK mbuf design~\cite{dpdkmbuf}, where a single memory buffer
 is used for both packet data and metadata.
-\item Separate from packet buffer: This design simply allocates a continuous memory region
-storing an array of packet metadata, and initialize them as much as we can.
+\item Separate from packet buffer: This design allocates a contiguous memory
+region storing an array of packet metadata, and initialize their packet
+independent fields.
 \end{itemize}
@@
 
-With the above design change, we found that using the FILO-ptr\_array in section~\ref{memorym} with
-separate from packet buffer design yields the best performance. Largely due to both data structures
-have better spatial locality and more batching friendly, when all data are allocated together.
-For example, setting up 32 packet metadatas in an array definitely incurs
-less cache misses than setting up in 32 elements in umem.
-And getting 32 free umem elements from ptr\_array has similar benefits.
-With the above decision, the OVS rxdrop shows around $19Mpps$, closed
-to the baseline xdpsock rxdrop performance.
+With the above design change, we find that using the LIFO-ptr\_array in
+Section~\ref{memorym} with separated packet metadata allocation yields the
+best performance. It is because both data structures have better spatial
+locality and there are more batching friendly.
+For example, accessing 32 packet metadata in an array incurs less cache
+misses than accessing the packet metadata 32 umem chunks.
+The ptr\_array for umempool management outperforms the other two designs
+for the similar reason. With the above design decisions, the OVS rxdrop
+can achieve similar performance as the baseline xdpsock rxdrop at around
+$19Mpps$.
 
@@
 %574.790 M/sec  
-We continued measuring the performance of OVS-AFXDP l2fwd and
-observed only $4Mpps$, compared to $17Mpps$ baseline xdpsock-l2fwd.
-We found that the OVS PMD thread under rxdrop
-has much fewer context switches compared to the l2fwd, indicating that
+With all the aforementioned optimizations, We continue measuring the
+performance of OVS-AFXDP l2fwd and observe only $4Mpps$, compared to
+$17Mpps$ baseline xdpsock-l2fwd.
+We find that the OVS PMD thread under rxdrop has much fewer context
+switches compared to the l2fwd, indicating that
 the PMD process spends much more time in kernel space than in userspace.
-By using strace, we saw that the OVS-AFXDP l2fwd experiement calls sendto system call
-at very high frequeuce, because from Figure~\ref{afxdptx}, we design to check
-the success of send (step 4) immediately after issusing send (step 3).
-We change the design by calling sendmsg syscall (step 3) only when TX ring is
-closed to full, e.g., when 3/4 ring elements have been used.
-In specific case, instead of issuing sendmsg syscall for a batch of 32 packets
-and making sure they are done, we only issue send when there are 512 outstanding
-packets.  With this change, the OVS-AFXDP l2fwd experiment shows around $14Mpps$.
+By using strace, we find that the OVS-AFXDP l2fwd experiment calls sendto
+system call at very high rate. It is because we design to check the
+completion of send immediately after issusing send as shown in step 3 and
+4 in Figure~\ref{afxdptx}.
+We modify the design by calling sendmsg syscall (step 3) only when TX ring is
+close to full, e.g., when 3/4 ring elements have been used.
+For example, instead of issuing sendmsg syscall for a batch of 32 packets
+and making sure they are finished, we only issue send when there are 512
+outstanding packets.  With this change, the OVS-AFXDP l2fwd experiment
+can achieve around $14Mpps$.
 

### 2018-10-26 paper.tex
Edit 4.4.5

@@
 \subsubsection{Summary}
-Through the step-by-step analysis, the key to performance is to keep
-the process CPU time in userspace, and making sure the userspace process
-is dedicated to do packet processing. Moreover, standard optmization
-techniques such as batching is critical. Here we apply batching in a
-couple of places such as issuing sendmsg syscall, packet reception and
-transmission. In summary, OVS-AFXDP shows performance below:
+We summarize the OVS-AFXDP in Table~\ref{compare}.
+Through the step-by-step analysis, the keys to performance improvement are to
+keep packet processing in userspace to avoid kernel userspace context switch,
+and allocating dedicated userspace processes for packet processing.
+Moreover, standard optimization techniques such as batching is critical
+to performance. We observe performance boost when apply batching to a couple
+of places such as issuing sendmsg syscall, packet reception and transmission.
 \begin{table}
@@
 \caption{\footnotesize
-Performance comparison of the xdpsock and and OVS-AFXDP.
+Performance comparison of the xdpsock and OVS-AFXDP.
 }

### 2018-10-26 paper.tex
comment out the table 1

@@
 This paper focuses on the OVS-eBPF and OVS-AFXDP projects.
-A rough comparison of the two project is shown in Table~\ref{proj-compare}.
+%A rough comparison of the two project is shown in Table~\ref{proj-compare}.
 The remainder of this paper is organized as follows:
@@
 Section~\ref{sec:conclusion} concludes the paper and discusses the future work.
-
+\mycomment{
 \begin{table}
@@
 \end{table}
-
+}
 \section{Background}
@@
 packets from its userspace interface.  One major use case of dpif-netdev
-is OVS-DPDK~\cite{ovsdpdk}, where the packet reception and transmission
+is OVS-DPDK, where the packet reception and transmission
 are all conducted in DPDK's userspace

### 2018-10-29 paper.tex
add page numbering
Fix typo in Figure 2

@@
-\documentclass[10pt]{sigplanconf}
+\documentclass[10pt,numbers,reprint]{sigplanconf}
 
@@
 The forwarding plane of OVS consists of two components;
-ovs-vswitch handles the complexity of the OpenFlow protocol, and the
+ovs-vswitchd handles the complexity of the OpenFlow protocol, and the
 datapath acts as a caching layer to optimize the performance.  A flow

### 2018-10-29 paper.tex
Minor fixes in ch1 and ch2

@@
 accessing packets from userspace application. Using AF\_XDP, the OVS-AFXDP project
-redirect packets to userspace, and processes the packets using OVS's full-fledged
+redirects packets to userspace, and processes the packets using OVS's full-fledged
 userspace datapath implementation, dpif-netdev. In this approach, OVS-AFXDP
@@
 The OVS datapath first performs packet parsing to extract relevant protocol
-headers from the packet and store it locally in a manner that is efficient for
+headers from the packet and stores it locally in a manner that is efficient for
 performing lookups (flow key), then it uses this information to look into the
@@
 contains an address that points to a chunk in umem. The address is not system's
-virtual or physical address but simply an offset within the entire umem memory
+virtual or physical address but simply an offset within the umem memory
 region.
 
-For example, to receive packets from XSK, firstly, a set of descriptors pointing
-to empty packet buffer need to be pushed into the Fill ring.  When a packet
+For example, to receive packets from XSK, firstly, the userspace program firstly
+pushes a set of descriptors pointing to empty packet buffer into the Fill ring.
+When a packet
 arrives, kernel pops descriptors from the Fill ring, fills in the data into the
@@
 fine-grained. XSK is bound to a specific queue on a device, so only the
-traffic/flow sent to the queue shows up in the XSK.
+traffic sent to the queue shows up in the XSK.
 

### 2018-10-29 paper.tex
Cite ovs-bpf project paper

@@
 
-Firstly, in OVS-eBPF project, we attach flow processing eBPF programs to TC.
-We start with the most aggressive goal that we plan to re-implement the entire
-features of OVS kernel datapath under net/openvswitch/* into eBPF code.
-We work around a couple of eBPF limitations, for example, the lack of TLV
-(Type-Length-Value) support leads us to redefine a binary kernel-user API using
-a fixed-length array; without a dedicated way to execute a packet, we
-create a dedicated device that attached a eBPF program to handle packet
-execute logic for userspace to kernel packet transmission.
-Currently, OVS-eBPF can satisfy most of the basic features for flow processing
-and tunneling protocol support, and we are investigating more complicated
-features such as connection tracking, NAT, (de)fragmentation, and ALG.
+Firstly, in OVS-eBPF project~\cite{extensible_ovs}, we attach flow processing
+eBPF programs to TC. We start with the most aggressive goal that we plan to
+re-implement the entire features of OVS kernel datapath under net/openvswitch/*
+into eBPF code. We work around a couple of eBPF limitations, for example, the
+lack of TLV (Type-Length-Value) support leads us to redefine a binary
+kernel-user API using a fixed-length array; without a dedicated way to execute
+a packet, we create a dedicated device that attached a eBPF program to handle
+packet execute logic for userspace to kernel packet transmission. Currently,
+OVS-eBPF can satisfy most of the basic features for flow processing and
+tunneling protocol support, and we are investigating more complicated features
+such as connection tracking, NAT, (de)fragmentation, and ALG.
 
@@
 There are several hook points where eBPF programs can be attached in recent
-Linux kernels. XDP~\cite{xdp,xdp2} is another eBPF program hook point where
+Linux kernels. XDP~\cite{xdp,xdp2,xdp-conext18} is another eBPF program hook point where
 its attachment point is at the lowest level of the network stack.
@@
 \item Flow key. This is the internal representation of the protocol headers
-    and metadata for the current packet being processed, defined in the P4~\cite{p4} language.
+    and metadata for the current packet being processed.
 \item Flow Table. This is a hash table whose key is the 32-bit hash value
@@
 of $0.28$ Mpps from baseline is the result of the deparser writing all known
-packet header metadata back to the packet, and due to extraneous memory copies
-introduced by the P4-to-eBPF compiler. With a more intelligent deparser
+packet header metadata back to the packet. With a more intelligent deparser
 implementation, and by using recent improvements to the eBPF

### 2018-10-29 paper.tex
Add XDP reference
Remove P4 related stuff
Add conclusion

@@
 
+The eBPF project originally had the ambitious goal of replacing the fixed
+kernel datapath with a dynamic datapath that may be injected on demand, while
+keeping the design and operation essentially the same. This would solve
+distribution and maintenance issues, as the version of the datapath would be
+distributed with the OVS userspace package; furthermore, it would outline a
+path to allow OVS to continue to process packets in the kernel without the use
+of an OVS-specific kernel module, acheiving the goals of other recent
+work~\cite{ovs_tc}. Collectively, this would allow the datapath logic to be
+extended and modified more easily, while reducing the maintenance burden for
+that code. Longer term, with a more flexible code base, many performance and
+feature improvements could be made and more easily, as the constraints around
+backwards compatibility of the fixed kernel ABI could be relaxed.
+
+While the basic use cases for implementing the OpenFlow forwarding model can be
+achieved in a straightforward manner in OVS-eBPF, implementing the full
+capabilities of the current OVS datapath is more difficult within the bounds of
+the BPF verification engine~\cite{extensible_ovs}. This paper explored an
+additional approach which makes use of another datapath implementation that OVS
+contains in userspace. Rather than building an extensible datapath and
+inserting this logic into the kernel at runtime, the packets can be efficiently
+directed to userspace for network processing using AF\_XDP sockets. The
+userspace datapath implementation has none of the distribution, maintenance or
+compatibility issues that the kernel implementation has, so at face value this
+appears to serve many of the original goals of the OVS-eBPF project.
+
+Over the course of exploring the use of AF\_XDP sockets, multiple design
+choices were made with the goal of optimal performance in mind. Future
+evaluation of this project should investigate the tradeoffs made with these
+decisions, and how they affect the configuration and deployment of the
+implementation. This implementation contains a lot of technical parallels with
+the OVS-DPDK project, which puts certain constraints on deployment such as the
+configuration of devices and the dedication of CPU and memory resources to the
+datapath. In some deployment environments, these may not be feasible
+constraints to place on OVS, and the existing kernel implementation does not
+require such constraints.
+
 \bibliography{paper.bib}

### 2018-10-29 paper.tex
Minor modification in ch 4

@@
 \vspace{-1.0em}
-\caption{OVS Architecture AF\_XDP}
+\caption{OVS Architecture with AF\_XDP}
 \label{ovsafxdp}
@@
 Therefore, packets can come from DPDK packet I/O library,
-a Linux AF\_PACKET socket API, or in our case, the AF\_XDP socket interface,
+a Linux AF\_PACKET socket API, or AF\_XDP socket interface,
 as long as each mechanism implements its own netdev interface.
@@
 In order to properly program Rx/Tx/Fill/Completion rings, we implement a
-umem memory management layer, call umempool. Basically, it's a data structure
+umem memory management layer, call umempool. It is a data structure
 maintaining the unused/available elements in umem with GET and PUT access
@@
 the Fill ring and waits for incoming packets.  When there are incoming packets,
-at step 2, the Fill ring's four buffer elements have been consumed and moved
+at step 2, the Fill ring's four buffer elements will be consumed and moved
 to the Rx ring.  In order to keep receiving packets, ovs-vswitchd gets another
@@
 and packets parsing, lookup and action execution.}
-With further investigation using Linux perf tools, our first improvement is to
-to enable OVS's Poll-Mode-Driver (PMD) mode to the afxdp netdev.
+With further investigation using Linux perf tools, our first improvement is
+to enable OVS's Poll-Mode-Driver (PMD) mode to the AF\_XDP netdev.
 
@@
 With the above design change, we find that using the LIFO-ptr\_array in
-Section~\ref{memorym} with separated packet metadata allocation yields the
+Section~\ref{memorym} with separated packet metadata allocation for
+metadata allocation yields the
 best performance. It is because both data structures have better spatial
-locality and there are more batching friendly.
+locality and they are more batching friendly.
 For example, accessing 32 packet metadata in an array incurs less cache
-misses than accessing the packet metadata 32 umem chunks.
+misses than accessing the packet metadata in 32 umem chunks.
 The ptr\_array for umempool management outperforms the other two designs
@@
 4 in Figure~\ref{afxdptx}.
-We modify the design by calling sendmsg syscall (step 3) only when TX ring is
+We modify this design by calling sendmsg syscall (step 3) only when TX ring is
 close to full, e.g., when 3/4 ring elements have been used.

### 2018-11-01 paper.tex
minor fix
Import LPC.cls
If missing iftex.sty,
  $ apt-get install texlive-generic-extra
Update to LPC template
Header in the even page mess up. Not sure how to fix it.

@@
-\documentclass[10pt,numbers,reprint]{sigplanconf}
+\documentclass[10pt,numbers,reprint,sigconf,authorversion]{lpc}
 
@@
 \begin{document}
+\pdfinfo{
+/Title (Bringing the Power of eBPF to Open vSwitch)
+/Author (William Tu, Joe Stringer, Yifeng Sun and Yi-Hung Wei)}
+
 \title{Bringing the Power of eBPF to Open vSwitch}
-\authorinfo{
-  \(
-  \begin{matrix}
-    % Alphabetical order.
-    \textrm{William Tu$^1$} & \textrm{Joe Stringer$^2$} & \textrm{Yifeng Sun$^1$} & \textrm{Yi-Hung Wei$^1$}\\
-    \textrm{u9012063@gmail.com} & \textrm{joe@cilium.io} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
-  \end{matrix}
-  \)
-}
-{
-  \(
-  \begin{matrix}
-    \textrm{$^1$VMware Inc.} & \textrm{$^2$Cilium.io}\\
-  \end{matrix}
-  \)
-}{}
+\author{William Tu}
+\affiliation{%
+      \institution{VMware Inc.}}
+\email{u9012063@gmail.com}
+
+\author{Joe Stringer}
+\affiliation{%
+      \institution{Cilium.io}}
+\email{joe@cilium.io}
+
+\author{Yifeng Sun}
+\affiliation{%
+      \institution{VMware Inc.}}
+\email{pkusunyifeng@gmail.com}
+
+\author{Yi-Hung Wei}
+\affiliation{%
+      \institution{VMware Inc.}}
+\email{yihung.wei@gmail.com}
+
+\acmDOI{}
+\setcounter{secnumdepth}{0}
+\setcopyright{none}
 
@@
 This paper describes two eBPF projects related to OVS: OVS-eBPF and OVS-AFXDP.
-
 The eBPF project originally had the ambitious goal of replacing the fixed

### 2018-11-04 paper.tex
shorten the conference string
fix afxdp tx operation description
fix typo, move figure a little right

--- 2018-10-29 paper.tex
@@
 \begin{verbatim}
-
             +-----------------------+
@@
                       | packets
-
 \end{verbatim}
@@
 Kernel consumes packet buffers asscoaited with the Tx ring, and pushes
-transmitted descriptors to the Tx ring.
+transmitted descriptors to the Completion ring.
 The userspace program then checks the Completion ring to determine whether the
@@
 At step 3, sendmsg syscall is issued to signal the kernel to start the transmission.
-As the send msg syscall in XSK is asynchronous, ovs-vswitchd needs to poll
+As the sendmsg syscall in XSK is asynchronous, ovs-vswitchd needs to poll
 the COMPLETION ring to make sure that these four packets have been sent out

### 2018-11-05 paper.tex
fix bib
remove deparser

@@
 recent high speed network devices, such as 40G/100G NICs.
-Performance evaluation~\cite{af_packet_v4,af_packet_performance} of AF\_PACKET
+Performance evaluation~\cite{af_packet_v4} of AF\_PACKET
 shows less than 2 million packets per second using single core.
@@
 If a matching flow has less than 32 actions to execute, the rest of the
-actions is no-op, and we short-cut the action execution to the deparser.
+actions is no-op, and we short-cut the no-op action execution. 
 If a matching flow has more than 32 actions, then the eBPF datapath delegates
@@
 the parser+lookup the first time the packet was received. The rest of the
-lookup, actions execution, and deparser is then executed as per the description
+lookup, and actions execution is then executed as per the description
 in the previous sections.
@@
 with incoming packets traversing through an eBPF parser+lookup program,
-an output action program, and the deparser program.
+and an output action program.
 Since the OVS eBPF is implemented based on tail calling these
@@
 actions processing, the datapath does not need to execute any modification to
-the packet contents, so the deparser program execution may be optimized out.
-
-To ensure invoking the deparser, the third experiment, set\_dst\_mac, involves
+the packet contents.
+
+The third experiment, set\_dst\_mac, involves
 altering packet data, specifically to modify the destination MAC address. In
-this case, the deparser writes stored metadata from actions processing back to
+this case, the OVS writes stored metadata from actions processing back to
 the packet. The performance in this experiment drops to $0.84$ Mpps. The drop
-of $0.28$ Mpps from baseline is the result of the deparser writing all known
-packet header metadata back to the packet. With a more intelligent deparser
+of $0.28$ Mpps from baseline is the result of writing all known
+packet header metadata back to the packet. With a more intelligent
 implementation, and by using recent improvements to the eBPF

### 2018-11-16 paper.tex
prepare for radio

@@
 \begin{document}
-\title{Bringing the Power of eBPF to Open vSwitch}
+\title{Faast Userspace OVS with AF\_XDP}
 \authorinfo{
@@
     % Alphabetical order.
-    \textrm{William Tu$^1$} & \textrm{Joe Stringer$^2$} & \textrm{Yifeng Sun$^1$} & \textrm{Yi-Hung Wei$^1$}\\
-    \textrm{u9012063@gmail.com} & \textrm{joe@cilium.io} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
+    \textrm{William Tu$^1$} & \textrm{Yifeng Sun$^1$} & \textrm{Yi-Hung Wei$^1$}\\
+    \textrm{u9012063@gmail.com} & \textrm{pkusunyifeng@gmail.com} & \textrm{yihung.wei@gmail.com} \\
   \end{matrix}
@@
 
-\section{OVS eBPF Datapath}
-\label{sec:ebpf}
-\myfigcross{ovsebpf}{5.5in}{The overall architecture of OVS eBPF datapath
-consists of multiple eBPF programs which are tail-called dynamically, maps
-which are shared between eBPF programs and userspace applications, and
-ovs-vswitchd as the management plane for all components.}
-
-\subsection{eBPF Configuration and Maps}
-The OVS eBPF datapath consists of one ELF-formatted object file which provides
-the full functionality of an OVS datapath. This object defines
-a set of maps and a set of eBPF programs which implement a subset of
-the datapath functionality.  To bootstrap, we load the eBPF program into
-the kernel using \verb+iproute+. One of the programs is marked
-within the ELF file to indicate that it should be
-attached to the hook point; the other programs are only executed via tail
-calls rooted in the eBPF hook point.  The ELF file also defines multiple persistent eBPF maps, which are
-pinned to the BPF filesystem ~\cite{bpfpmap}
-for sharing between different eBPF programs and \verb+ovs-vswitchd+.
-The OVS datapath requires the following eBPF maps:
-\begin{itemize}
-\item Flow key. This is the internal representation of the protocol headers
-    and metadata for the current packet being processed.
-\item Flow Table. This is a hash table whose key is the 32-bit hash value
-of the flow key, from both packet and its metadata,
-and value equals an array of actions to execute upon the flow.
-\item Stats Table. This is similar to the flow table, but rather than holding
-an array of actions to execute for the packet, it contains packet and byte
-statistics for the flow.
-\item Perf Ring Buffer~\cite{ebpfperf,ebpfperf2}. The perf event map
-allows an eBPF program to put user-defined data into a ring buffer which
-may be read from a userspace program. \verb+ovs-vswitchd+ memory maps
-this ring buffer to read packets and metadata from the eBPF program for
-flow miss upcall processing.
-\item Program Array. This map allows eBPF programs to tail call other eBPF
-programs. When the BPF loader inserts eBPF programs into the kernel, it assigns
-unique indexes for each program and stores these into the map.  At run time,
-an eBPF program will tail call another program by referring to an index within
-this map.
-\end{itemize}
-
-% locking and synchronization
-The OVS eBPF program is triggered by the TC ingress hook associated with a
-particular network device. Multiple instances of the same eBPF program
-may be triggered simultaneously on different cores which are receiving traffic
-from different network devices.  The eBPF maps, unless specified, have a single
-instance across all cores.  Access to map entries are protected by the kernel
-RCU~\cite{rcu} mechanism which makes it safe to read concurrently.  However,
-there are no built-in mechanisms to protect writers to the maps.
-For the flow table map, OVS avoids the race by ensuring that
-only \verb+ovs-vswitchd+ inserts or removes elements from the map from a single
-thread.  For flow statistics, atomic operations are used to avoid race
-conditions.  Other maps such as the flow key perf ring buffer maps use
-per-cpu instances to manage synchronization.
-
-\subsection{Header Parsing and Lookup}
-\label{parsing}
-When a packet arrives on the TC ingress hook, the OVS eBPF datapath begins
-executing a series of programs, beginning with the parser/lookup program
-as shown in Figure~\ref{ovsebpf:fig}.  The eBPF parser program
-consists of two components; standard protocol parsing and Linux-specific
-metadata parsing.  The protocol parsing is executed directly on the packet
-bytes based on standardized protocols, while the platform-specific metadata
-parsing must occur on the {\em context} provided by the eBPF environment.
-
-The resulting code will assemble the protocol headers and metadata,
-collectively known as the flow key. This flow key is then used to look for an
-entry in the flow table map, to get an array of actions to execute. If there is
-no entry in the flow table map, then the packet and the flow key will be
-written to the perf event map for further processing by \verb+ovs-vswitchd+.
-
-\subsection{Action Execution}
-\label{execution}
-When a lookup is successful the eBPF gets a list of actions to be executed,
-such as outputting the packet to a certain port, or pushing a VLAN tag. The
-list of actions is configured in \verb+ovs-vswitchd+ and may be a variable
-length depending on the desired network processing behaviour. For example, an
-L2 switch doing unknown broadcast sends packet to all its current ports.  The
-OVS datapath's actions is derived from the OpenFlow action specification and
-the OVSDB schema for \verb+ovs-vswitchd+.
-
-One might expect to intuitively write an eBPF program to iterate across the
-list of actions to execute, with each iteration of the loop dealing with a
-particular action in the list.  Unfortunately, this type of iteration implies
-dynamic loops, which are restricted within the eBPF forwarding model. Moreover,
-the variable number of actions also implies that there is no way to guarantee
-the bounded program size, which is limited to 4,096 eBPF bytecode instructions.
-
-To solve these challenges, we first break each type of action logic into an
-independent eBPF program and tail call from one eBPF action program to
-another, as shown in Figure~\ref{ovsebpf:fig}.  This alleviates
-the problem from having 4k instructions for the {\em entire} action list to 4k
-instructions {\em per action}.  Our proof-of-concept implementation shows that
-this limitation is sufficient for all existing actions.  Additionally, the
-design allows each action to be implemented and verified independently.
-Second, to solve the dynamic looping problem, we convert the variable length
-list of actions into a fixed length 32-element array.  As a result,
-flow table lookup always returns an array of 32 actions to be executed,
-and the LLVM compiler unrolls the loop to pass the BPF verifier.
-If a matching flow has less than 32 actions to execute, the rest of the
-actions is no-op, and we short-cut the no-op action execution. 
-If a matching flow has more than 32 actions, then the eBPF datapath delegates
-the execution to userspace datapath, i.e., the slow path. For many use cases,
-this is sufficient; there is room to further optimize this path in the future
-if common cases require more than 32 actions.
-
-Each action also requires certain action-specific metadata to execute.
-For example, an output action would require an {\em ifindex} to
-indicate the output interface, while a push\_vlan action needs the {\em VLAN ID}
-to be inserted into the VLAN tag.  To accommodate this, the array element not
-only contains the action to execute, but also the metadata required by the
-particular action. The size of each element must be big enough to hold the
-metadata for any action, as it is a fixed sized array. Future work may
-relax this requirement.
-
-Each eBPF program executes in an isolated environment. As a consequence
-of breaking the action list into individual eBPF action programs, some state needs to be
-transferred between the current eBPF program and the next.
-The state to transfer from one eBPF action program to another includes
-(1) The flow key, so that the currently executing eBPF action program can
-lookup the flow table, and (2) The action index in the actions list, so that
-the eBPF action program knows it is executing the $N$th element in the list,
-and at the end of its processing, to tail-call the $n+1$th action program.
-Currently eBPF could use either per-CPU maps as scratch buffers or the
-context's packet control block (\verb+cb+) to pass data between programs. In our
-design, we use one 32-bit value in the \verb+cb+ to keep the action
-index, and per-cpu maps to save the flow key.
-%As a result, there is additional overhead of map lookup per action.
-
-% an example to illustrate
-Figure~\ref{ovsebpf:fig} also demonstrates an example eBPF execution of a
-packet forwarded to port 2 as well as port mirroring to a VLAN port with VLAN ID
-100 at port 3.  Once the packet is parsed,
-the flow table lookup returns an action list of output:2, push\_vlan:100,
-output:3.  At the end of the parser+lookup program, it tail calls the eBPF
-output program as the first step to kick start action execution.
-The execution of this output program {\em overwrites} the caller's stack,
-so it has to look up the flow table map to retrieve the flow key,
-and execute the output action.  Once done, the output action increments the
-action index from 0 to 1, saves it in \verb+cb+, and tail-calls the next action
-program, which is push\_vlan.  The push\_vlan eBPF program again 
-looks up the flow table, and fetches the action metadata at index 1
-and executes the push VLAN action using a BPF helper function.
-The third output action follows the same procedure and finally sends the
-packet out to port 3.
-\saveforlater{Appendix A shows the eBPF helper functions used in
-each eBPF action program implementation.}
-
-\subsection{Flow Miss Upcall and Installation}
-One of the important tasks of the OVS datapath is to pass any packet that
-misses its flow table to the slow path to get instructions for
-further processing.  In the existing OVS {\em kernel} datapath implementation,
-the missed packet is sent to \verb+ovs-vswitchd+, which processes the packet,
-inserts a flow into the flow table, and re-injects the packet into the kernel
-datapath using the Linux generic netlink interface.  For the eBPF datapath, this
-design implies two requirements: (1) a way for the eBPF program to communicate
-with userspace, and (2) a mechanism for the userspace to re-insert packet into
-the eBPF program.
-
-% upcall: how ebpf work 
-To address the first requirement, we use the support for Linux perf ring
-buffers and the \verb+skb_event_output()+ eBPF helper function~\cite{ebpfperf}
-which allows the eBPF program to pass data to the userspace through the Linux perf
-ring buffer~\cite{ebpfperf}.  During miss upcall processing, the eBPF program
-will insert the full packet and the current flow key into the perf ring buffer.
-To receive the data from userspace, \verb+ovs-vswitchd+ runs a thread which
-maps the ring buffer to its virtual memory using mmap system call, and polls
-the buffer using the poll system call. If there is incoming data from the ring buffer, the
-thread is woken up, it reads from the buffer and processes the packet and
-metadata. The result of this processing will be inserted into the flow table
-map.
-% down call: how to re-insert packet
-To address the second requirement, we construct a dedicated Linux TAP device
-which also has the OVS eBPF datapath program attached to it.
-\verb+ovs-vswitchd+ sends the missed packet using an AF\_PACKET socket,
-triggering the underlying eBPF program for further processing.
-This program is very similar to the
-previously-used parser+lookup program, with minor changes. Specifically, this
-packet was originally received on one device, however when \verb+ovs-vswitchd+
-sends the packet on the TAP device, the eBPF program is triggered for the TAP
-device instead. So, the platform metadata for incoming port misidentifies the
-source as the TAP device. To ensure that the packet lookup occurs correctly,
-\verb+ovs-vswitchd+ prepends the port number to the beginning of the packet
-data, then when the eBPF program for the dedicated TAP device processes the
-packet, it will read this port number into the metadata, then strip this port
-from the packet. The resulting packet is identical to the originally-received
-packet, and now the metadata will match the metadata originally generated by
-the parser+lookup the first time the packet was received. The rest of the
-lookup, and actions execution is then executed as per the description
-in the previous sections.
-
-\subsection{Evaluation}\label{evaluation}
-To quantify the performance of the OVS eBPF datapath, we measure the packet
-forwarding rate in millions of packets per second (Mpps), using 64-byte
-packets under different forwarding scenarios. The hardware testbed
-consists of two Intel Xeon E5 2650 servers, each with an Intel 10GbE
-X540-AT2 dual port NIC, with the two ports of the Intel NIC on one server
-connected to the two ports on the identical NIC on the other server. The
-OVS eBPF datapath is installed on one server, acting as a bridge to
-forward packets from one port on the NIC to the other, and vice-versa.
-The other server acts as packet generating host, which runs DPDK
-packet-gen sending at the maximum packet rate of
-$14.88$Mpps. This server sends to one port to the target server, and
-receives the forwarded packets on the other port. All experiments use
-only one CPU core running Linux kernel 4.9-rc6.
-\begin{table}
-\centering
-\small
-\begin{tabular}{|c | c | c| c|}
-\hline
- {\bf Linux Bridge} & {\bf OVS Kernel } & {\bf Linux TC} & {\bf OVS eBPF}\\ \hline
- 1.6 Mpps & 1.4 Mpps & 1.9 Mpps & 1.12 Mpps\\ \hline
-\end{tabular}
-\caption{\footnotesize
-Baseline port-to-port forwarding rate using existing Linux subsystems and
-OVS kernel and eBPF.
-}
-\label{table:baseperf}
-\end{table}
-
-\textbf{Baseline Forwarding Performance.}
-We start by conducting a simple port-to-port packet forwarding experiment,
-i.e., receiving packets from one port and outputting to the other,
-using the Linux native bridge, Linux TC, and OVS kernel datapath, as shown in
-Table~\ref{table:baseperf}.
-The native Linux bridge is a simple L2 mac-learning switch with no
-programmability, showing the forwarding rate of $1.6$ Mpps.
-The OVS kernel module, which does additional flow key extraction,
-shows a slower forwarding rate of $1.4$ Mpps.
-For forwarding packets using TC, we have an eBPF program loaded
-into TC, with the program only having one BPF helper function call, the
-bpf\_skb\_redirect, that redirects the incoming packet from one interface
-to another.  Since TC accesses the incoming packets closest to the driver
-layer, it shows the highest performance of $1.9$ Mpps.
-Finally, we measure our proposed OVS eBPF forwarding speed,
-with incoming packets traversing through an eBPF parser+lookup program,
-and an output action program.
-Since the OVS eBPF is implemented based on tail calling these
-additional eBPF programs at TC, we observe the overhead of $0.78$ Mpps,
-a reduction from $1.9$ Mpps to $1.12$ Mpps.
-
-\textbf{Forwarding with action execution.}
-To further investigate the overhead,
-we program one additional action executed before the packet is
-forwarded to the other port.  Since the Linux bridge has no
-programmability, we only compare the OVS kernel datapath with the OVS eBPF
-datapath.  Table~\ref{table:ebpfperf} shows forwarding
-packets while executing the additional action type: hash, push\_vlan,
-set\_dst\_mac, and set\_gre\_tunnel, with the OVS kernel and eBPF datapaths.
-The forwarding performance of these results is bounded by the CPU cost.
-In the case of the hash and vlan actions, the NIC allows the processing
-to be offloaded from the CPU, resulting in minimal performance overhead for
-executing these actions.  As a result, the forwarding rate exhibits little
-to no overhead above the baseline.
-Moreover, for the hash and push\_vlan actions, due to the NIC offloading of
-actions processing, the datapath does not need to execute any modification to
-the packet contents.
-
-The third experiment, set\_dst\_mac, involves
-altering packet data, specifically to modify the destination MAC address. In
-this case, the OVS writes stored metadata from actions processing back to
-the packet. The performance in this experiment drops to $0.84$ Mpps. The drop
-of $0.28$ Mpps from baseline is the result of writing all known
-packet header metadata back to the packet. With a more intelligent
-implementation, and by using recent improvements to the eBPF
-API~\cite{dpa1,dpa2}, the performance gap compared with the native kernel
-implementation is expected to shrink.
-
-%The Set DST\_MAC result of $0.84$ Mpps makes the packet modification using the P4-generated deparser.
-Finally, the set\_get\_tunnel action experiment yielded $0.48$ Mpps, which
-represents the additional cost of tunnelling traffic; such traffic must
-traverse the Linux network stack twice, once for overlay traffic and
-once for underlay traffic. The eBPF result is also more expensive than
-the existing Linux implementation, but with a narrower performance gap
-than the earlier experiments. The majority of processing in this case occurs
-outside of OVS, so the overhead of the eBPF datapath has less effect.
-
-\begin{table}
-\centering
-\small
-\begin{tabular}{|c | p{1.4cm} | p{1.4cm}| c|}
-\hline
-{\bf Action} & {\bf eBPF DP} & {\bf Kernel DP} & {\bf Overhead}\\ \hline 
-hash & 1.12 & 1.34 & 16\%\\ \hline
-push\_vlan & 1.11 & 1.32 & 15\% \\ \hline
-set\_dst\_mac & 0.84 &  1.28 & 34\% \\ \hline
-%set\_dst\_mac (no parser) & 1.14 & -- & --\\ \hline
-set\_gre\_tunnel & 0.48 & 0.57 & 8\%\\ \hline 
-\end{tabular}
-\caption{\footnotesize Comparison of single core forwarding rate in Mpps
-with eBPF and Kernel Datapath, with the additional
-action executed before forwarding the packet to another port.}
-\label{table:ebpfperf}
-\end{table}
-
 \section{Userspace OVS with AF\_XDP}
@@
 
+\subsection{Lockless and Talkless Ring}
+Describe lockless and talkless single producer single consumer ring design.
+How to achieve lockless.
+How to achieve talkless.
+
+\subsection{Time Budget for Line Rate Processing}
+see Jesper's post
+Improving Linux networking performance
+
 \subsection{Performance Evaluation and Optimization}
@@
 
+\subsection{NAPI Busy Polling}
+Cite Eric's work, get NAPI id from XSK, pass NAPI id to busy poll.
+
+\subsection{Spectre and Meltdown Mitigation}
+remove indirect call, remove switch-statement with > 5 cases
+
 \subsubsection{Summary}
